### 2022-04-13 papers/main.tex
Add model description

% !TeX root = main.tex
\documentclass[number,times]{elsarticle}

\input{preamble.tex}

\begin{document}

\title{Description of a Deterministic Unit Commitment Model with Probabilistic Reserve constraints and Network Representation}

\maketitle

\newpage

\tableofcontents

\newpage

\section{Model description}

\subsection{Sets}

\begin{itemize}
    \item $G$ - Generators
    \item $G_n$ - Generators located at node $n$
    \item $GD$ - Dispatchable / thermal / conventional generators
    \item $GR$ - Renewable / variable generators
    \item $N$ - Nodes or buses in the network
    \item $B$ - Lines or branches
    \item $T$ - Time steps / intervals / slices
    \item $L^+$ - Upward reserve levels
    \item $L^-$ - Downward reserve levels
\end{itemize}

\subsection{Parameters}

\begin{itemize}
    \item $D_{nt}$ - Demand
    \item $PTDF_{nl}$ - Power Transfer Distribution Factor
    \item $AF_{gt}$ - Availability Factor
    \item $K_g$ - Capacity
    \item $P^{min}_g$ - Minimum power output
    \item $P^{max}_g$ - Maximum power output
    \item $D^{L^+}_{lt}$ - Upward reserve requirement
    \item $D^{L^-}_{lt}$ - Downward reserve requirement
\end{itemize}

\subsection{Variables}

All variables are positive.

\begin{itemize}
    \item $q_{gt}$ - Generation
    \item $\hat{q}_{gt}$ - Generation above the minimum stable operating point (0 in the case of renewables).
    \item $ls_{nt}$ - Load shedding
    \item $inj_{nt}$ - Node injection
    \item $f_{lt}$ - Branch flow
    \item $r^+_{gt}$ - Total upward reserve provision
    \item $r^-_{gt}$ - Total downward reserve provision
    \item $r^{L^+}_{glt}$ - Upward reserve provision for reserve level $l$
    \item $r^{L^-}_{glt}$ - Downward reserve provision for reserve level $l$
    \item $rs_{glt}$ - Upward reserve shedding for reserve level $l$
\end{itemize}

\subsection{Objective}

\begin{align}
    \min \quad & \sum_{t \in T} \sum_{g \in G} C^{var}_{g} \cdot \hat{q}_{gt} \nonumber                                          \\
               & \sum_{t \in T} \sum_{l \in L^+} P^{L^+} \cdot \sum_{g \in G} C^{var}_{g} \cdot r^{L^+}_{glt} + C^{shed} \cdot rs_{lt} \\
               & \sum_{t \in T} \sum_{l \in L^-} P^{L^-} \cdot C^{var}_{g} \cdot r^{L^-}_{glt} \nonumber
\end{align}

From the top line to the bottom, the costs are those of dispatching generators and activating upwards or downwards reserves.

Costs related to unit commitment, $z_{gt}$, have been omitted for brevity.

\subsection{Constraints}

The power balance:

\begin{equation}
    \sum_{g \in G_n} q_{gt} + ls_{nt} = D_{nt} + inj_{nt} \quad n \in N, \; t \in T
\end{equation}

Note the use of the set $G_n$ to only allow generators at node $n$ to contribute to the power balance.

Network constraints:

\begin{align}
    f_{bt} = \sum_{n \in N} PTDF_{nb} \cdot inj_{nt} & \quad b \in B, \; t \in T \\
    -F_b \leq f_{bt} \leq F_b                        & \quad b \in B, \; t \in T \\
    \sum_{n \in N} inj_{nt} = 0                      & \quad n \in N, \; t \in T \\
\end{align}

Constraints on generator output:

\begin{align}
    q_{gt} - r^{-}_{gt} \geq 0                    & \quad g \in GR, \; t \in T \\
    q_{gt} + r^{+}_{gt} \leq AF_{gt} \cdot K_g    & \quad g \in GR, \; t \in T \\
    q_{gt} - r^{-}_{gt} \geq P^{min} \cdot z_{gt} & \quad g \in GD, \; t \in T \\
    q_{gt} + r^{+}_{gt} \leq P^{max} \cdot z_{gt} & \quad g \in GD, \; t \in T
\end{align}

For brevity and clarity, constraints on ramping and minimum up and down times are omitted. In some sense, the commitment variable $z_{gt}$ can be thought of as given.

The constraints on reserve provision are as follows:

\begin{align}
    r^{+}_{gt} = \sum_{l \in L^+} r^{L^+}_{glt}           & \quad g \in G, \; t \in T   \\
    r^{-}_{gt} = \sum_{l \in L^-} r^{L^-}_{glt}           & \quad g \in G, \; t \in T   \\
    D^{L^+}_{gt} = \sum_{g \in G} r^{L^+}_{glt} + rs_{lt} & \quad l \in L^+, \; t \in T \\
    D^{L^-}_{gt} = \sum_{g \in G} r^{L^-}_{glt}           & \quad l \in L^-, \; t \in T \\
\end{align}

There are several matters to note here:

\begin{itemize}
    \item The operating reserve balance is performed over the entire network, not per node.
    \item The operating reserve balance is split into reserve levels. Higher reserve levels (values of $l$) are less likely to occur.
    \item It is possible to shed upward reserves, and this is more likely to occur for higher reserve levels.
\end{itemize}

The following constraints attempt to take network constraints into account (albeit very weakly):

\begin{align}
    \sum_{g \in G_n} r^{L^+}_{glt} = rinj^{L^+}_{nlt}                & \quad n \in N, l \in L^+, \; t \in T    \\
    \sum_{g \in G_n} r^{L^-}_{glt} = rinj^{L^-}_{nlt}                & \quad n \in N, l \in L^-, \; t \in T    \\
    rf^{L^+}_{blt} = \sum_{n \in N} PTDF_{nb} \cdot rinj^{L^+}_{nlt} & \quad b \in B, l \in L^+, \; t \in T    \\
    rf^{L^-}_{blt} = \sum_{n \in N} PTDF_{nb} \cdot rinj^{L^-}_{nlt} & \quad b \in B, l \in L^-, \; t \in T    \\
    -F_{b} \leq rf^{L^+}_{blt} \leq F_b                              & \quad b \in B, \; l \in L^+, \; t \in T \\
    -F_{b} \leq rf^{L^-}_{blt} \leq F_b                              & \quad b \in B, \; l \in L^-, \; t \in T \\
    \sum_{n \in N} rinj^{L^+}_{nt} = 0                               & \quad l \in L^+, \; n \in N, \; t \in T \\
    \sum_{n \in N} rinj^{L^-}_{nt} = 0                               & \quad l \in L^-, \; n \in N, \; t \in T \\
\end{align}

Since imbalances are aggregated across the network, a particular reserve level activation is not associated with an imbalance at the nodal level. The above constraints therefore enforce that for each reserve level $l$ and node $n$, there exists some combination of node injections and generator dispatches which would satisfy the network constraints AND the imbalance across the entire network. Given the formulation here, which uses reserve levels, i.e. quantiles, to represent forecast errors, it is difficult to come up with more stringent conditions.

\end{document}

### 2022-04-13 scripts/scenario_bias_check.jl
Investigate forecast scenario biases

@@ -1,0 +1,57 @@
+include(joinpath(@__DIR__, "..", "intro.jl"))
+sn = script_name(@__FILE__)
+mkrootdirs(joinpath(datadir("sims"), sn))
+
+opts_vec = options_3_days(sn)
+scens = load_scenarios(opts)
+
+# Only analyse particular source of uncertainty
+# buses = ["GOUY"]
+buses = String[]
+buses_str = reduce(*, buses .* "_")
+for scens_cp in vcat(scens, [Dict(k => v) for (k, v) in scens])
+    unc_srcs = reduce(*, collect(keys(scens_cp)) .* "_")[1:(end - 1)]
+    plt_name = length(buses) > 0 ? buses_str * unc_srcs : unc_srcs
+    @info "Plotting $unc_srcs..."
+
+    # Get net load forecast error per node
+    mult = Dict("Load" => -1, "Wind" => 1, "Solar" => 1)
+    net_load_forecast_error_dict = scenarios_2_net_load_forecast_error(
+        opts, scens_cp, mult
+    )
+    if length(buses) > 0
+        net_load_forecast_error_dict = Dict(
+            k => v for (k,v) in net_load_forecast_error_dict if k in buses
+        )
+    end
+
+    # Sum up uncertainty over entire network
+    total_NLFE = sum(v for (k, v) in net_load_forecast_error_dict)
+
+    # Plot timeseries and bounds
+    y = mean(total_NLFE; dims=2)[:]
+    ymax = [quantile(total_NLFE[i, :], 0.95) for i in 1:size(total_NLFE, 1)]
+    ymin = [quantile(total_NLFE[i, :], 0.05) for i in 1:size(total_NLFE, 1)]
+    Plots.plot(
+        y;
+        ribbon=(y .- ymin, ymax .- y),
+        ylab="Net load forecast error (90% CI)",
+        title=unc_srcs,
+    )
+    Plots.savefig(
+        joinpath(datadir("sims", sn, "uncertainty_sources=$(plt_name).png"))
+    )
+end
+
+# # Plot histograms
+# plt_vec = Plots.Plot[]
+# for i in 1:size(total_NLFE, 1)
+#     push!(plt_vec, histogram(total_NLFE[i, :] ./ 1e4))
+# end
+# Plots.plot(
+#     plt_vec...;
+#     layout=(4, 6),
+#     lab="",
+#     # xlims=(-2, 5),
+#     # xticks=-2:2:5
+# )

### 2022-04-13 src/process_input.jl

@@ -292,7 +292,7 @@ function powermodels_2_GEPPR(grid_data_path, grid_red_path)
                 "installedCapacity" =>
                     Dict(string(k) => Float64(v) for (k, v) in cap[t]), # [MW]
                 "averageGenerationCost" => 0.0,
-                "variableOperationAndMaintenanceCost" => 0.0, 
+                "variableOperationAndMaintenanceCost" => 0.0,
                 # Otherwise reserve activation doesn't work
             ) for t in res_names
         ),
@@ -585,10 +585,12 @@ end
 
 function scenarios_2_GEPPR(opts::Dict, scens)
     @unpack upward_reserve_levels, downward_reserve_levels = opts
-    
+
     # Get net load forecast error per node
     mult = Dict("Load" => -1, "Wind" => 1, "Solar" => 1)
-    net_load_forecast_error_dict = scenarios_2_net_load_forecast_error(opts, scens, mult)
+    net_load_forecast_error_dict = scenarios_2_net_load_forecast_error(
+        opts, scens, mult
+    )
 
     # Sum up uncertainty over entire network
     total_NLFE = sum(v for (k, v) in net_load_forecast_error_dict)
@@ -604,7 +606,9 @@ function scenarios_2_GEPPR(opts::Dict, scens)
     return D⁺, D⁻, P⁺, P⁻, Dmid⁺, Dmid⁻
 end
 
-function scenarios_2_net_load_forecast_error(opts::Dict, scens, mult)
+function scenarios_2_net_load_forecast_error(
+    opts::Dict, scens, mult; init_val=zeros(24, 1000)
+)
     pm = PowerModels.parse_file(grid_red_path)
 
     # Convolute scenarios to get total net load forecast error
@@ -625,9 +629,11 @@ function scenarios_2_net_load_forecast_error(opts::Dict, scens, mult)
         )
 
         # Net load is negative -> upward reserves are activated
-        # Net load is positive -> downward reserves are activated 
-        net_load_forecast_error_dict[name] = sum(
-            scen_mat * mult[source] for (source, scen_mat) in bus_scen_mat
+        # Net load is positive -> downward reserves are activated
+        net_load_forecast_error_dict[name] = reduce(
+            +,
+            scen_mat * mult[source] for (source, scen_mat) in bus_scen_mat;
+            init=init_val,
         )
     end
     return net_load_forecast_error_dict
[…]

### 2022-04-14 scripts/DUCPR_reserve_shedding_sensitivity.jl
Include reserve level network cons

@@ -3,7 +3,7 @@ sn = script_name(@__FILE__)
 
 opts_vec = options_3_days(sn)
 opts_vec = map(
-    opts -> (opts["vars_2_save"] = [:z, :q, :ls, :rsl⁺]; opts), opts_vec
+    opts -> (opts["vars_2_save"] = [:z, :q, :ls, :rsL⁺]; opts), opts_vec
 )
 opts_vec = vcat(
     [

### 2022-04-14 src/GEPPR.jl
Include reserve level network cons

@@ -3,7 +3,8 @@ include(srcdir("util.jl"))
 
 ### UTIL
 function param_and_config(opts::Dict)
-    @unpack optimization_horizon, rolling_horizon, include_downward_reserves = opts
+    @unpack optimization_horizon, rolling_horizon, include_downward_reserves =
+        opts
     is_linear = (opts["unit_commitment_type"] == "none")
     GEPPR_dir = datadir("pro", "GEPPR")
     configFiles =
@@ -111,8 +112,7 @@ run_GEPPR(opts_vec) = [
         run_GEPPR(opts)
     catch
         @warn "Optimisation failed"
-    end
-    for opts in opts_vec
+    end for opts in opts_vec
 ]
 
 function GEPPR.load_GEP(opts::Dict, path::String)
@@ -147,8 +147,12 @@ end
 
 function apply_operating_reserves!(gep::GEPM, opts::Dict)
     @unpack operating_reserves_type,
-    operating_reserves_sizing_type, upward_reserve_levels,
-    downward_reserve_levels = opts
+    operating_reserves_sizing_type,
+    upward_reserve_levels,
+    downward_reserve_levels,
+    upward_reserve_levels_included_in_redispatch,
+    downward_reserve_levels_included_in_redispatch = opts
+
     operating_reserves_type == "none" && return gep
     @assert operating_reserves_type == "probabilistic"
     @assert operating_reserves_sizing_type == "given"
@@ -164,6 +168,12 @@ function apply_operating_reserves!(gep::GEPM, opts::Dict)
     modify_parameter!(gep, "reserveSizingType", "given")
     modify_parameter!(gep, "includeReserveActivationCosts", true)
     modify_parameter!(gep, "includeDownwardReserves", true)
+    modify_parameter!(
+        gep, "upwardReserveLevelsIncludedInNetworkRedispatch", 1:10
+    )
+    modify_parameter!(
+        gep, "downwardReserveLevelsIncludedInNetworkRedispatch", 1:10
+    )
     L⁺ = gep[:I, :sets, :L⁺] = 1:upward_reserve_levels
     L⁻ = gep[:I, :sets, :L⁻] = 1:downward_reserve_levels
     ORBZ = GEPPR.get_set_of_operating_reserve_balancing_zones(gep)
@@ -190,7 +200,7 @@ end
 
 function constrain_reserve_shedding!(gep::GEPM, opts::Dict)
[…]

### 2022-04-14 src/opts.jl

@@ -14,7 +14,9 @@ function options(pairs...)
         "optimization_horizon" => [1,48],
         "rolling_horizon" => false,
         "upward_reserve_levels" => 10,
+        "upward_reserve_levels_included_in_redispatch" => 1:10,
         "downward_reserve_levels" => 10,
+        "downward_reserve_levels_included_in_redispatch" => 1:10,
         "include_downward_reserves" => true,
         "reserve_shedding_limit" => 1.0
     )

### 2022-04-20 src/process_input.jl
Fix coverage and net load calc

@@ -611,7 +611,7 @@ function scenarios_2_GEPPR(opts::Dict, scens)
         @unpack upward_reserve_levels, downward_reserve_levels = opts
 
         # Get net load forecast error per node
-        mult = Dict("Load" => -1, "Wind" => 1, "Solar" => 1)
+        mult = Dict("Load" => 1, "Wind" => -1, "Solar" => -1)
         net_load_forecast_error_dict = scenarios_2_net_load_forecast_error(
             opts, scens, mult
         )
@@ -624,7 +624,7 @@ function scenarios_2_GEPPR(opts::Dict, scens)
             transpose(total_NLFE);
             n_up=upward_reserve_levels,
             n_down=downward_reserve_levels,
-            coverage=10, # Number of scenarios ignored on tail ends
+            coverage=Int(round(size(total_NLFE, 2)*0.01)), # Number of scenarios ignored on tail ends
         )
 
         @save eval(file_name) D⁺ D⁻ P⁺ P⁻ Dmid⁺ Dmid⁻

### 2022-04-20 src/GEPPR.jl
Allow copperplate model

@@ -45,6 +45,9 @@ function param_and_config(opts::Dict)
             "Optimisation length too long for unit commitment model, consider setting `rolling_horizon = true`.",
         )
     end
+    if opts["copperplate"] == true
+        param["forceCopperPlateModel"] = true
+    end
     return configFiles, param
 end
 

### 2022-04-20 src/opts.jl
Allow copper plate option

@@ -18,7 +18,8 @@ function options(pairs...)
         "downward_reserve_levels" => 10,
         "downward_reserve_levels_included_in_redispatch" => [],
         "include_downward_reserves" => true,
-        "reserve_shedding_limit" => 1.0
+        "reserve_shedding_limit" => 1.0,
+        "copperplate" => false,
     )
     for (k,v) in pairs
         opts[k] = v

### 2022-04-20 scripts/scenario_bias_check.jl
Investigating feasability issues

@@ -3,55 +3,47 @@ sn = script_name(@__FILE__)
 mkrootdirs(joinpath(datadir("sims"), sn))
 
 opts_vec = options_3_days(sn)
-scens = load_scenarios(opts)
 
 # Only analyse particular source of uncertainty
-# buses = ["GOUY"]
-buses = String[]
-buses_str = reduce(*, buses .* "_")
-for scens_cp in vcat(scens, [Dict(k => v) for (k, v) in scens])
-    unc_srcs = reduce(*, collect(keys(scens_cp)) .* "_")[1:(end - 1)]
-    plt_name = length(buses) > 0 ? buses_str * unc_srcs : unc_srcs
-    @info "Plotting $unc_srcs..."
+function plot_net_load_forecast_uncertainty(opts, scens, buses=String[])
+    scens = load_scenarios(opts)
 
-    # Get net load forecast error per node
-    mult = Dict("Load" => -1, "Wind" => 1, "Solar" => 1)
-    net_load_forecast_error_dict = scenarios_2_net_load_forecast_error(
-        opts, scens_cp, mult
-    )
-    if length(buses) > 0
-        net_load_forecast_error_dict = Dict(
-            k => v for (k,v) in net_load_forecast_error_dict if k in buses
+    # buses = ["GOUY"]
+    buses = String[]
+    buses_str = reduce(*, buses .* "_")
+    for scens_cp in vcat(scens, [Dict(k => v) for (k, v) in scens])
+        unc_srcs = reduce(*, collect(keys(scens_cp)) .* "_")[1:(end - 1)]
+        plt_name = length(buses) > 0 ? buses_str * unc_srcs : unc_srcs
+        @info "Plotting $unc_srcs..."
+
+        # Get net load forecast error per node
+        mult = Dict("Load" => -1, "Wind" => 1, "Solar" => 1)
+        net_load_forecast_error_dict = scenarios_2_net_load_forecast_error(
+            opts, scens_cp, mult
+        )
+        if length(buses) > 0
+            net_load_forecast_error_dict = Dict(
+                k => v for (k,v) in net_load_forecast_error_dict if k in buses
+            )
+        end
+
+        # Sum up uncertainty over entire network
+        total_NLFE = sum(v for (k, v) in net_load_forecast_error_dict)
+
+        # Plot timeseries and bounds
[…]

### 2022-04-20 scripts/infeasibility_debug.jl
Investigating feasability issues

@@ -1,0 +1,26 @@
+include(joinpath(@__DIR__, "..", "intro.jl"))
+sn = script_name(@__FILE__)
+
+# No network, no UC, no network redispatch
+opts_vec = options_3_days(sn)
+opts_vec = [
+    merge(opts, Dict(
+        "reserve_shedding_limit" => 0.0,
+        "unit_commitment_type" => "none",
+        "vars_2_save" => [:z, :q, :ls, :rsL⁺],
+        "copperplate" => true
+    ))
+]
+gep_vec = run_GEPPR(opts_vec)
+# Day 41 is still broken!!! Hmmmm...
+
+# No UC
+opts_vec = options_3_days(sn)
+opts_vec = [
+    merge(opts, Dict(
+        "reserve_shedding_limit" => 0.0,
+        "unit_commitment_type" => "none",
+        "vars_2_save" => [:z, :q, :ls, :rsL⁺],
+    ))
+]
+gep_vec = run_GEPPR(opts_vec)

### 2022-04-20 src/GEPPR.jl

@@ -226,11 +226,15 @@ function constrain_reserve_shedding!(gep::GEPM, opts::Dict)
     )
     L⁺ = GEPPR.get_set_of_upward_reserve_levels(gep)
 
+    norm_factor = Dict(
+        t => sum(D⁺[z, l, Y[1], P[1], t] for z in ORBZ, l in L⁺) for t in T
+    )
     gep[:M, :constraints, :reserveSheddingLimit] = @constraint(
         gep.model,
         [z = ORBZ, y in Y, p in P, t = T],
         sum(rsL⁺[n, l, y, p, t] for n in ORBZ2N[z], l in L⁺) /
-        sum(D⁺[z, l, y, p, t] for z in ORBZ, l in L⁺) <= reserve_shedding_limit
+        (iszero(norm_factor[t]) ? 1.0 : norm_factor[t]) <=
+            reserve_shedding_limit
     )
     return gep
 end
@@ -277,10 +281,8 @@ function save_gep_for_security_analysis(gep::GEPM, path::String)
                 (g, n) => Dict("q" => q[(g, n), y, p, atval(t, typeof(q))])
                 for (g, n) in GRN
             ),
-            "load_shed" => Dict(
-                n => ls[n,y,p,atval(t, typeof(ls))]
-                for n in N
-            )
+            "load_shed" =>
+                Dict(n => ls[n, y, p, atval(t, typeof(ls))] for n in N),
         )
     end
     @save eval(path) UC_results

### 2022-04-20 src/opts.jl

@@ -20,6 +20,8 @@ function options(pairs...)
         "include_downward_reserves" => true,
         "reserve_shedding_limit" => 1.0,
         "copperplate" => false,
+        "vars_2_save" => Symbol[],
+        "exprs_2_save" => Symbol[],
     )
     for (k,v) in pairs
         opts[k] = v

### 2022-04-21 scripts/infeasibility_debug.jl
Saving

@@ -1,26 +1,113 @@
 include(joinpath(@__DIR__, "..", "intro.jl"))
 sn = script_name(@__FILE__)
+mkrootdirs(simsdir(sn))
 
-# No network, no UC, no network redispatch
+# Linear, no reserves, no network - simplest case
 opts_vec = options_3_days(sn)
+map(opts -> rm(opts["save_path"]; force=true, recursive=true), opts_vec)
 opts_vec = [
-    merge(opts, Dict(
-        "reserve_shedding_limit" => 0.0,
-        "unit_commitment_type" => "none",
-        "vars_2_save" => [:z, :q, :ls, :rsL⁺],
-        "copperplate" => true
-    ))
+    merge(
+        opts,
+        Dict(
+            "unit_commitment_type" => "none",
+            "operating_reserves_type" => "none",
+            "copperplate" => true,
+        ),
+    ) for opts in opts_vec
 ]
 gep_vec = run_GEPPR(opts_vec)
+Plots.savefig(
+    plot_dispatch(gep_vec[2], 1), simsdir(sn, "day_5_lin_no_reserve_no_net.png")
+)
+
+# Linear, no reserves, with network
+opts_vec = options_3_days(sn)
+map(opts -> rm(opts["save_path"]; force=true, recursive=true), opts_vec)
+opts_vec = [
+    merge(
+        opts,
+        Dict(
+            "unit_commitment_type" => "none",
+            "operating_reserves_type" => "none",
+        ),
+    ) for opts in opts_vec
+]
+gep_vec = run_GEPPR(opts_vec)
+Plots.savefig(
+    plot_dispatch(gep_vec[2], 1), simsdir(sn, "day_5_lin_no_reserve.png")
+)
+
+# Linear, no reserves, with network AND STORAGE
+opts_vec = options_3_days(sn)
+map(opts -> rm(opts["save_path"]; force=true, recursive=true), opts_vec)
+opts_vec = [
+    merge(
+        opts,
+        Dict(
+            "unit_commitment_type" => "none",
+            "operating_reserves_type" => "none",
+            "include_storage" => true,
+        ),
+    ) for opts in opts_vec
+]
+gep_vec = run_GEPPR(opts_vec)
+Plots.savefig(
+    plot_dispatch(gep_vec[2], 1), simsdir(sn, "day_5_lin_no_reserve_w_storage.png")
[…]

### 2022-04-21 scripts/process_input.jl

@@ -8,7 +8,7 @@ process_belderbos_data(grid_path)
 # Process PowerModels into GEPPR format
 powermodels_2_GEPPR(grid_path, grid_red_path)
 
-# Run a linear operational model
+# Run a linear operational model for the entire year
 opts = options(
     "include_storage" => true,
     "unit_commitment_type" => "none",

### 2022-04-21 src/util.jl
Saving

@@ -23,6 +23,9 @@ end
 
 scenario_id(opts::Dict) = parse(Int, split(opts["load_scenario_data_paths"], "_")[end])
 
+# DIRs
+simsdir(args...) = datadir("sims", args...)
+
 # PATHS
 grid_path = datadir("pro", "grid.json")
 grid_red_path = datadir("pro", "GEPPR", "grid_red.json")

### 2022-04-21 scripts/process_input.jl
Modify day choice

@@ -20,7 +20,26 @@ opts = options(
 gep = run_GEPPR(opts)
 
 # Identify 3 days to investigate - no load shedding at all, almost load shedding and load shedding
-days_to_run_models_on(gep, "days_for_analysis.csv")
+days, TVec = days_to_run_models_on(gep, "days_for_analysis.csv")
+
+# Check the dispatch at the network level for each case
+for T in TVec
+    Plots.savefig(
+        plot_dispatch(gep, 1; T=T, plot_state_of_charge=false),
+        simsdir(sn, "dispatch_T=$T.png"),
+    )
+end
+
+# Plot daily load shedding for entire year
+ls = sum(
+    reshape(
+        dropdims(sum(gep[:loadShedding].data; dims=(1, 2, 3)); dims=(1, 2, 3)),
+        24,
+        :,
+    );
+    dims=1,
+)[:]
+Plots.plot(ls, xlab="Day", ylab="Load shedding [MW]", lab="")
 
 # Save storage dispatch as a node injection
 storage_dispatch_2_node_injection(gep, grid_path, grid_wo_store_path)

### 2022-04-21 src/plot.jl

@@ -132,7 +132,7 @@ function plot_dispatch(
             xlab="Time [h]",
             legend=:outerright,
             xlims=(T[1], T[end] + 1),
-            ylims=(0, maximum(D) * 1.1),
+            # ylims=(0, maximum(D) * 1.1),
         )
 
         StatsPlots.groupedbar!(

### 2022-04-21 src/analysis.jl
Modify day choice

@@ -49,13 +49,19 @@ function days_to_run_models_on(gep::GEPM, filename::String)
     dem_net_sum = [sum(v[i] for (k, v) in dem_net) for i in 1:8_760]
     dem_net_day = reshape(dem_net_sum, 24, :)
     dem_net_max_day = maximum(dem_net_day; dims=1)
+
+    # The days
     day_no_scarce = findmin(dem_net_max_day)[2][2]
-    day_some_scarce = findfirst(dem_net_max_day .> -1000)[2]
+    day_even_less_scarce = findfirst(x -> (-10000 < x < -6000), dem_net_max_day)[2]
+    day_less_scarce = findfirst(x -> (-6000 < x < -2000), dem_net_max_day)[2]
+    day_some_scarce = findfirst(x -> (-2000 < x < -1000), dem_net_max_day)[2]
     day_scarce = findmax(dem_net_max_day)[2][2]
     df = DataFrame(;
-        type=["No scarcity", "Scarcity in redispatch", "Scarcity in day ahead"],
+        type=["Max RAM", "RAM>4000MW", "RAM>1000MW", "RAM>500MW", "Min RAM"],
         days=[
             day_no_scarce
+            day_even_less_scarce
+            day_less_scarce
             day_some_scarce
             day_scarce
         ]
@@ -64,5 +70,5 @@ function days_to_run_models_on(gep::GEPM, filename::String)
         (i - 1) * 24 + 1:i * 24 for i in df[:,"days"]
     ]
     CSV.write(datadir("pro", filename), df)
-    return day_no_scarce, day_some_scarce, day_scarce
+    return df[:,"days"], df[:,"timesteps"]
 end

### 2022-04-27 papers/main.tex
Investigating feasability issues
Update model description

@@
 
-All variables are positive.
+All variables are positive apart from injetion variables which may also be negative.
 
@@
     \item $r^{L^-}_{glt}$ - Downward reserve provision for reserve level $l$
-    \item $rs_{glt}$ - Upward reserve shedding for reserve level $l$
+    \item $rs_{nlt}$ - Upward reserve shedding for reserve level $l$
+    \item $rc_{nlt}$ - Downward reserve provided by day ahead load shedding for reserve level $l$
+    \item $rinj_{nlt}^{L^+}$ - Possible node injection due to activation of upward reserve level $l$
+    \item $rinj_{nlt}^{L^-}$ - Possible node injection due to activation of downward reserve level $l$
+    \item $rf_{nbt}^{L^+}$ - Possible branch flow due to activation of upward reserve level $l$
+    \item $rf_{nbt}^{L^-}$ - Possible branch flow due to activation of downward reserve level $l$
 \end{itemize}
@@
     \min \quad & \sum_{t \in T} \sum_{g \in G} C^{var}_{g} \cdot \hat{q}_{gt} \nonumber                                                                  \\
-               & \sum_{t \in T} \sum_{l \in L^+} P^{L^+} \cdot \sum_{g \in G} C^{var}_{g} \cdot r^{L^+}_{glt} + C^{shed} \cdot rs_{lt} \\
-               & \sum_{t \in T} \sum_{l \in L^-} P^{L^-} \cdot C^{var}_{g} \cdot r^{L^-}_{glt} \nonumber
+               & + \sum_{t \in T} \sum_{l \in L^+} \sum_{n \in N} P^{L^+} \cdot \sum_{g \in G} C^{var}_{g} \cdot r^{L^+}_{glt} + C^{shed} \cdot rs_{nlt} \\
+               & - \sum_{t \in T} \sum_{l \in L^-} \sum_{n \in N} P^{L^-} \cdot C^{var}_{g} \cdot r^{L^-}_{gnlt} + C^{shed} \cdot rc_{nlt} \nonumber
 \end{align}
@@
 
-Note the use of the set $G_n$ to only allow generators at node $n$ to contribute to the power balance.
+Note the use of the set $G_n$ to only allow generators at node $n$ to contribute to the power balance. Another way of describing this would have been through an incidence matrix.
 
@@
 
-For brevity and clarity, constraints on ramping and minimum up and down times are omitted. In some sense, the commitment variable $z_{gt}$ can be thought of as given.
+For brevity and clarity, constraints on ramping and minimum up and down times are omitted.
 
@@
 \begin{align}
-    r^{+}_{gt} = \sum_{l \in L^+} r^{L^+}_{glt}           & \quad g \in G, \; t \in T   \\
-    r^{-}_{gt} = \sum_{l \in L^-} r^{L^-}_{glt}           & \quad g \in G, \; t \in T   \\
-    D^{L^+}_{gt} = \sum_{g \in G} r^{L^+}_{glt} + rs_{lt} & \quad l \in L^+, \; t \in T \\
-    D^{L^-}_{gt} = \sum_{g \in G} r^{L^-}_{glt}           & \quad l \in L^-, \; t \in T \\
+    D^{L^+}_{lt} = \sum_{g \in G} r^{L^+}_{glt} + \sum_{n \in N} rs_{nlt} & \quad l \in L^+, \; t \in T \\
+    D^{L^-}_{lt} = \sum_{g \in G} r^{L^-}_{glt} + \sum_{n \in N} rc_{nlt} & \quad l \in L^-, \; t \in T \\
+    \sum_{l \in L^-} rc_{nlt} \leq ls_{nt}                                 & \quad n \in N, \; t \in T \\
+    r^{+}_{gt} = \sum_{l \in L^+} r^{L^+}_{gnlt}                           & \quad g \in G, \; t \in T   \\
+    r^{-}_{gt} = \sum_{l \in L^-} r^{L^-}_{gnlt}                           & \quad g \in G, \; t \in T 
 \end{align}
@@
     \item It is possible to shed upward reserves, and this is more likely to occur for higher reserve levels.
+This model is therefore able to make a tradeoff between day ahead adequacy and real time operational security, albeit crudely.
+    \item Shedding load in day ahead allows additional downward reserves to be provided through the variable $rc_{nlt}$.
+Implicitly this assumes that load can be `activated' in real time to provide downward reserves.
 \end{itemize}
@@
 \begin{align}
-    \sum_{g \in G_n} r^{L^+}_{glt} = rinj^{L^+}_{nlt}                & \quad n \in N, l \in L^+, \; t \in T    \\
-    \sum_{g \in G_n} r^{L^-}_{glt} = rinj^{L^-}_{nlt}                & \quad n \in N, l \in L^-, \; t \in T    \\
+    \sum_{g \in G_n} r^{L^+}_{glt} + rs_{nlt} = rinj^{L^+}_{nlt}                & \quad n \in N, l \in L^+, \; t \in T    \\
+    \sum_{g \in G_n} r^{L^-}_{glt} + rc_{nlt} = rinj^{L^-}_{nlt}                & \quad n \in N, l \in L^-, \; t \in T    \\
     rf^{L^+}_{blt} = \sum_{n \in N} PTDF_{nb} \cdot rinj^{L^+}_{nlt} & \quad b \in B, l \in L^+, \; t \in T    \\
@@
     \sum_{n \in N} rinj^{L^+}_{nt} = 0                               & \quad l \in L^+, \; n \in N, \; t \in T \\
-    \sum_{n \in N} rinj^{L^-}_{nt} = 0                               & \quad l \in L^-, \; n \in N, \; t \in T \\
+    \sum_{n \in N} rinj^{L^-}_{nt} = 0                               & \quad l \in L^-, \; n \in N, \; t \in T
 \end{align}

### 2022-04-27 scripts/DUCPR_reserve_shedding_sensitivity.jl
Update model description

@@ -10,7 +10,7 @@ opts_vec = vcat(
                 "reserve_shedding_limit" => v,
                 "save_path" => joinpath(opts["save_path"], "RSL=$v"),
                 "time_out" => 600,
-                "vars_2_save" => [:z, :q, :ls, :rsL⁺],
+                "vars_2_save" => [:z, :q, :ls, :rsL⁺, :rsL⁻],
                 "exprs_2_save" => [:loadShedding]
             ),
         ) for v in 1:-0.1:0, opts in opts_vec

### 2022-04-28 scripts/DUCPR_reserve_shedding_sensitivity.jl
Change scenario_id -> month_day

@@ -33,7 +33,7 @@ for i in 1:length(opts_vec)
     gep_vec[i] == nothing && continue
     apply_operating_reserves!(gep_vec[i], opts_vec[i])
 end
-sid_vec = [scenario_id(opts_vec[i]) for i in 1:length(opts_vec)]
+sid_vec = [month_day(opts_vec[i]) for i in 1:length(opts_vec)]
 rsl_vec = [opts_vec[i]["reserve_shedding_limit"] for i in 1:length(opts_vec)]
 ls = Dict(
     (sid_vec[i], rsl_vec[i]) => try
@@ -69,8 +69,9 @@ rsL⁺ = Dict(
         NaN
     end for i in 1:length(opts_vec)
 )
-x = reshape([sum(ls[sid_vec[i], rsl_vec[i]]) for i in 1:length(opts_vec)], :, 3)
-y = reshape([rsL⁺[sid_vec[i], rsl_vec[i]] for i in 1:length(opts_vec)], :, 3)
+nd = length(options_diff_days(sn))
+x = reshape([sum(ls[sid_vec[i], rsl_vec[i]]) for i in 1:length(opts_vec)], :, nd)
+y = reshape([rsL⁺[sid_vec[i], rsl_vec[i]] for i in 1:length(opts_vec)], :, nd)
 Plots.plot(
     x,
     y;
@@ -78,5 +79,5 @@ Plots.plot(
     markershape=:star5,
     xlab="Day ahead load shedding [MWh]",
     ylab="Reserve shedding [MWh]",
-    lab=["208" "5" "41"],
+    lab=["214" "210" "136" "309"],
 )

### 2022-04-28 src/util.jl

@@ -21,7 +21,7 @@ function Base.parse(::Type{UnitRange{T}}, str::AbstractString) where T<:Integer
     return parse(T, split_str[1]), parse(T, split_str[2])
 end
 
-scenario_id(opts::Dict) = parse(Int, split(opts["load_scenario_data_paths"], "_")[end])
+month_day(opts::Dict) = split(opts["load_scenario_data_paths"], "1000SC_BELDERBOS_load_")[end]
 
 # DIRs
 simsdir(args...) = datadir("sims", args...)

### 2022-04-28 src/process_input.jl
Change scenario_id -> month_day

@@ -579,7 +579,7 @@ function load_scenarios(
 end
 
 function load_scenarios(opts::Dict; for_GEPPR=false)
-    scen_id = scenario_id(opts)
+    scen_id = month_day(opts)
 
     # If scenarios are being loaded for GEPPR, this step may be superfluous
     GEPPR_file_name = datadir("scenarios", "GEPPR_frmt_month=$(scen_id).jld2")
@@ -608,7 +608,7 @@ function load_scenarios(opts::Dict; for_GEPPR=false)
 end
 
 function scenarios_2_GEPPR(opts::Dict, scens)
-    scen_id = scenario_id(opts)
+    scen_id = month_day(opts)
     file_name = datadir("scenarios", "GEPPR_frmt_month=$(scen_id).jld2")
     mkrootdirs(dirname(file_name))
     if isfile(file_name)

### 2022-05-03 scripts/DUCPR_reserve_shedding_sensitivity.jl
Modify network redispatch constraint

@@ -10,12 +10,13 @@ opts_vec = vcat(
                 "reserve_shedding_limit" => v,
                 "save_path" => joinpath(opts["save_path"], "RSL=$v"),
                 "time_out" => 600,
-                "upward_reserve_levels_included_in_redispatch" => [10],
-                "downward_reserve_levels_included_in_redispatch" => [10],
+                "upward_reserve_levels_included_in_redispatch" => [],
+                "downward_reserve_levels_included_in_redispatch" => [],
                 "vars_2_save" => [:z, :q, :ls, :rsL⁺, :rsL⁻, :e],
                 "exprs_2_save" => [:loadShedding]
             ),
-        ) for v in 1:-0.1:0, opts in opts_vec
+        ) for v in 0.1:-0.02:0, opts in opts_vec
+        # ) for v in [0.02], opts in opts_vec
     ]...,
 )
 gep = run_GEPPR(opts_vec[1])
@@ -33,7 +34,7 @@ map(
 
 # Plot
 for i in 1:length(opts_vec)
-    gep_vec[i] == nothing && continue
+    gep_vec[i] === nothing && continue
     apply_operating_reserves!(gep_vec[i], opts_vec[i])
 end
 sid_vec = [month_day(opts_vec[i]) for i in 1:length(opts_vec)]

### 2022-05-03 src/GEPPR.jl
Security analysis uses keys from PM dict

@@ -289,9 +289,9 @@ function save(gep::GEPM, opts::Dict)
 end
 
 """
-    save_gep_for_security_analysis(gep::GEPM)
+    save_gep_for_security_analysis(gep::GEPM, path::String)
 
-Saves data in the format: hour -> generator (with associated bus) -> value
+Saves data in the format: hour -> generator (with associated bus) -> values / name / bus etc.
 """
 function save_gep_for_security_analysis(gep::GEPM, path::String)
     q = gep[:q]
@@ -303,34 +303,38 @@ function save_gep_for_security_analysis(gep::GEPM, path::String)
     GDN = GEPPR.get_set_of_nodal_dispatchable_generators(gep)
     GRN = GEPPR.get_set_of_nodal_intermittent_generators(gep)
     STN = GEPPR.get_set_of_nodal_storage_technologies(gep)
+    TN2idx = tech_node_to_idx(gep)
     y, p = first.([Y, P])
     for t in T
         UC_results[t] = Dict(
-            "gen" => [
-                Dict(
-                    "bus" => n,
-                    "name" => g,
-                    "q" => q[(g, n), y, p, atval(t, typeof(q))],
-                    "z" => z[(g, n), y, p, atval(t, typeof(z))],
-                ) for (g, n) in GDN
-            ],
-            "res" => [
-                Dict(
-                    "bus" => n,
-                    "name" => g,
-                    "q" => q[(g, n), y, p, atval(t, typeof(q))],
-                ) for (g, n) in GRN
-            ],
-            "store" => [
-                Dict(
+            "gen" => Dict(
+                TN2idx[(g,n)] => 
+                    Dict(
+                        "bus" => n,
+                        "name" => g,
+                        "q" => q[(g, n), y, p, atval(t, typeof(q))],
+                        "z" => z[(g, n), y, p, atval(t, typeof(z))],
+                    )
+                for (g, n) in GDN
+            ),
+            "res" => Dict(
+                TN2idx[(g,n)] => Dict(
+                        "bus" => n,
+                        "name" => g,
+                        "q" => q[(g, n), y, p, atval(t, typeof(q))],
+                    )
+                for (g, n) in GRN
+            ),
+            "store" => Dict(
+                TN2idx[(st,n)] => Dict(
                     "bus" => n,
                     "name" => st,
[…]

### 2022-05-04 scripts/DUCPR_reserve_shedding_sensitivity.jl
Save storage dispatch, also look at net redispatch

@@ -8,14 +8,14 @@ opts_vec = vcat(
             opts,
             Dict(
                 "reserve_shedding_limit" => v,
-                "save_path" => joinpath(opts["save_path"], "RSL=$v"),
+                "save_path" => joinpath(opts["save_path"], "RSL=$(v)_L=$L"),
                 "time_out" => 600,
-                "upward_reserve_levels_included_in_redispatch" => [],
-                "downward_reserve_levels_included_in_redispatch" => [],
-                "vars_2_save" => [:z, :q, :ls, :rsL⁺, :rsL⁻, :e],
+                "upward_reserve_levels_included_in_redispatch" => L,
+                "downward_reserve_levels_included_in_redispatch" => L,
+                "vars_2_save" => [:z, :q, :ls, :rsL⁺, :rsL⁻, :e, :sc, :sd],
                 "exprs_2_save" => [:loadShedding]
             ),
-        ) for v in 0.1:-0.02:0, opts in opts_vec
+        ) for v in 0.1:-0.02:0, L in [Int[],1:10], opts in opts_vec
         # ) for v in [0.02], opts in opts_vec
     ]...,
 )

### 2022-05-04 src/GEPPR.jl
Save storage dispatch, also look at net redispatch

@@ -133,9 +133,6 @@ function GEPPR.load_GEP(opts::Dict, path::String)
     return load_GEP(path)
 end
 
-# TODO: alternative run_GEPPR which keeps GEPPR the same in between runs
-# TODO: so as to save model load time
-
 function apply_initial_commitment!(gep::GEPM, opts::Dict)
     @unpack save_path, optimization_horizon, initial_commitment_data_path = opts
     isempty(initial_commitment_data_path) && return nothing

### 2022-05-04 papers/main.tex

@@
     \item $rf_{nbt}^{L^-}$ - Possible branch flow due to activation of downward reserve level $l$
+    \item $d_{nlt}^{L^+}$ - Possible imbalance on node $n$ for upward reserve level $l$
+    \item $d_{nlt}^{L^-}$ - Possible imbalance on node $n$ for downward reserve level $l$
 \end{itemize}
@@
 \begin{align}
-    \sum_{g \in G_n} r^{L^+}_{glt} + rs_{nlt} = rinj^{L^+}_{nlt}                & \quad n \in N, l \in L^+, \; t \in T    \\
-    \sum_{g \in G_n} r^{L^-}_{glt} + rc_{nlt} = rinj^{L^-}_{nlt}                & \quad n \in N, l \in L^-, \; t \in T    \\
-    rf^{L^+}_{blt} = \sum_{n \in N} PTDF_{nb} \cdot rinj^{L^+}_{nlt} & \quad b \in B, l \in L^+, \; t \in T    \\
-    rf^{L^-}_{blt} = \sum_{n \in N} PTDF_{nb} \cdot rinj^{L^-}_{nlt} & \quad b \in B, l \in L^-, \; t \in T    \\
-    -F_{b} \leq rf^{L^+}_{blt} \leq F_b                              & \quad b \in B, \; l \in L^+, \; t \in T \\
-    -F_{b} \leq rf^{L^-}_{blt} \leq F_b                              & \quad b \in B, \; l \in L^-, \; t \in T \\
+    \sum_{g \in G_n} r^{L^+}_{glt} + rs_{nlt} = d^{L^+}_{nlt} + rinj^{L^+}_{nlt} & \quad n \in N, \; l \in L^+, \; t \in T \\
+    \sum_{g \in G_n} r^{L^-}_{glt} + rc_{nlt} = d^{L^-}_{nlt} + rinj^{L^-}_{nlt} & \quad n \in N, \; l \in L^-, \; t \in T \\
+    rf^{L^+}_{blt} = \sum_{n \in N} PTDF_{nb} \cdot rinj^{L^+}_{nlt}             & \quad b \in B, \; l \in L^+, \; t \in T \\
+    rf^{L^-}_{blt} = \sum_{n \in N} PTDF_{nb} \cdot rinj^{L^-}_{nlt}             & \quad b \in B, \; l \in L^-, \; t \in T \\
+    -F_{b} \leq f_{bt} + rf^{L^+}_{blt} \leq F_b                                 & \quad b \in B, \; l \in L^+, \; t \in T \\
+    -F_{b} \leq f_{bt} + rf^{L^-}_{blt} \leq F_b                                 & \quad b \in B, \; l \in L^-, \; t \in T \\
     \sum_{n \in N} rinj^{L^+}_{nt} = 0                                           & \quad l \in L^+, \; n \in N, \; t \in T \\
@@
 Since imbalances are aggregated across the network, a particular reserve level activation is not associated with an imbalance at the nodal level.
-The above constraints therefore enforce that for each reserve level $l$ and node $n$, there exists some combination of node injections and generator dispatches which would satisfy the network constraints AND the imbalance across the entire network.
+The above constraints therefore enforce that for each reserve level $l$ and node $n$, there exists some combination of nodal imbalance, node injections, generator dispatches and line flows which would satisfy the network constraints AND the imbalance across the entire network.
 Given the formulation here, which uses reserve levels, i.e.
-quantiles, to represent forecast errors, it is difficult to come up with more stringent conditions.
+quantiles, over the entire network to represent forecast errors, it is difficult to come up with more stringent conditions.
 

### 2022-05-04 src/GEPPR.jl
Include store dispatch in security analsysis

@@ -294,6 +294,8 @@ function save_gep_for_security_analysis(gep::GEPM, path::String)
     q = gep[:q]
     z = gep[:z]
     e = gep[:e]
+    sc = gep[:sc]
+    sd = gep[:sd]
     ls = gep[:loadShedding]
     UC_results = Dict{Integer,Dict}()
     N, Y, P, T = GEPPR.get_set_of_nodes_and_time_indices(gep)
@@ -327,6 +329,8 @@ function save_gep_for_security_analysis(gep::GEPM, path::String)
                     "bus" => n,
                     "name" => st,
                     "e" => e[(st, n), y, p, atval(t - 1, typeof(e))],
+                    "discharge" => sd[(st, n), y, p, atval(t, typeof(sd))],
+                    "charge" => sc[(st, n), y, p, atval(t, typeof(sc))],
                 ) for (st, n) in STN
             ),
             "load_shed" => Dict(

### 2022-05-06 papers/main.tex
Errors in network redispatch formulation

@@
 \begin{align}
-    \sum_{g \in G_n} r^{L^+}_{glt} + rs_{nlt} = d^{L^+}_{nlt} + rinj^{L^+}_{nlt} & \quad n \in N, \; l \in L^+, \; t \in T \\
-    \sum_{g \in G_n} r^{L^-}_{glt} + rc_{nlt} = d^{L^-}_{nlt} + rinj^{L^-}_{nlt} & \quad n \in N, \; l \in L^-, \; t \in T \\
+    \sum_{g \in G_n, l'=1:l} (r^{L^+}_{glt} + rs_{nlt}) = d^{L^+}_{nlt} + rinj^{L^+}_{nlt} & \quad n \in N, \; l \in L^+, \; t \in T \\
+    - \sum_{g \in G_n, l'=1:l} (r^{L^-}_{glt} + rc_{nlt}) = d^{L^-}_{nlt} + rinj^{L^-}_{nlt} & \quad n \in N, \; l \in L^-, \; t \in T \\
+    \sum_{n \in N} d^{L^+}_{nlt} = \sum_{l'=1:l} D^{L^+}_{l't}                             & \quad l \in L^+, \; t \in T             \\
+    \sum_{n \in N} d^{L^-}_{nlt} = -\sum_{l'=1:l} D^{L^-}_{l't}                            & \quad l \in L^-, \; t \in T             \\
     rf^{L^+}_{blt} = \sum_{n \in N} PTDF_{nb} \cdot rinj^{L^+}_{nlt}                       & \quad b \in B, \; l \in L^+, \; t \in T \\

### 2022-05-06 scripts/DUCPR_simple.jl

@@ -1,0 +1,80 @@
+include(joinpath(@__DIR__, "..", "intro.jl"))
+sn = script_name(@__FILE__)
+mkrootdirs(plotsdir(sn))
+
+opts_vec = options_diff_days(sn)
+init_soc = 0.0
+opts_vec = vcat(
+    [
+        merge(
+            opts,
+            Dict(
+                "time_out" => 600,
+                "save_path" => joinpath(
+                    opts["save_path"], "RSL=$(v)_init_SOC=$(init_soc)"
+                ),
+                "reserve_shedding_limit" => v,
+                "initial_state_of_charge" => init_soc,
+            ),
+        ) for opts in opts_vec, v in [1.0, 0.0]
+    ]...,
+)
+gep_vec = run_GEPPR(opts_vec)
+GC.gc() # Who knows, maybe this will help
+map(
+    i -> if gep_vec[i] === nothing
+        nothing
+    else
+        save_gep_for_security_analysis(gep_vec[i], opts_vec[i])
+    end,
+    eachindex(opts_vec),
+)
+
+nd = length(options_diff_days(sn))
+sid_vec = [month_day(opts_vec[i]) for i in 1:length(opts_vec)]
+rsl_vec = [opts_vec[i]["reserve_shedding_limit"] for i in 1:length(opts_vec)]
+rsl_mat = reshape(rsl_vec, nd, 2)
+ls = Dict(
+    (sid_vec[i], rsl_vec[i]) => try
+        sum(gep_vec[i][:loadShedding].data; dims=(1, 2, 3))[:]
+    catch
+        GEPPR.SVC(NaN)
+    end for i in 1:length(opts_vec)
+)
+rsLt = Dict(
+    (sid_vec[i], rsl_vec[i]) => try
+        gep_vec[i][:rsL⁺]
+    catch
+        GEPPR.SVC(NaN)
+    end for i in 1:length(opts_vec)
+)
+rsL⁺ = Dict(
+    (sid_vec[i], rsl_vec[i]) => try
+        sum(
+            rsLt[(sid_vec[i], rsl_vec[i])][n, l, 1, 1, t]
+            # rsLt[(sid_vec[i], rsl_vec[i])][n, l, 1, 1, t] *
+            # P⁺[(sid_vec[i], rsl_vec[i])][l, 1, 1, t] 
+            for n in GEPPR.get_set_of_nodes(gep_vec[i]),
[…]

### 2022-05-06 scripts/DUCPR_reserve_shedding_sensitivity.jl
Errors in network redispatch formulation

@@ -1,5 +1,6 @@
 include(joinpath(@__DIR__, "..", "intro.jl"))
 sn = script_name(@__FILE__)
+mkrootdirs(plotsdir(sn))
 
 opts_vec = options_diff_days(sn)
 opts_vec = vcat(
@@ -12,53 +13,58 @@ opts_vec = vcat(
                 "time_out" => 600,
                 "upward_reserve_levels_included_in_redispatch" => L,
                 "downward_reserve_levels_included_in_redispatch" => L,
-                "vars_2_save" => [:z, :q, :ls, :rsL⁺, :rsL⁻, :e, :sc, :sd],
-                "exprs_2_save" => [:loadShedding]
             ),
-        ) for opts in opts_vec, L in [Int[],1:10], v in 0.1:-0.02:0 
+        ) for opts in opts_vec, L in [Int[], 1:10], v in 0.1:-0.02:0
         # ) for v in [0.02], opts in opts_vec
     ]...,
 )
-gep = run_GEPPR(opts_vec[1])
+# gep = run_GEPPR(opts_vec[1])
 # gep_vec = run_GEPPR(opts_vec)
 gep_vec = run_GEPPR(opts_vec; load_only=true)
 GC.gc() # Who knows, maybe this will help
 map(
-    i ->
-        if gep_vec[i] === nothing
-            nothing
-        else
-            save_gep_for_security_analysis(gep_vec[i], opts_vec[i])
-        end,
+    i -> if gep_vec[i] === nothing
+        nothing
+    else
+        save_gep_for_security_analysis(gep_vec[i], opts_vec[i])
+    end,
     eachindex(opts_vec),
 )
 
 sid_vec = [month_day(opts_vec[i]) for i in 1:length(opts_vec)]
 rsl_vec = [opts_vec[i]["reserve_shedding_limit"] for i in 1:length(opts_vec)]
-L_vec = [length(opts["upward_reserve_levels_included_in_redispatch"]) for opts in opts_vec]
+L_vec = [
+    length(opts["upward_reserve_levels_included_in_redispatch"]) for
+    opts in opts_vec
+]
 
 # Plot the minimum number of units committed each day
 u_min = DataFrame(
     "u_min" => [
         try
[…]

### 2022-05-06 src/GEPPR.jl

@@ -37,6 +37,7 @@ function param_and_config(opts::Dict)
         ),
         "imposePeriodicityConstraintOnStorage" =>
             rolling_horizon ? false : true,
+        "storageTechnologies" => Dict()
     )
     if length(optimization_horizon[1]:optimization_horizon[end]) > 100 &&
         is_linear == false &&
@@ -47,6 +48,17 @@ function param_and_config(opts::Dict)
     end
     if opts["copperplate"] == true
         param["forceCopperPlateModel"] = true
+    end
+    if ismissing(opts["initial_state_of_charge"]) == false
+        init_soc = opts["initial_state_of_charge"]
+        @assert 0 <= init_soc <= 1
+        storage_data = YAML.load(open(datadir("pro", "GEPPR", "storage.yaml")))["storageTechnologies"]
+        param["storageTechnologies"] = Dict(
+            st => Dict(
+                "initialStateOfCharge" =>
+                    v["installedEnergyCapacity"] * init_soc,
+            ) for (st,v) in storage_data
+        )
     end
     return configFiles, param
 end
@@ -116,13 +128,15 @@ function run_GEPPR(opts::Dict; load_only=false)
     return gep
 end
 
-run_GEPPR(opts_vec; kwargs...) = [
-    try
-        run_GEPPR(opts; kwargs...)
-    catch
-        @warn "Optimisation failed"
-    end for opts in opts_vec
-]
+function run_GEPPR(opts_vec; kwargs...)
+    return [
+        try
+            run_GEPPR(opts; kwargs...)
+        catch
+            @warn "Optimisation failed"
+        end for opts in opts_vec
+    ]
+end
 
 function GEPPR.load_GEP(opts::Dict, path::String)
     @load eval(joinpath(path, "config.jld2")) dictConfig
@@ -249,12 +263,13 @@ function prevent_simultaneous_charge_and_discharge!(gep::GEPM, opts::Dict)
     STN = GEPPR.get_set_of_nodal_storage_technologies(gep)
     SCK = GEPPR.get_storage_charging_capacity_expression(gep)
     SDK = GEPPR.get_storage_discharging_capacity_expression(gep)
-    gep[:M, :variables, :o] = o = @variable(
-        gep.model,
-        [stn = STN, y = Y, p = P, t = T],
-        binary = true,
-        base_name = "o"
-    )
+    gep[:M, :variables, :o] =
+        o = @variable(
+            gep.model,
[…]

### 2022-05-06 src/opts.jl

@@ -22,6 +22,7 @@ function options(pairs...)
         "copperplate" => false,
         "vars_2_save" => Symbol[],
         "exprs_2_save" => Symbol[],
+        "initial_state_of_charge" => missing
     )
     for (k, v) in pairs
         opts[k] = v
@@ -38,7 +39,8 @@ function options_diff_days(sn)
         # "initial_commitment_data_path" => datadir("sims", "rolling_UC_full_year"),
         "operating_reserves_type" => "probabilistic",
         "operating_reserves_sizing_type" => "given",
-        "vars_2_save" => [:z, :q],
+        "vars_2_save" => [:z, :q, :ls, :rL⁺, :rL⁻, :rsL⁺, :rsL⁻, :e, :sc, :sd],
+        "exprs_2_save" => [:loadShedding],
         "time_out" => 60*10
     )
     opts_vec = [

### 2022-05-06 src/GEPPR.jl
Fix initial SOC for storage

@@ -287,6 +287,17 @@ function prevent_simultaneous_charge_and_discharge!(gep::GEPM, opts::Dict)
     return gep
 end
 
+function apply_initial_state_of_charge!(gep::GEPM, opts::Dict)
+    E_init = GEPPR.get_storage_initial_state_of_charge(gep, default=missing)
+    e = GEPPR.get_storage_state_of_charge_var(gep)
+    STN = GEPPR.get_set_of_nodal_storage_technologies(gep)
+    N, Y, P, T = GEPPR.get_set_of_nodes_and_time_indices(gep)
+    for stn in STN
+        fix(e[stn,Y[1],P[1],T[1]-1], E_init[stn], force=true)
+    end
+    return gep
+end
+
 function save(gep::GEPM, opts::Dict)
     @unpack save_path = opts
     vars_2_save = get(opts, "vars_2_save", nothing)

### 2022-05-07 scripts/DUCPR_simple.jl
Fixi initial SOC issue

@@ -76,5 +76,7 @@ Plots.plot(
     xlab="Reserve shedding limit [0 - 1]",
     ylab="Day ahead load shedding [MWh]",
 )
+Plots.savefig(plotsdir(sn, "load_shedding_vs_RSL.png"))
 
 plot_dispatch(gep_vec[1], 1)
+Plots.savefig(plotsdir(sn, "dispatch_$(sid_vec[1]).png"))

### 2022-05-07 src/GEPPR.jl

@@ -114,6 +114,7 @@ function run_GEPPR(opts::Dict; load_only=false)
             apply_initial_commitment!(gep, opts)
             constrain_reserve_shedding!(gep, opts)
             prevent_simultaneous_charge_and_discharge!(gep, opts)
+            apply_initial_state_of_charge!(gep, opts)
             optimize_GEP_model!(gep)
             save_optimisation_values!(gep)
         else

### 2022-05-09 scripts/DUCPR_RSL_debug.jl
Add reserve provision costs

@@ -3,29 +3,40 @@ sn = script_name(@__FILE__)
 mkrootdirs(plotsdir(sn))
 
 opts = options_diff_days(sn)[4]
+rm(opts["save_path"]; force=true, recursive=true)
+opts["unit_commitment_type"] = "none"
 opts["initial_state_of_charge"] = 0.0
-opts_vec = [
-    merge(
-        opts,
-        Dict(
-            "unit_commitment_type" => "none",
-            "include_storage" => false,
-            "save_path" => joinpath(opts["save_path"], "simplest"),
-        ),
-    )
-    merge(
-        opts,
-        Dict(
-            "unit_commitment_type" => "none",
-            "save_path" => joinpath(opts["save_path"], "simpler"),
-        ),
-    )
-    merge(
-        opts,
-        Dict(
-            "unit_commitment_type" => "none",
-            "save_path" => joinpath(opts["save_path"], "full"),
-        ),
-    )
-]
-gep_vec = run_GEPPR(opts_vec[1:2])
+opts["reserve_provision_cost"] = 200.0
+gep = run_GEPPR(opts)
+
+@show opts["reserve_provision_cost"]
+@show sum(gep[:rsL⁺])
+@show sum(gep[:loadShedding])
+nothing
+
+
+# opts_vec = [
+#     merge(
+#         opts,
+#         Dict(
+#             "unit_commitment_type" => "none",
+#             "include_storage" => false,
+#             "save_path" => joinpath(opts["save_path"], "simplest"),
+#         ),
+#     )
+#     merge(
+#         opts,
+#         Dict(
+#             "unit_commitment_type" => "none",
+#             "save_path" => joinpath(opts["save_path"], "simpler"),
+#         ),
+#     )
+#     merge(
+#         opts,
+#         Dict(
+#             "unit_commitment_type" => "none",
+#             "save_path" => joinpath(opts["save_path"], "full"),
+#         ),
+#     )
+# ]
+# gep_vec = run_GEPPR(opts_vec[1:2])

### 2022-05-09 src/GEPPR.jl

@@ -3,9 +3,17 @@ include(srcdir("util.jl"))
 
 ### UTIL
 function param_and_config(opts::Dict)
-    @unpack optimization_horizon, rolling_horizon, include_downward_reserves =
-        opts
+    @unpack optimization_horizon,
+    rolling_horizon,
+    include_downward_reserves,
+    include_storage,
+    copperplate,
+    initial_state_of_charge,
+    reserve_provision_cost = opts
+
     is_linear = (opts["unit_commitment_type"] == "none")
+    init_soc = opts["initial_state_of_charge"]
+
     GEPPR_dir = datadir("pro", "GEPPR")
     configFiles =
         joinpath.(
@@ -18,9 +26,8 @@ function param_and_config(opts::Dict)
                 "RES.yaml",
             ],
         )
-    if opts["include_storage"]
-        push!(configFiles, joinpath(GEPPR_dir, "storage.yaml"))
-    end
+
+    # Param
     param = @suppress Dict{String,Any}(
         "optimizer" => optimizer(opts),
         "unitCommitmentConstraintType" => opts["unit_commitment_type"],
@@ -38,8 +45,34 @@ function param_and_config(opts::Dict)
         ),
         "imposePeriodicityConstraintOnStorage" =>
             rolling_horizon ? false : true,
-        "storageTechnologies" => Dict(),
+        "dispatchableGeneration" => Dict(),
+        "intermittentGeneration" => Dict(),
     )
+
+    # In case including storage
+    if include_storage
+        push!(configFiles, joinpath(GEPPR_dir, "storage.yaml"))
+        opts["storageTechnologies"] = Dict()
+    end
+
+    # Reserve provision costs
+    if reserve_provision_cost > 0.0
+        rpc = reserve_provision_cost
+        disp_data = YAML.load(open(datadir("pro", "GEPPR", "units.yaml")))["dispatchableGeneration"]
+        res_data = YAML.load(open(datadir("pro", "GEPPR", "RES.yaml")))["intermittentGeneration"]
+        for (k, v) in disp_data
+            param["dispatchableGeneration"][k] = Dict(
+                "reserveProvisionCost" => rpc
+            )
+        end
+        for (k, v) in res_data
+            param["intermittentGeneration"][k] = Dict(
+                "reserveProvisionCost" => rpc
+            )
+        end
+    end
+
+    # Error if this optimization horizon too long
     if length(optimization_horizon[1]:optimization_horizon[end]) > 100 &&
         is_linear == false &&
[…]

### 2022-05-09 src/opts.jl

@@ -23,7 +23,8 @@ function options(pairs...)
         "vars_2_save" => Symbol[],
         "exprs_2_save" => Symbol[],
         "initial_state_of_charge" => missing,
-        "replace_storage_dispatch_with_node_injection" => false
+        "replace_storage_dispatch_with_node_injection" => false,
+        "reserve_provision_cost" => 0.0
     )
     for (k, v) in pairs
         opts[k] = v

### 2022-05-10 src/process_input.jl
Remove power to gas

@@ -198,33 +198,33 @@ function process_belderbos_data(grid_path)
         )
     end
 
-    # Power to Gas
-    h2_sh = gen["hydrogen_storage"]
-    duration = 1000
-    for row in XLSX.eachrow(h2_sh)
-        row["A"] in ("ID", "(integer)") && continue
-        id += 1
-        network["storage"][string(id)] = Dict(
-            "index" => id,
-            "storage_bus" => row["C"],
-            "ps" => 0.0,
-            "qs" => 0.0,
-            "energy" => 0.0,
-            "energy_rating" => row["D"],
-            "charge_rating" => row["D"] / duration,
-            "discharge_rating" => row["D"] / duration,
-            "charge_efficiency" => 0.7,
-            "discharge_efficiency" => 0.7,
-            "qmin" => 0.0,
-            "qmax" => 0.0,
-            "r" => 0.0,
-            "x" => 0.0,
-            "p_loss" => 0.0,
-            "q_loss" => 0.0,
-            "status" => 1,
-            "name" => "P2G_$(id)",
-        )
-    end
+    # # Power to Gas
+    # h2_sh = gen["hydrogen_storage"]
+    # duration = 1000
+    # for row in XLSX.eachrow(h2_sh)
+    #     row["A"] in ("ID", "(integer)") && continue
+    #     id += 1
+    #     network["storage"][string(id)] = Dict(
+    #         "index" => id,
+    #         "storage_bus" => row["C"],
+    #         "ps" => 0.0,
+    #         "qs" => 0.0,
+    #         "energy" => 0.0,
+    #         "energy_rating" => row["D"],
+    #         "charge_rating" => row["D"] / duration,
+    #         "discharge_rating" => row["D"] / duration,
+    #         "charge_efficiency" => 0.7,
+    #         "discharge_efficiency" => 0.7,
+    #         "qmin" => 0.0,
+    #         "qmax" => 0.0,
+    #         "r" => 0.0,
+    #         "x" => 0.0,
+    #         "p_loss" => 0.0,
+    #         "q_loss" => 0.0,
+    #         "status" => 1,
+    #         "name" => "P2G_$(id)",
+    #     )
+    # end
 
     # Save the dictionary
     exp_pro = datadir("pro")

### 2022-05-10 src/GEPPR.jl
Fix mixup with timeseries without storage

@@ -33,9 +33,9 @@ function param_and_config(opts::Dict)
         "unitCommitmentConstraintType" => opts["unit_commitment_type"],
         "relativePathTimeSeriesCSV" =>
             if opts["replace_storage_dispatch_with_node_injection"]
-                "timeseries.csv"
-            else
                 "timeseries_wo_storage.csv"
+            else
+                "timeseries.csv"
             end,
         "includeDownwardReserves" => include_downward_reserves,
         "relativePathMatpowerData" => basename(grid_red_path),

### 2022-05-10 src/process_input.jl
Fix mixup with timeseries without storage

@@ -339,6 +339,7 @@ function powermodels_2_GEPPR(grid_data_path, grid_red_path)
                 ]...,
             ) for t in res_names
         ]...,
+        "Weight" => [1.0 for i in 1:nT]
     )
     name = if isempty(network["storage"])
         "timeseries_wo_storage.csv"

### 2022-05-10 papers/main.tex
Saving

@@
     \sum_{n \in N} d^{L^+}_{nlt} = \sum_{l'=1:l} D^{L^+}_{l't}                             & \quad l \in L^+, \; t \in T             \\
-    \sum_{n \in N} d^{L^-}_{nlt} = -\sum_{l'=1:l} D^{L^-}_{l't}                            & \quad l \in L^-, \; t \in T             \\
+    \sum_{n \in N} d^{L^-}_{nlt} = - \sum_{l'=1:l} D^{L^-}_{l't}                            & \quad l \in L^-, \; t \in T             \\
     rf^{L^+}_{blt} = \sum_{n \in N} PTDF_{nb} \cdot rinj^{L^+}_{nlt}                       & \quad b \in B, \; l \in L^+, \; t \in T \\

### 2022-05-10 scripts/DUCPR_RSL_debug.jl
Saving

@@ -6,10 +6,11 @@ opts = options_diff_days(sn)[4]
 rm(opts["save_path"]; force=true, recursive=true)
 opts["unit_commitment_type"] = "none"
 opts["initial_state_of_charge"] = 0.0
-opts["reserve_provision_cost"] = 200.0
+opts["upward_reserve_levels_included_in_redispatch"] = 1:10
+opts["downward_reserve_levels_included_in_redispatch"] = 1:10
 gep = run_GEPPR(opts)
 
-@show opts["reserve_provision_cost"]
+@show opts["reserve_provision_coupwardst"]
 @show sum(gep[:rsL⁺])
 @show sum(gep[:loadShedding])
 nothing

### 2022-05-10 scripts/interaction_check.jl
Keep including hydrogen storage in model

@@ -5,5 +5,6 @@ mkrootdirs(simsdir(sn))
 
 opts = options_diff_days(sn)[1]
 opts["initial_state_of_charge"] = 0.0
+opts["time_out"] = 120
 gep = run_GEPPR(opts)
 d = save_gep_for_security_analysis(gep, opts)

### 2022-05-10 src/process_input.jl

@@ -198,33 +198,33 @@ function process_belderbos_data(grid_path)
         )
     end
 
-    # # Power to Gas
-    # h2_sh = gen["hydrogen_storage"]
-    # duration = 1000
-    # for row in XLSX.eachrow(h2_sh)
-    #     row["A"] in ("ID", "(integer)") && continue
-    #     id += 1
-    #     network["storage"][string(id)] = Dict(
-    #         "index" => id,
-    #         "storage_bus" => row["C"],
-    #         "ps" => 0.0,
-    #         "qs" => 0.0,
-    #         "energy" => 0.0,
-    #         "energy_rating" => row["D"],
-    #         "charge_rating" => row["D"] / duration,
-    #         "discharge_rating" => row["D"] / duration,
-    #         "charge_efficiency" => 0.7,
-    #         "discharge_efficiency" => 0.7,
-    #         "qmin" => 0.0,
-    #         "qmax" => 0.0,
-    #         "r" => 0.0,
-    #         "x" => 0.0,
-    #         "p_loss" => 0.0,
-    #         "q_loss" => 0.0,
-    #         "status" => 1,
-    #         "name" => "P2G_$(id)",
-    #     )
-    # end
+    # Power to Gas
+    h2_sh = gen["hydrogen_storage"]
+    duration = 1000
+    for row in XLSX.eachrow(h2_sh)
+        row["A"] in ("ID", "(integer)") && continue
+        id += 1
+        network["storage"][string(id)] = Dict(
+            "index" => id,
+            "storage_bus" => row["C"],
+            "ps" => 0.0,
+            "qs" => 0.0,
+            "energy" => 0.0,
+            "energy_rating" => row["D"],
+            "charge_rating" => row["D"] / duration,
+            "discharge_rating" => row["D"] / duration,
+            "charge_efficiency" => 0.7,
+            "discharge_efficiency" => 0.7,
+            "qmin" => 0.0,
+            "qmax" => 0.0,
+            "r" => 0.0,
+            "x" => 0.0,
+            "p_loss" => 0.0,
+            "q_loss" => 0.0,
+            "status" => 1,
+            "name" => "P2G_$(id)",
+        )
+    end
 
     # Save the dictionary
     exp_pro = datadir("pro")
@@ -315,7 +315,7 @@ function powermodels_2_GEPPR(grid_data_path, grid_red_path)
                             v["charge_rating"], v["discharge_rating"]
                         ]),
[…]

### 2022-05-10 scripts/interaction_check.jl
Continuing OPF debugging

@@ -5,6 +5,11 @@ mkrootdirs(simsdir(sn))
 
 opts = options_diff_days(sn)[1]
 opts["initial_state_of_charge"] = 0.0
+# opts["operating_reserves_type"] = "none"
 opts["time_out"] = 120
 gep = run_GEPPR(opts)
 d = save_gep_for_security_analysis(gep, opts)
+
+df = overall_network_results(gep)
+df[:,"Load - net generation - load shed"]
+CSV.write(joinpath(opts["save_path"], "aggregate_results.csv"), df)

### 2022-05-10 src/GEPPR.jl

@@ -154,6 +154,8 @@ function run_GEPPR(opts::Dict; load_only=false)
             constrain_reserve_shedding!(gep, opts)
             prevent_simultaneous_charge_and_discharge!(gep, opts)
             apply_initial_state_of_charge!(gep, opts)
+            # fix_model_variable!(gep, :sc, GEPPR.SVC(0.0))
+            # fix_model_variable!(gep, :sd, GEPPR.SVC(0.0))
             optimize_GEP_model!(gep)
             save_optimisation_values!(gep)
         else

### 2022-05-11 scripts/interaction_check.jl
Fixed infeasible OPF

@@ -13,3 +13,55 @@ d = save_gep_for_security_analysis(gep, opts)
 df = overall_network_results(gep)
 df[:,"Load - net generation - load shed"]
 CSV.write(joinpath(opts["save_path"], "aggregate_results.csv"), df)
+
+# pfl is from Efthymios' power flow which I'm not pushing to github
+function line_flow_comparison(gep, pfl, t)
+    f_1 = gep[:f][:,:,:,t]
+    # Im = GEPPR.get_incidence_matrix(gep)
+    Im = calc_basic_incidence_matrix(gep.networkData)
+    N, Y, P, Y = GEPPR.get_set_of_nodes_and_time_indices(gep)
+    L = GEPPR.get_set_of_lines(gep)
+    L2Lidx = GEPPR.get_line_indices(gep)
+    node_index_2_node_name = Dict(
+        v["index"] => v["string"] for (k,v) in gep.networkData["bus"]
+    )
+
+    my_flow = Float64[]
+    ef_flow = Float64[]
+    lidx_vec = Int64[]
+    t_br = Int64[]
+    f_br = Int64[]
+    t_br_str = String[]
+    f_br_str = String[]
+
+    for l in L
+        lidx = L2Lidx[l]
+        br_vec = abs.(Im[lidx,:])
+        n1, n2 = findall(br_vec .== 1)
+        idx_opf = (lidx, n1, n2)
+        push!(ef_flow, pfl[idx_opf])
+        push!(my_flow, f_1[l,1,1]/100)
+        push!(lidx_vec, lidx)
+        push!(t_br, n1)
+        push!(t_br_str, node_index_2_node_name[n1])
+        push!(f_br, n2)
+        push!(f_br_str, node_index_2_node_name[n2])
+    end
+
+    df = DataFrame(
+        "Line_name" => L,
+        "Line_index" => lidx_vec,
+        "From_branch_name" => f_br_str,        
+        "From_branch_idx" => f_br,
+        "To_branch_name" => t_br_str,        
+        "To_branch_idx" => t_br,
+        "UC_flow" => my_flow,
+        "OPF_flow" => ef_flow
+    )
+
+    CSV.write(joinpath(opts["save_path"], "line_flows_hour=$(t).csv"), df)
+
+    return df
+end
+
+df = line_flow_comparison(gep, pfl, 5113)

### 2022-05-18 src/GEPPR.jl
Fix bug in tech_node_2_idx

@@ -537,7 +537,7 @@ function save_gep_for_security_analysis(gep::GEPM, path::String)
 end
 
 function tech_node_to_idx(gep::GEPM)
-    d = gep.networkData
+    nd = gep.networkData
     GDN = GEPPR.get_set_of_nodal_dispatchable_generators(gep)
     GRN = GEPPR.get_set_of_nodal_intermittent_generators(gep)
     STN = GEPPR.get_set_of_nodal_storage_technologies(gep)
@@ -550,7 +550,7 @@ function tech_node_to_idx(gep::GEPM)
                         bus_idx_2_name[string(pair[2]["bus"])] ==
                         string(gn[2]) && pair[2]["name"] == gn[1]
                     ),
-                    collect(n["res"]),
+                    collect(nd["res"]),
                 )][1],
             ) for gn in GRN
         ),

### 2022-05-18 scripts/main_model_runs.jl
Update main model runs

@@ -1,7 +1,6 @@
 include(joinpath(@__DIR__, "..", "intro.jl"))
 sn = script_name(@__FILE__)
 mkrootdirs(plotsdir(sn))
-rm(simsdir(sn); force=true, recursive=true)
 
 opts = options_diff_days(sn)[4]
 opts["initial_state_of_charge"] = 0.0
@@ -38,11 +37,55 @@ opts_vec = [
         opts,
         Dict(
             "operating_reserves_type" => "none",
-            "save_path" =>
-                joinpath(opts["save_path"], "base_UC=true_DANet=true_PSCD=true"),
+            "save_path" => joinpath(
+                opts["save_path"], "base_UC=true_DANet=true_PSCD=true"
+            ),
         ),
     ),
 ]
+opts_vec = vcat(
+    opts_vec...,
+    [
+        merge(
+            opts,
+            Dict(
+                "operating_reserves_type" => "probabilistic",
+                "reserve_shedding_limit" => v,
+                "save_path" => "$(opts["save_path"])_RSV=$(v)",
+            ),
+        ) for opts in opts_vec, v in [0.0, 0.5, 1.0]
+    ]...,
+)
+opts_vec = vcat(
+    opts_vec...,
+    [
+        merge(
+            opts,
+            Dict(
+                "upward_reserve_levels_included_in_redispatch" => 1:10,
+                "downward_reserve_levels_included_in_redispatch" => 1:10,
+                "save_path" => "$(opts["save_path"])_L⁺=1:10_L⁻=1:10",
+            ),
+        ) for opts in opts_vec if opts["reserve_shedding_limit"] < 1.0
+    ]...,
+)
+opts_vec = vcat(
+    opts_vec...,
+    [
+        merge(
+            opts,
+            Dict(
+                "absolute_limit_on_nodal_imbalance" => true,
+                "save_path" => "$(opts["save_path"])_AbsIm=true",
+            ),
+        ) for opts in opts_vec if opts["reserve_shedding_limit"] < 1.0 && (
+            isempty(opts["upward_reserve_levels_included_in_redispatch"]) ==
+            false ||
+            isempty(opts["downward_reserve_levels_included_in_redispatch"]) ==
+            false
+        )
+    ],
+)
 
 function main_model_run(opts; make_plots=true)
     gep = run_GEPPR(opts)
@@ -54,6 +97,7 @@ function main_model_run(opts; make_plots=true)
 
[…]

### 2022-05-18 papers/main.tex
Updating scripts

@@
 
-\title{Description of a Deterministic Unit Commitment Model with Probabilistic Reserve constraints and Network Representation}
+\title{Resource Adequacy and Operational Security Interaction in the EPOC 2030-50 Project}
 
@@
 
-\section{Model description}
+\section{Description of a Deterministic Unit Commitment Model with Probabilistic Reserve (DUC-PR) constraints}
 
@@
 
+\section{Input data - stylised Belgian grid with a large amount of renewables}
+
+\section{Analysis of DUC-PR model results}
+
+This analysis relates only to the 309th day of the year (November 5th).
+
+\subsection{No operating reserves}
+
+\begin{table}
+    \begin{tabular}{ccccc}
+        \toprule
+        UC  & DANet & PSCD & Load shedding [MWh] & Objective \\
+            &       &      & 0.0                 & 101,916 \\
+        \xm &       &      & 0.0                 & 212,558 \\
+        \xm & \xm   &      & 0.0                 & 252,327 \\
+        \xm & \xm   & \xm  & 0.0                 & 252,327 \\
+    \end{tabular}
+    \caption{Analysis of model results when operating reserves are not included.
+UC = Unit Commitment constraints, DANet = Day Ahead Network constraints, PSCD = Prevent Simulataneous Charging and Discharging constraints.}
+\end{table}
+
 \end{document}

### 2022-05-18 src/GEPPR.jl

@@ -461,6 +461,8 @@ function save(gep::GEPM, opts::Dict)
         JSON.print(f, opts, 4)
     end
 
+    save_gep_for_security_analysis(gep, opts)
+
     vars_2_save = get(opts, "vars_2_save", nothing)
     exprs_2_save = get(opts, "exprs_2_save", nothing)
     if vars_2_save !== nothing
@@ -482,10 +484,10 @@ Saves data in the format: hour -> generator (with associated bus) -> values / na
 """
 function save_gep_for_security_analysis(gep::GEPM, path::String)
     q = gep[:q]
-    z = gep[:z]
-    e = gep[:e]
-    sc = gep[:sc]
-    sd = gep[:sd]
+    z = gep[:z, SVC(missing)]
+    e = gep[:e, SVC(missing)]
+    sc = gep[:sc, SVC(missing)]
+    sd = gep[:sd, SVC(missing)]
     ls = gep[:loadShedding]
     UC_results = Dict{Integer,Dict}()
     N, Y, P, T = GEPPR.get_set_of_nodes_and_time_indices(gep)

### 2022-05-18 scripts/DUCPR_RSL_debug.jl
Updating scripts

@@ -15,9 +15,9 @@ opts["convex_hull_limit_on_nodal_imbalance"] = true
 opts["n_scenarios_for_convex_hull_calc"] => 1_000
 # opts["prevent_simultaneous_charge_and_discharge"] = false
 # opts["upward_reserve_levels_included_in_redispatch"] = [9,10]
-opts["upward_reserve_levels_included_in_redispatch"] = 1:10
+# opts["upward_reserve_levels_included_in_redispatch"] = 1:10
 # opts["downward_reserve_levels_included_in_redispatch"] = [9,10]
-opts["downward_reserve_levels_included_in_redispatch"] = 1:10
+# opts["downward_reserve_levels_included_in_redispatch"] = 1:10
 
 RSL_vec = [1.0, 0.5, 0.0]
 opts_vec = [
@@ -41,3 +41,14 @@ for i in 1:length(opts_vec)
     display(plot_dispatch(gep_vec[i], 1))
     display(plot_reserves_simple(gep_vec[i], 1))
 end
+
+opts = options_diff_days(sn)[4]
+rm(opts["save_path"]; force=true, recursive=true)
+opts["initial_state_of_charge"] = 0.0
+opts["convex_hull_limit_on_nodal_imbalance"] = true
+opts["n_scenarios_for_convex_hull_calc"] => 1_000
+opts["include_downward_reserves"] = false
+opts["upward_reserve_levels_included_in_redispatch"] = 1:10
+opts["downward_reserve_levels_included_in_redispatch"] = 1:10
+
+gep = run_GEPPR(opts)

### 2022-05-18 scripts/main_model_runs.jl

@@ -2,6 +2,15 @@ include(joinpath(@__DIR__, "..", "intro.jl"))
 sn = script_name(@__FILE__)
 mkrootdirs(plotsdir(sn))
 
+function res_act_net(opts)
+    return opts["reserve_shedding_limit"] < 1.0 && (
+        isempty(opts["upward_reserve_levels_included_in_redispatch"]) ==
+        false ||
+        isempty(opts["downward_reserve_levels_included_in_redispatch"]) ==
+        false
+    )
+end
+
 opts = options_diff_days(sn)[4]
 opts["initial_state_of_charge"] = 0.0
 opts_vec = [
@@ -78,12 +87,7 @@ opts_vec = vcat(
                 "absolute_limit_on_nodal_imbalance" => true,
                 "save_path" => "$(opts["save_path"])_AbsIm=true",
             ),
-        ) for opts in opts_vec if opts["reserve_shedding_limit"] < 1.0 && (
-            isempty(opts["upward_reserve_levels_included_in_redispatch"]) ==
-            false ||
-            isempty(opts["downward_reserve_levels_included_in_redispatch"]) ==
-            false
-        )
+        ) for opts in opts_vec if res_act_net(opts)
     ],
 )
 
@@ -132,7 +136,7 @@ for opts in opts_vec
         @warn "$(opts["save_path"]) failed."
         println(string(e))
     end
-end 
+end
 
 function analyse_main_model_runs(gep_vec, opts_vec)
     df = DataFrame(
@@ -141,6 +145,14 @@ function analyse_main_model_runs(gep_vec, opts_vec)
         "DANet" => [opts["copperplate"] == false for opts in opts_vec],
         "PSCD" => [
             opts["prevent_simultaneous_charge_and_discharge"] == true for
+            opts in opts_vec
+        ],
+        "OR" =>
+            [opts["operating_reserves_type"] != "none" for opts in opts_vec],
+        "RSV" => [opts["reserve_shedding_limit"] for opts in opts_vec],
+        "RANet" => [res_act_net(opts) for opts in opts_vec],
+        "AbsImb" => [
+            opts["absolute_limit_on_nodal_imbalance"] == true for
             opts in opts_vec
         ],
         "Reserve Shedding" =>

### 2022-05-23 papers/main.tex
Continue writing up results

@@
 
-\section{Input data - stylised Belgian grid with a large amount of renewables}
+\section{Input data - stylised Belgian grid with a large amount of variable renewable energy sources}
 
@@
 \subsection{No operating reserves}
+
+\tabref{tab:results_no_OR} lists results for an increasing level of technical constraints or increasingly inflexible system.
+While the objective more than doubles going from the simple linear economic dispatch model with no network constraints to the unit commitment model with network constraints, no load shedding occurs.
+Preventing simultaneous charging and discharging constraints does not affect results at all, implying that additional energy consumption from storage does not aid congestion.
 
@@
     \caption{Analysis of model results when operating reserves are not included.
-UC = Unit Commitment constraints, DANet = Day Ahead Network constraints, PSCD = Prevent Simulataneous Charging and Discharging constraints.}
+UC = Unit Commitment constraints, DANet = Day Ahead Network constraints, PSCD = Prevent Simultaneous Charging and Discharging constraints.}\label{tab:results_no_OR}
+\end{table}
+
+\subsection{Probabilistic operating reserves with no reserve activation network constraints}
+
+\begin{itemize}
+    \item Unit commitment constraints lead to unavoidable load shedding. 
+    \item Results should be interpreted with caution however, since reserve provision is quite strict (storage can't provide reserves, non spinning reserves not modelled).
+    \item Network constraints are less problematic than unit commitment (cost increase of ~20,000 euros instead of 500,000) (perhaps thanks to storage, who knows).
+    \item Preventing simultaneous charge and discharge has a small but noticeable effect.
+\end{itemize}
+
+\begin{table}
+    \begin{tabular}{ccccccc}
+        \toprule
+        UC  & DANet & PSCD & RSL & Load shedding [MWh] & Reserve Shedding [MWh] & Objective \\
+            &       &      & 0.0 & 0.0                 & 0.0                    & 117,290   \\
+            &       &      & 0.5 & 0.0                 & 0.0                    & 117,290   \\
+            &       &      & 1.0 & 0.0                 & 0.0                    & 117,290   \\
+        \midrule
+        \xm &       &      & 0.0 & 1,669               & 0.0                    & 645,679   \\
+        \xm &       &      & 0.5 & 1,669               & 0.0                    & 645,679   \\
+        \xm &       &      & 1.0 & 1,669               & 0.0                    & 645,679   \\
+        \midrule
+        \xm & \xm   &      & 0.0 & 1,669               & 0.0                    & 665,010   \\
+        \xm & \xm   &      & 0.5 & 1,669               & 136                    & 657,010   \\
+        \xm & \xm   &      & 1.0 & 1,669               & 136                    & 657,010   \\
+        \midrule
+        \xm & \xm   & \xm  & 0.0 & 1,669               & 0.0                    & 665,705   \\
+        \xm & \xm   & \xm  & 0.5 & 1,669               & 136                    & 665,696   \\
+        \xm & \xm   & \xm  & 1.0 & 1,669               & 136                    & 665,696   \\
+    \end{tabular}
+    \caption{Analysis of model results with operating reserves but no reserve activation network constraints.
+UC = Unit Commitment constraints, DANet = Day Ahead Network constraints, PSCD = Prevent Simultaneous Charging and Discharging constraints, RSL = Reserve Shedding Limit.}\label{tab:results_no_RANet}
+\end{table}
+
+\subsection{Probabilistic operating reserves with reserve activation network constraints}
+
+\begin{itemize}
+    \item Unit commitment constraints lead to unavoidable load shedding. 
+    \item Results should be interpreted with caution however, since reserve provision is quite strict (storage can't provide reserves, non spinning reserves not modelled).
+    \item Network constraints are less problematic than unit commitment (cost increase of ~20,000 euros instead of 500,000) (perhaps thanks to storage, who knows).
+    \item Preventing simultaneous charge and discharge has a small but noticeable effect.
+\end{itemize}
+
+\begin{table}
+    \begin{tabular}{ccccccc}
+        \toprule
+        UC  & DANet & PSCD & RSL & Load shedding [MWh] & Reserve Shedding [MWh] & Objective \\
+            &       &      & 0.0 & 0.0                 & 0.0                    & 117,290   \\
+            &       &      & 0.5 & 0.0                 & 0.0                    & 117,290   \\
+            &       &      & 1.0 & 0.0                 & 0.0                    & 117,290   \\
+        \midrule
+        \xm &       &      & 0.0 & 1,669               & 0.0                    & 645,679   \\
+        \xm &       &      & 0.5 & 1,669               & 0.0                    & 645,679   \\
+        \xm &       &      & 1.0 & 1,669               & 0.0                    & 645,679   \\
+        \midrule
+        \xm & \xm   &      & 0.0 & 1,669               & 0.0                    & 665,010   \\
+        \xm & \xm   &      & 0.5 & 1,669               & 136                    & 657,010   \\
+        \xm & \xm   &      & 1.0 & 1,669               & 136                    & 657,010   \\
+        \midrule
+        \xm & \xm   & \xm  & 0.0 & 1,669               & 0.0                    & 665,705   \\
+        \xm & \xm   & \xm  & 0.5 & 1,669               & 136                    & 665,696   \\
+        \xm & \xm   & \xm  & 1.0 & 1,669               & 136                    & 665,696   \\
+    \end{tabular}
+    \caption{Analysis of model results with operating reserves but no reserve activation network constraints.
+UC = Unit Commitment constraints, DANet = Day Ahead Network constraints, PSCD = Prevent Simultaneous Charging and Discharging constraints, RSL = Reserve Shedding Limit.}\label{tab:results_no_RANet}
 \end{table}

### 2022-05-23 scripts/main_model_runs.jl
Continue writing up results

@@ -3,7 +3,7 @@ sn = script_name(@__FILE__)
 mkrootdirs(plotsdir(sn))
 
 function res_act_net(opts)
-    return opts["reserve_shedding_limit"] < 1.0 && (
+    return opts["reserve_shedding_limit"] <= 1.0 && (
         isempty(opts["upward_reserve_levels_included_in_redispatch"]) ==
         false ||
         isempty(opts["downward_reserve_levels_included_in_redispatch"]) ==
@@ -90,6 +90,18 @@ opts_vec = vcat(
         ) for opts in opts_vec if res_act_net(opts)
     ],
 )
+opts_vec = vcat(
+    opts_vec...,
+    [
+        merge(
+            opts,
+            Dict(
+                "load_multiplier" => 1.5,
+                "save_path" => "$(opts["save_path"])_LoadMult=1.5",
+            ),
+        ) for opts in opts_vec if res_act_net(opts)
+    ],
+)
 
 function main_model_run(opts; make_plots=true)
     gep = run_GEPPR(opts)
@@ -155,6 +167,7 @@ function analyse_main_model_runs(gep_vec, opts_vec)
             opts["absolute_limit_on_nodal_imbalance"] == true for
             opts in opts_vec
         ],
+        "LM" => [opts["load_multiplier"] for opts in opts_vec],
         "Reserve Shedding" =>
             sum.([
                 ismissing(gep[:rsL⁺]) ? SVC(0.0) : gep[:rsL⁺] for gep in gep_vec
@@ -163,6 +176,31 @@ function analyse_main_model_runs(gep_vec, opts_vec)
         "Objective" => [gep[:objective] for gep in gep_vec],
     )
     CSV.write(joinpath(opts["save_path"], "summary.csv"), df)
+
+    function analyse_effect_of_constraints(df, path)
+        mkrootdirs(path)
+
+        df_no_OR = filter(row -> row["OR"] == false, df)
+        CSV.write(joinpath(path, "no_OR.csv"), df_no_OR)
+
+        df_no_RANet = filter(row -> row["OR"] == true && row["RANet"] == false && row["AbsImb"] == false, df)
+        DataFrames.sort!(df_no_RANet, ["UC", "DANet", "PSCD"])
+        CSV.write(joinpath(path, "no_RANet.csv"), df_no_RANet)
+
[…]

### 2022-05-24 papers/main.tex
Saving while I retrieve a script

@@
 
+The amount of reserve shedding can be limited to a fraction of the total upward reserve requirements $RSL$:
+
+\begin{equation}
+    \sum_{n \in N, l \in L^+} rs_{nlt} \leq RSL \cdot \sum_{l \in L^+} D^{L^+}_{l't} \quad t \in T
+\end{equation}
+
 There are several matters to note here:
@@
     \item It is possible to shed upward reserves, and this is more likely to occur for higher reserve levels.
-This model is therefore able to make a tradeoff between day ahead adequacy and real time operational security, albeit crudely.
+This model is therefore able to make a trade-off between day ahead adequacy and real time operational security, albeit crudely.
     \item Shedding load in day ahead allows additional downward reserves to be provided through the variable $rc_{nlt}$.
@@
 The above constraints therefore enforce that for each reserve level $l$ and node $n$, there exists some combination of nodal imbalance, node injections, generator dispatches and line flows which would satisfy the network constraints AND the imbalance across the entire network.
+
 Given the formulation here, which uses reserve levels, i.e.
 quantiles, over the entire network to represent forecast errors, it is difficult to come up with more stringent conditions.
+However, we devised two ways of doing so.
+The first is to apply box constraints to the possible nodal imbalances:
+
+\begin{align}
+    d_{nt}^{min} \leq $d_{nlt}^{L^+}$ \leq d_{nt}^{max} & \quad n \in N, \; l \in L^+, \; t \in T
+    d_{nt}^{min} \leq $d_{nlt}^{L^-}$ \leq d_{nt}^{max} & \quad n \in N, \; l \in L^-, \; t \in T
+\end{align}
+
+It is possible to calculate $d_{nt}^{min}$ and $d_{nt}^{max}$ since the quantiles for $D^{L^+}$ and $D^{L^-}$ are calculated based on forecast error scenarios which \emph{are} defined at the nodal level.
+These limits are shown in
 
@@
 
-\begin{table}
+\begin{table}[ht]
+    \centering
     \begin{tabular}{ccccc}
@@
         UC  & DANet & PSCD & Load shedding [MWh] & Objective \\
+        \midrule
             &       &      & 0.0                 & 101,916   \\
@@
         \xm & \xm   & \xm  & 0.0                 & 252,327   \\
+        \bottomrule
     \end{tabular}
@@
     \item Unit commitment constraints lead to unavoidable load shedding. 
-    \item Results should be interpreted with caution however, since reserve provision is quite strict (storage can't provide reserves, non spinning reserves not modelled).
+    \item This result should be interpreted with caution however, since reserve provision is quite strict (storage can't provide reserves, non spinning reserves not modelled).
+If more flexibility would be available in terms of reserve provision then this might not occur.
     \item Network constraints are less problematic than unit commitment (cost increase of ~20,000 euros instead of 500,000) (perhaps thanks to storage, who knows).
@@
 
-\begin{table}
+\begin{table}[ht]
+    \centering
+    \footnotesize
     \begin{tabular}{ccccccc}
@@
         UC  & DANet & PSCD & RSL & Load shedding [MWh] & Reserve Shedding [MWh] & Objective \\
+        \midrule
             &       &      & 0.0 & 0.0                 & 0.0                    & 117,290   \\
@@
         \xm & \xm   & \xm  & 1.0 & 1,669               & 136                    & 665,696   \\
+        \bottomrule
     \end{tabular}
@@
 
-\begin{itemize}
-    \item Unit commitment constraints lead to unavoidable load shedding. 
-    \item Results should be interpreted with caution however, since reserve provision is quite strict (storage can't provide reserves, non spinning reserves not modelled).
-    \item Network constraints are less problematic than unit commitment (cost increase of ~20,000 euros instead of 500,000) (perhaps thanks to storage, who knows).
-    \item Preventing simultaneous charge and discharge has a small but noticeable effect.
-\end{itemize}
-
-\begin{table}
+\begin{table}[ht]
+    \centering
+    \footnotesize
     \begin{tabular}{ccccccc}
         \toprule
-        UC  & DANet & PSCD & RSL & Load shedding [MWh] & Reserve Shedding [MWh] & Objective \\
-            &       &      & 0.0 & 0.0                 & 0.0                    & 117,290   \\
-            &       &      & 0.5 & 0.0                 & 0.0                    & 117,290   \\
-            &       &      & 1.0 & 0.0                 & 0.0                    & 117,290   \\
+        UC  & DANet & PSCD & RSL & AbsImb & Load shedding [MWh] & Reserve Shedding [MWh] & Objective \\
         \midrule
-        \xm &       &      & 0.0 & 1,669               & 0.0                    & 645,679   \\
-        \xm &       &      & 0.5 & 1,669               & 0.0                    & 645,679   \\
-        \xm &       &      & 1.0 & 1,669               & 0.0                    & 645,679   \\
-        \midrule
-        \xm & \xm   &      & 0.0 & 1,669               & 0.0                    & 665,010   \\
-        \xm & \xm   &      & 0.5 & 1,669               & 136                    & 657,010   \\
-        \xm & \xm   &      & 1.0 & 1,669               & 136                    & 657,010   \\
-        \midrule
-        \xm & \xm   & \xm  & 0.0 & 1,669               & 0.0                    & 665,705   \\
-        \xm & \xm   & \xm  & 0.5 & 1,669               & 136                    & 665,696   \\
-        \xm & \xm   & \xm  & 1.0 & 1,669               & 136                    & 665,696   \\
+        \bottomrule
     \end{tabular}
     \caption{Analysis of model results with operating reserves but no reserve activation network constraints.
-UC = Unit Commitment constraints, DANet = Day Ahead Network constraints, PSCD = Prevent Simultaneous Charging and Discharging constraints, RSL = Reserve Shedding Limit.}\label{tab:results_no_RANet}
+UC = Unit Commitment constraints, DANet = Day Ahead Network constraints, PSCD = Prevent Simultaneous Charging and Discharging constraints, RSL = Reserve Shedding Limit, AbsImb = limit on absolute value of nodal imbalances}\label{tab:results_no_RANet}
 \end{table}

### 2022-05-24 scripts/nodal_imbalance_investigation.jl
Explain convex hull constraints

@@ -1,0 +1,62 @@
+include(joinpath(@__DIR__, "..", "intro.jl"))
+sn = script_name(@__FILE__)
+mkrootdirs(plotsdir(sn))
+
+function scatter_nodal_imbalance(dNLFE, n1, n2)
+    x = net_load_forecast_error_dict[n1][:]
+    y = net_load_forecast_error_dict[n2][:]
+    v = [[x[i], y[i]] for i in eachindex(x)]
+    hull = convex_hull(v; backend=CDDLib.Library())
+    p = vrep(hull)
+    p = removevredundancy(p, Gurobi.Optimizer)
+    p = polyhedron(p)
+    plt = Plots.scatter(x, y; xlabel=n1, ylabel=n2, lab="Imbalance [MWh]")
+    Plots.plot!(plt, p, lab="Convex hull", alpha=0.5)
+    return plt
+end
+
+opts = opts_vec[4]
+scens = load_scenarios(opts)
+mult = Dict("Load" => 1, "Wind" => -1, "Solar" => -1)
+net_load_forecast_error_dict = scenarios_2_net_load_forecast_error(
+    opts, scens, mult
+)
+
+n1 = "TIHANGE 1"
+n2 = "TIHANGE 2"
+plt = scatter_nodal_imbalance(net_load_forecast_error_dict, "TIHANGE 1", "TIHANGE 2")
+Plots.savefig(plotsdir(sn, "convex_hull_$(n1)_$(n2).png"))
+
+NLFE_vec = Dict(k => v[:] for (k, v) in net_load_forecast_error_dict)
+df = DataFrame(NLFE_vec)
+temp = mean_and_std.(eachcol(df))
+m = first.(temp)
+std = last.(temp)
+node_names = [n for n in names(df)]
+x = 0:2:length(node_names)*2
+Plots.bar(
+    x,
+    m;
+    lab="",
+    xticks=(x, node_names),
+    rotation=90,
+    margin=10mm,
+    ylabel="Mean imbalance [MWh]",
+)
+Plots.savefig(plotsdir(sn, "mean_imbalance.png"))
+# Plots.bar(m; yerror=std, xticks=[n for n in names(df)])
+
+y_min = minimum.(eachcol(df))
+y_max = maximum.(eachcol(df))
+StatsPlots.groupedbar(
+    x[1:end-1],
+    hcat(y_min, y_max);
+    lab="",
+    bar_position=:stack,
[…]

### 2022-05-24 papers/main.tex
Explain convex hull constraints

@@
 \newpage
+
+\section{Interaction description}
+
+All code can be found at \href{https://github.com/junglegobs/ASoSEPOC}{https://github.com/junglegobs/ASoSEPOC}.
 
@@
 
-Costs related to unit commitment, $z_{gt}$, have been omitted for brevity.
+Costs related to unit commitment, $z_{gt}$, have been omitted for brevity though they are included in the model.
 
@@
 \begin{align}
-    d_{nt}^{min} \leq $d_{nlt}^{L^+}$ \leq d_{nt}^{max} & \quad n \in N, \; l \in L^+, \; t \in T
-    d_{nt}^{min} \leq $d_{nlt}^{L^-}$ \leq d_{nt}^{max} & \quad n \in N, \; l \in L^-, \; t \in T
+    d_{nt}^{min} \leq d_{nlt}^{L^+} \leq d_{nt}^{max} & \quad n \in N, \; l \in L^+, \; t \in T \\
+    d_{nt}^{min} \leq d_{nlt}^{L^-} \leq d_{nt}^{max} & \quad n \in N, \; l \in L^-, \; t \in T
 \end{align}
@@
 It is possible to calculate $d_{nt}^{min}$ and $d_{nt}^{max}$ since the quantiles for $D^{L^+}$ and $D^{L^-}$ are calculated based on forecast error scenarios which \emph{are} defined at the nodal level.
-These limits are shown in
+These limits are shown in Figure \ref{fig:imbalance_range}.
+Clearly it is simply not possible to have an imbalance on some nodes, such as Coo (the location of Belgium's pumped hydro unit).
+Constraining $d_{nlt}^{L^+}$ and $d_{nlt}^{L^-}$ in this way is referred to as AbsImb later on.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.9\textwidth]{../plots/nodal_imbalance_investigation/imbalance_range}
+    \caption{Range of possible imbalances for all hours of day 309.\label{fig:imbalance_range}}
+\end{figure}
+
+The other possibility is to restrict $d_{nlt}^{L^+}$ and $d_{nlt}^{L^-}$ to lie within the convex hull of all imbalances.
+This is illustrated in 2 dimensions in Figure \ref{fig:convex_hull_tihange} for the imbalances at Tihange 1 and 2.
+Clearly the imbalances are highly correlated, and constraining the possible nodal imbalances to lie within the convex hull exploits this in a way that box constraints would not be able to.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.9\textwidth]{../plots/nodal_imbalance_investigation/convex_hull_TIHANGE 1_TIHANGE 2}
+    \caption{Illustration of convex hull constraints on $d_{nlt}^{L^+}$ and $d_{nlt}^{L^-}$.
+Blue circles are imbalance scenarios, shaded area is the convex hull, $d_{nt}^{min}$ and $d_{nt}^{max}$ are the x limits for TIHANGE 1 and the y limits for TIHANGE 2.
+\label{fig:convex_hull_tihange}}
+\end{figure}
 
 \section{Input data - stylised Belgian grid with a large amount of variable renewable energy sources}
+
+\todo[inline]{For now, see README.md}
 
@@
     \footnotesize
-    \begin{tabular}{ccccccc}
+    \begin{tabular}{cccccccc}
         \toprule

### 2022-05-25 papers/main.tex
Noticed that load shed where there is no load

@@
 \section{Interaction description}
+
+\todo[inline]{Fill in}
 
@@
 
+\subsubsection{Day ahead constraints}
+
 The power balance:
@@
 For brevity and clarity, constraints on ramping and minimum up and down times are omitted.
+
+\subsubsection{Operating reserve (activation) constraints}
 
@@
 
+\subsubsection{Operating reserve activation network constraints} \label{sec:operating_reserve_network_activation_constraints}
+
 The following constraints attempt to take network constraints into account (albeit very weakly):
@@
 \begin{align}
-    \sum_{g \in G_n, l'=1:l} (r^{L^+}_{glt} + rs_{nlt}) = d^{L^+}_{nlt} + rinj^{L^+}_{nlt}   & \quad n \in N, \; l \in L^+, \; t \in T \\
+    \sum_{g \in G_n, l'=1:l} (r^{L^+}_{glt} + rs_{nlt}) = d^{L^+}_{nlt} + rinj^{L^+}_{nlt}   & \quad n \in N, \; l \in L^+, \; t \in T \label{eq:reserve_network_activation_constraints_1} \\
     - \sum_{g \in G_n, l'=1:l} (r^{L^-}_{glt} + rc_{nlt}) = d^{L^-}_{nlt} + rinj^{L^-}_{nlt} & \quad n \in N, \; l \in L^-, \; t \in T                                                     \\
@@
     \sum_{n \in N} rinj^{L^+}_{nt} = 0                                                       & \quad l \in L^+, \; n \in N, \; t \in T                                                     \\
-    \sum_{n \in N} rinj^{L^-}_{nt} = 0                                                       & \quad l \in L^-, \; n \in N, \; t \in T
+    \sum_{n \in N} rinj^{L^-}_{nt} = 0                                                       & \quad l \in L^-, \; n \in N, \; t \in T \label{eq:reserve_network_activation_constraints_2}
 \end{align}
@@
 The above constraints therefore enforce that for each reserve level $l$ and node $n$, there exists some combination of nodal imbalance, node injections, generator dispatches and line flows which would satisfy the network constraints AND the imbalance across the entire network.
+
+\subsubsection{Further constraining reserve activation network constraints} \label{sec:further_constraining_operating_reserve_network_activation_constraints}
 
@@
 
+At the time of writing, this convex hull restriction led to model infeasibilities and so it is not
+
 \section{Input data - stylised Belgian grid with a large amount of variable renewable energy sources}
@@
 While the objective more than doubles going from the simple linear economic dispatch model with no network constraints to the unit commitment model with network constraints, no load shedding occurs.
-Preventing simultaneous charging and discharging constraints does not affect results at all, implying that additional energy consumption from storage does not aid congestion.
+Preventing simultaneous charging and discharging constraints does not affect results at all, implying that additional energy consumption (which is not physically possible) from storage does not aid congestion in this context.
 
-\begin{table}[ht]
+\begin{table}[H]
     \centering
@@
 
+The dispatch for the model with and without unit commitment constraints and with network constraints is shown in \figref{fig:dispatch_base}.
+The addition of unit commitment constraints leads to storage dispatching and conventional units generating throughout the day.
+Interestingly, the amount of solar generation and storage dispatch is higher in the economic dispatch model and the model with unit commitment and network constraints.
+This may be because batteries and solar generators are located in buses where there is also load, favouring their usage.
+
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}[t]{0.7\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base/dispatch}
+        \caption{Economic dispatch model}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.7\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_UC=true/dispatch}
+        \caption{Unit commitment model}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.7\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_UC=true_DANet=true/dispatch}
+        \caption{Unit commitment model with day ahead network constraints}
+    \end{subfigure}
+    \caption{Dispatch schedules with and without operating reserves.\label{fig:dispatch_base}}
+\end{figure}
+
 \subsection{Probabilistic operating reserves with no reserve activation network constraints}
@@
 If more flexibility would be available in terms of reserve provision then this might not occur.
-    \item Network constraints are less problematic than unit commitment (cost increase of ~20,000 euros instead of 500,000) (perhaps thanks to storage, who knows).
+    \item Network constraints are less problematic than unit commitment (cost increase of \~20,000 euros instead of \~500,000).
     \item Preventing simultaneous charge and discharge has a small but noticeable effect.
@@
 
+\figref{fig:dispatch_simple_reserves} shows the dispatch for 3 models with operating reserves, the first a simple economic dispatch, then a unit commitment and then a unit commitment with network constraints.
+Similarly to before, network constraints lead to greater usage of solar and storage compared to the unit commitment model without.
+
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}[t]{0.7\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_RSV=0.0/dispatch}
+        \caption{No UC, no DANet, no PSCD, RSL is 0.0}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.7\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_UC=true_RSV=0.0/dispatch}
+        \caption{UC, no DANet, no PSCD, RSL is 0.0}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.7\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_UC=true_DANet=true_RSV=0.0/dispatch}
+        \caption{UC, DANet, no PSCD, RSL is 0.0}
+    \end{subfigure}
+    \caption{Dispatch schedules with and without operating reserves.\label{fig:dispatch_simple_reserves}}
+\end{figure}
+
+\begin{itemize}
+    \item What's happening to reduce reserve shedding without reducing load shedding? Consider only the difference between $RSL$ 0.5 and 0.0 for the UC and DANet case.
+    \item For $RSL = 0.5 $, load shedding occurs across all nodes for the 10th hour of the day. Reserve shedding occurs on one node, LATOUR, for the 9th hour.
+    \item The commitment for the two values of $RSL$ is not different.
+    \item The load shed in both cases is the same i.e. it is not shed in different places. It is also equal everywhere, which seems strange...
+    \item To avoid reserve shedding, the cost increase appears to be simply from regulating the generation of 6 conventional generators - Eeklo Nord, Meerhout, Monce and the generators at TIHANGE 1 and TIHANGE 3.
+Various renewable generators also modify their output.
+    \item This leads me to believe that economic load shedding is occurring, and that more conventional generators are generating and providing reserves in order to satisfy the reserve shedding limit.
+\end{itemize}
+
+The fact that upward operating reserves can be entirely satisfied while load shedding occurs signals to me that reserve provision is easier to provide than satisfying load.
+This is almost certainly due to the weak reserve activation network constraints.
+Further constraining these using the methods described in \secref{sec:further_constraining_operating_reserve_network_activation_constraints} attempts to address this, as will be detailed in the next section.
+
 \subsection{Probabilistic operating reserves with reserve activation network constraints}
+
+Including network reserve activation constraints, that is Constraints \ref{eq:reserve_network_activation_constraints_1} through \ref{eq:reserve_network_activation_constraints_1}, led to no discernible change in objective.
+This particular instance is not able to make a trade-off between day ahead adequacy and real time operational security.
+As a first attempt to overcome this issue, the load at all nodes was multiplied by a factor of 1.5.
+This gave the results shown in \tabref{tab:results_LM_1_5}.
 
@@
         \toprule
-        UC  & DANet & PSCD & RSL & AbsImb & Load shedding [MWh] & Reserve Shedding [MWh] & Objective \\
+        UC  & DANet & PSCD & AbsImb & RSL & Load shedding [MWh] & Reserve Shedding [MWh] & Objective \\
         \midrule
+            &       &      &        & 0.0 & 34,860              & 0                      & 7,827,724 \\
+            &       &      &        & 0.5 & 16,346              & 19,257                 & 3,934,989 \\
+            &       &      &        & 1.0 & 7,521               & 28,209                 & 2,463,656 \\
+        \midrule
+        \xm &       &      &        & 0.0 & 34,860              & 0                      & 7,969,802 \\
+        \xm &       &      &        & 0.5 & 16,672              & 19,002                 & 4,133,637 \\
+        \xm &       &      &        & 1.0 & 9,190               & 26,809                 & 2,892,401 \\
+        \midrule
+        \xm & \xm   &      &        & 0.0 & 42,008              & 0                      & 9,572,369 \\
+        \xm & \xm   &      &        & 0.5 & 16,950              & 28,096                 & 4,310,407 \\
+        \xm & \xm   &      &        & 1.0 & 15,482              & 29,604                 & 4,029,268 \\
+        \midrule
+        \xm & \xm   & \xm  &        & 0.0 & 42,008              & 0                      & 9,572,828 \\
+        \xm & \xm   & \xm  &        & 0.5 & 16,950              & 28,096                 & 4,310,407 \\
+        \xm & \xm   & \xm  &        & 1.0 & 15,482              & 29,604                 & 4,029,268 \\
+        \midrule
+        \xm & \xm   & \xm  & \xm    & 0.0 & 42,008              & 0                      & 9,572,828 \\
+        \xm & \xm   & \xm  & \xm    & 0.5 & 16,950              & 28,096                 & 4,310,407 \\
+        \xm & \xm   & \xm  & \xm    & 1.0 & 15,482              & 29,604                 & 4,029,268 \\
         \bottomrule
     \end{tabular}
-    \caption{Analysis of model results with operating reserves but no reserve activation network constraints.
-UC = Unit Commitment constraints, DANet = Day Ahead Network constraints, PSCD = Prevent Simultaneous Charging and Discharging constraints, RSL = Reserve Shedding Limit, AbsImb = limit on absolute value of nodal imbalances}\label{tab:results_no_RANet}
+    \caption{Analysis of model results with operating reserves and with reserve activation network constraints with load multiplied by a factor of 1.5.
+UC = Unit Commitment constraints, DANet = Day Ahead Network constraints, PSCD = Prevent Simultaneous Charging and Discharging constraints, RSL = Reserve Shedding Limit.}\label{tab:results_no_RANet}
 \end{table}
 
+With reference to \tabref{tab:results_LM_1_5}:
+
+\begin{itemize}
+    \item Even in this case, the reserve activation network constraints are not binding, since the AbsImb constraints do not alter the objective function.
+    \item On the other hand, the trade-off between reserve shedding and load shedding can be seen now in all cases. This trade-off is also obvious when observing the objective function.
+    \item Unit commitment constraints do relatively little to increase total costs, but this is unsurprising given that costs are dominated by load shedding.
+    \item Day ahead network constraints increase load shedding considerably i.e. congestion also plays a role in increasing load shedding.
+\end{itemize}
+
+Closer investigation at load shedding patterns for the model with UC and DANet (but no PSCD and AbsImb) reveals that all load shedding occurs at the start of the day (when the battery has not charged yet).
+
+\todo[inline]{Load is shed at nodes where no load exists, which is clearly unphysical.}
+
 \end{document}

### 2022-05-25 scripts/main_model_runs.jl
Noticed that load shed where there is no load

@@ -3,12 +3,17 @@ sn = script_name(@__FILE__)
 mkrootdirs(plotsdir(sn))
 
 function res_act_net(opts)
-    return opts["reserve_shedding_limit"] <= 1.0 && (
+    return (
+        opts["copperplate"] == false &&
         isempty(opts["upward_reserve_levels_included_in_redispatch"]) ==
         false ||
         isempty(opts["downward_reserve_levels_included_in_redispatch"]) ==
         false
     )
+end
+
+function has_OR(gep::GEPM)
+    return isnothing(gep_1[:I, :uncertainty]) == false
 end
 
 opts = options_diff_days(sn)[4]
@@ -75,7 +80,7 @@ opts_vec = vcat(
                 "downward_reserve_levels_included_in_redispatch" => 1:10,
                 "save_path" => "$(opts["save_path"])_L⁺=1:10_L⁻=1:10",
             ),
-        ) for opts in opts_vec if opts["reserve_shedding_limit"] < 1.0
+        ) for opts in opts_vec if opts["reserve_shedding_limit"] <= 1.0
     ]...,
 )
 opts_vec = vcat(
@@ -103,17 +108,20 @@ opts_vec = vcat(
     ],
 )
 
-function main_model_run(opts; make_plots=true)
+function main_model_run(opts; make_plots=false)
     gep = run_GEPPR(opts)
 
-    if make_plots
+    dis_file = joinpath(opts["save_path"], "dispatch.png")
+    res_file = joinpath(opts["save_path"], "ls_and_rs.png")
+
+    if make_plots && (isfile(dis_file) == false && isfile(res_file) == false)
         # Plot dispatch
         plt = plot_dispatch(gep, 1)
-        Plots.savefig(plt, joinpath(opts["save_path"], "dispatch.png"))
+        Plots.savefig(plt, dis_file)
 
         # If has reserves, plot these
         if opts["operating_reserves_type"] != "none"
-            apply_operating_reserves!(gep, opts)
+            has_OR(gep) == false && apply_operating_reserves!(gep, opts)
             plt = plot_reserves_simple(gep, 1)
             Plots.savefig(plt, joinpath(opts["save_path"], "reserves.png"))
         end
@@ -134,16 +142,18 @@ function main_model_run(opts; make_plots=true)
[…]

### 2022-05-25 scripts/main_model_runs.jl
Investigate load shedding patterns

@@ -148,8 +148,6 @@ function main_model_run(opts; make_plots=false)
     return gep
 end
 
-main_model_run(opts_vec[63])
-
 gep_vec = GEPM[]
 for opts in opts_vec
     try
@@ -293,8 +291,8 @@ end
 # how_is_reserve_shed_without_increasing_load_shedding(gep_vec, opts_vec)
 
 function load_shedding_distribution(gep_vec, opts_vec, idx)
-    plotly()
-    gep = gep_vec[63]
+    pyplot()
+    gep = gep_vec[idx]
 
     ls = dropdims(gep[:loadShedding].data; dims=(2, 3))
     plt = Plots.plot(
@@ -318,7 +316,6 @@ function load_shedding_distribution(gep_vec, opts_vec, idx)
     Plots.savefig(plt, plotsdir(sn, "ls_per_node_idx=$(idx).png"))
 
     D = dropdims(GEPPR.get_demand(gep); dims=(2, 3))
-
     norm_ls = (sum(ls; dims=2) ./ sum(D; dims=2))'
     plt = Plots.bar(
         hcat(x...),
@@ -339,6 +336,11 @@ function load_shedding_distribution(gep_vec, opts_vec, idx)
         markershape=:diamond,
         markercolor=:red,
         markersize=4,
+        lab="Unphysical load shedding"
     )
-    return Plots.savefig(plt, plotsdir(sn, "ls_per_node_idx=$(idx).png"))
+    Plots.savefig(plt, plotsdir(sn, "ls_per_node_idx=$(idx).png"))
+
+    return nothing
 end
+
+load_shedding_distribution(gep_vec, opts_vec, 63)

### 2022-05-25 scripts/main_model_runs.jl
Change initial state of charge

@@ -13,11 +13,11 @@ function res_act_net(opts)
 end
 
 function has_OR(gep::GEPM)
-    return isnothing(gep_1[:I, :uncertainty]) == false
+    return isnothing(gep[:I, :uncertainty]) == false
 end
 
 opts = options_diff_days(sn)[4]
-opts["initial_state_of_charge"] = 0.0
+opts["initial_state_of_charge"] = 0.5
 opts_vec = [
     merge(
         opts,

### 2022-05-27 papers/main.tex
Change initial state of charge
Alter reserve shedding sensitivity analysis

@@
         \midrule
-            &       &      & 0.0                 & 101,916   \\
-        \xm &       &      & 0.0                 & 212,558   \\
-        \xm & \xm   &      & 0.0                 & 252,327   \\
-        \xm & \xm   & \xm  & 0.0                 & 252,327   \\
+            &       &      & 0.0                 & 101,915   \\
+        \xm &       &      & 0.0                 & 109,148   \\
+        \xm & \xm   &      & 0.0                 & 227,495   \\
+        \xm & \xm   & \xm  & 0.0                 & 227,492   \\
         \bottomrule
@@
         \midrule
-            &       &      & 0.0 & 0.0                 & 0.0                    & 117,290   \\
-            &       &      & 0.5 & 0.0                 & 0.0                    & 117,290   \\
-            &       &      & 1.0 & 0.0                 & 0.0                    & 117,290   \\
+            &       &      & 0.0 & 0.0                 & 0.0                    & 101,345   \\
+            &       &      & 0.5 & 0.0                 & 0.0                    & 101,014   \\
+            &       &      & 1.0 & 0.0                 & 0.0                    & 101,014   \\
         \midrule
-        \xm &       &      & 0.0 & 1,669               & 0.0                    & 645,679   \\
-        \xm &       &      & 0.5 & 1,669               & 0.0                    & 645,679   \\
-        \xm &       &      & 1.0 & 1,669               & 0.0                    & 645,679   \\
+        \xm &       &      & 0.0 & 1.669               & 0.0                    & 617,536   \\
+        \xm &       &      & 0.5 & 1.669               & 0.0                    & 617,536   \\
+        \xm &       &      & 1.0 & 1.669               & 0.0                    & 617,536   \\
         \midrule
-        \xm & \xm   &      & 0.0 & 1,669               & 0.0                    & 665,010   \\
-        \xm & \xm   &      & 0.5 & 1,669               & 136                    & 657,010   \\
-        \xm & \xm   &      & 1.0 & 1,669               & 136                    & 657,010   \\
+        \xm & \xm   &      & 0.0 & 1.669               & 0.0                    & 650,647   \\
+        \xm & \xm   &      & 0.5 & 1.669               & 136                    & 650,231   \\
+        \xm & \xm   &      & 1.0 & 1.669               & 136                    & 650,231   \\
         \midrule
-        \xm & \xm   & \xm  & 0.0 & 1,669               & 0.0                    & 665,705   \\
-        \xm & \xm   & \xm  & 0.5 & 1,669               & 136                    & 665,696   \\
-        \xm & \xm   & \xm  & 1.0 & 1,669               & 136                    & 665,696   \\
+        \xm & \xm   & \xm  & 0.0 & 1.669               & 0.0                    & 650,647   \\
+        \xm & \xm   & \xm  & 0.5 & 1.669               & 136                    & 650,231   \\
+        \xm & \xm   & \xm  & 1.0 & 1.669               & 136                    & 650,263   \\
         \bottomrule
@@
         \midrule
-            &       &      &        & 0.0 & 34,860              & 0                      & 7,827,724 \\
-            &       &      &        & 0.5 & 16,346              & 19,257                 & 3,934,989 \\
-            &       &      &        & 1.0 & 7,521               & 28,209                 & 2,463,656 \\
+            &       &      &        & 0.0 & 32,366              & 0                      & 7,290,909 \\
+            &       &      &        & 0.5 & 0.0                 & 36,269                 & 447,102   \\
+            &       &      &        & 1.0 & 0.0                 & 36,082                 & 440,467   \\
         \midrule
-        \xm &       &      &        & 0.0 & 34,860              & 0                      & 7,969,802 \\
-        \xm &       &      &        & 0.5 & 16,672              & 19,002                 & 4,133,637 \\
-        \xm &       &      &        & 1.0 & 9,190               & 26,809                 & 2,892,401 \\
+        \xm &       &      &        & 0.0 & 32,366              & 0                      & 7,433,785 \\
+        \xm &       &      &        & 0.5 & 1,669               & 34,379                 & 924,686   \\
+        \xm &       &      &        & 1.0 & 1,669               & 34,308                 & 919,363   \\
         \midrule
-        \xm & \xm   &      &        & 0.0 & 42,008              & 0                      & 9,572,369 \\
-        \xm & \xm   &      &        & 0.5 & 16,950              & 28,096                 & 4,310,407 \\
-        \xm & \xm   &      &        & 1.0 & 15,482              & 29,604                 & 4,029,268 \\
+        \xm & \xm   &      &        & 0.0 & 33,929              & 0                      & 7,822,959 \\
+        \xm & \xm   &      &        & 0.5 & 1,967               & 39,234                 & 1,249,422 \\
+        \xm & \xm   &      &        & 1.0 & 1,941               & 39,160                 & 1,245,647 \\
         \midrule
-        \xm & \xm   & \xm  &        & 0.0 & 42,008              & 0                      & 9,572,828 \\
-        \xm & \xm   & \xm  &        & 0.5 & 16,950              & 28,096                 & 4,310,407 \\
-        \xm & \xm   & \xm  &        & 1.0 & 15,482              & 29,604                 & 4,029,268 \\
+        \xm & \xm   & \xm  &        & 0.0 & 33,929              & 0                      & 7,822,959 \\
+        \xm & \xm   & \xm  &        & 0.5 & 1,967               & 39,234                 & 1,249,422 \\
+        \xm & \xm   & \xm  &        & 1.0 & 1,941               & 39,160                 & 1,245,647 \\
         \midrule
-        \xm & \xm   & \xm  & \xm    & 0.0 & 42,008              & 0                      & 9,572,828 \\
-        \xm & \xm   & \xm  & \xm    & 0.5 & 16,950              & 28,096                 & 4,310,407 \\
-        \xm & \xm   & \xm  & \xm    & 1.0 & 15,482              & 29,604                 & 4,029,268 \\
+        \xm & \xm   & \xm  & \xm    & 0.0 & 33,929              & 0                      & 7,822,959 \\
+        \xm & \xm   & \xm  & \xm    & 0.5 & 1,967               & 39,234                 & 1,249,422 \\
+        \xm & \xm   & \xm  & \xm    & 1.0 & 1,941               & 39,160                 & 1,245,647 \\
         \bottomrule
@@
 \end{table}
+
+\todo[inline]{Objective is weird for some cases...}
 
@@
 
-Closer investigation at load shedding patterns for the model with UC and DANet (but no PSCD and AbsImb) reveals that all load shedding occurs at the start of the day (when the battery has not charged yet).
+% Closer investigation at load shedding patterns for the model with UC and DANet (but no PSCD and AbsImb) reveals that all load shedding occurs at the start of the day (when the battery has not charged yet).
 
-\todo[inline]{Load is shed at nodes where no load exists, which is clearly unphysical.}
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{../plots/main_model_runs/ls_per_node_idx=63}
+        \caption{}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{../plots/main_model_runs/ls_per_node_idx=63}
+        \caption{}
+    \end{subfigure}
+    \caption{}
+\end{figure}
 

### 2022-05-27 scripts/DUCPR_reserve_shedding_sensitivity.jl

@@ -2,25 +2,28 @@ include(joinpath(@__DIR__, "..", "intro.jl"))
 sn = script_name(@__FILE__)
 mkrootdirs(plotsdir(sn))
 
-opts_vec = options_diff_days(sn)
-opts_vec = vcat(
-    [
-        merge(
-            opts,
-            Dict(
-                "reserve_shedding_limit" => v,
-                "save_path" => joinpath(opts["save_path"], "RSL=$(v)_L=$L"),
-                "time_out" => 600,
-                "upward_reserve_levels_included_in_redispatch" => L,
-                "downward_reserve_levels_included_in_redispatch" => L,
-            ),
-        ) for opts in opts_vec, L in [Int[], 1:10], v in 0.1:-0.02:0
-        # ) for v in [0.02], opts in opts_vec
-    ]...,
+opts = options_diff_days(sn)[4]
+opts = merge(
+    opts,
+    Dict(
+        "time_out" => 600,
+        "upward_reserve_levels_included_in_redispatch" => 1:10,
+        "downward_reserve_levels_included_in_redispatch" => 1:10,
+        "initial_state_of_charge" => 0.5,
+        "load_multiplier" => 1.5,
+        "absolute_limit_on_nodal_imbalance" => true,
+    ),
 )
-# gep = run_GEPPR(opts_vec[1])
-# gep_vec = run_GEPPR(opts_vec)
-gep_vec = run_GEPPR(opts_vec; load_only=true)
+opts_vec = vcat(
+    [merge(
+        opts,
+        Dict(
+            "reserve_shedding_limit" => v,
+            "save_path" => joinpath(opts["save_path"], "RSL=$(v)_L=$L"),
+        ),
+    ) v in 1.0:-0.1:0]...,
+)
+gep_vec = run_GEPPR(opts_vec)
 GC.gc() # Who knows, maybe this will help
 map(
     i -> if gep_vec[i] === nothing
@@ -31,117 +34,84 @@ map(
     eachindex(opts_vec),
 )
 
-sid_vec = [month_day(opts_vec[i]) for i in 1:length(opts_vec)]
 rsl_vec = [opts_vec[i]["reserve_shedding_limit"] for i in 1:length(opts_vec)]
-L_vec = [
[…]

### 2022-05-27 scripts/main_model_runs.jl
Explain interaction

@@ -305,7 +305,7 @@ function load_shedding_distribution(gep_vec, opts_vec, idx)
     plt = Plots.bar(
         hcat(x...),
         sum(ls; dims=2)';
-        ylab="Load shedding / Load [-]",
+        ylab="Load shedding [MWh]",
         xticks=(x, N),
         lab="",
         tickfontrotation=90,
@@ -313,7 +313,7 @@ function load_shedding_distribution(gep_vec, opts_vec, idx)
         lw=0.0,
         c=:blue,
     )
-    Plots.savefig(plt, plotsdir(sn, "ls_per_node_idx=$(idx).png"))
+    Plots.savefig(plt, plotsdir(sn, "ls_per_node_abs_idx=$(idx).png"))
 
     D = dropdims(GEPPR.get_demand(gep); dims=(2, 3))
     norm_ls = (sum(ls; dims=2) ./ sum(D; dims=2))'
@@ -343,4 +343,4 @@ function load_shedding_distribution(gep_vec, opts_vec, idx)
     return nothing
 end
 
-load_shedding_distribution(gep_vec, opts_vec, 63)
+load_shedding_distribution(gep_vec, opts_vec, 56)

### 2022-05-27 papers/main.tex
Explain interaction

@@
 
+\input{acronyms}
+
 \newpage
@@
 
-\todo[inline]{Fill in}
+Adequacy assessments and operational security analyses are two fundamental exercises in determining and ensuring the reliable operation of the electric power system.
+The former assesses the ability of the power system to satisfy demand for electricity in the day ahead stage, while the latter assesses (or ensures, depending on the context) that an unexpected event, such as an outage of a transmission line or generator or lower or greater than expected realisation of demand or renewable generation, does not lead to cascading failures and a possible blackout.
+
+Typically, these exercises are performed independently. This interaction combines the two by feeding the results of an adequacy assessment to an operational security analysis.
+
+In doing so, we deviate from typical practices in a number of ways.
+First, the adequacy assessment is done for a limited number of days instead of many Monte Carlo years for the sake of clarity and brevity.
+Indeed, doing this exercise for many Monte Carlo years would drastically increase the computational complexity while the aim of this exercise is to illustrate how such an interaction could be performed.
+
+Secondly, the adequacy assessment is done using an unconventional \ac{UC} model, a modified version of the \ac{DUCPR} model described in \cite{Bruninx2017}.
+This model can be seen as a compromise between a computationally expensive stochastic \ac{UC} model and a more tractable though less accurate deterministic \ac{UC} model with operating reserve requirements.
+The treatment of operating reserves in this model allows a trade-off between shedding load in the day ahead stage, which would reduce day ahead adequacy, and shedding operating reserves, which should reduce operational security.
+By making this trade-off explicit and verifying the results in an operational security analysis we aim to elaborate on the results of \cite{Hermans2018}, which did not consider the possibility that shedding operating reserves would lead to insecure operation of the electric power system.
 
@@
 \section{Description of a Deterministic Unit Commitment Model with Probabilistic Reserve (DUC-PR) constraints}
+
+Sets, variables and costs related to unit commitment and storage operation have been omitted for brevity though they are included in the model.
+The interested reader is referred to \href{https://github.com/junglegobs/ASoSEPOC}{https://github.com/junglegobs/ASoSEPOC} for the relevant code.
 
@@
 
-All variables are positive apart from injetion variables which may also be negative.
+All variables are positive apart from node injection variables which may also be negative.
 
@@
 From the top line to the bottom, the costs are those of dispatching generators and activating upwards or downwards reserves.
-
-Costs related to unit commitment, $z_{gt}$, have been omitted for brevity though they are included in the model.
 
@@
 
-% Closer investigation at load shedding patterns for the model with UC and DANet (but no PSCD and AbsImb) reveals that all load shedding occurs at the start of the day (when the battery has not charged yet).
+Closer investigation at load shedding patterns for the model with UC and DANet (but no PSCD and AbsImb) and $RSL = 0.0$ reveals that load shedding is at it's highest in the mornings (8 - 10 AM) and in the evenings (17 - 20).
+No load shedding occurs in the middle of the day, which happens to be when load is lowest and solar generation is greatest.
+Load shedding is roughly equally distributed throughout the day in absolute though not relative terms.
 
@@
         \centering
-        \includegraphics[width=\textwidth]{../plots/main_model_runs/ls_per_node_idx=63}
+        \includegraphics[width=\textwidth]{../plots/main_model_runs/ls_timeseries_idx=56}
+        \caption{}
+    \end{subfigure}
+    \newline
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{../plots/main_model_runs/ls_per_node_abs_idx=56}
         \caption{}
@@
         \centering
-        \includegraphics[width=\textwidth]{../plots/main_model_runs/ls_per_node_idx=63}
+        \includegraphics[width=\textwidth]{../plots/main_model_runs/ls_per_node_idx=56}
         \caption{}
     \end{subfigure}
-    \caption{}
+    \caption{Load shedding distribution in time (top figure, aggregated over all nodes) and space (aggregated over the entire day) for UC, DANet, LM = 1.5 and $RSL = 0.5$.}
 \end{figure}

### 2022-05-27 scripts/DUCPR_reserve_shedding_sensitivity.jl

@@ -14,15 +14,15 @@ opts = merge(
         "absolute_limit_on_nodal_imbalance" => true,
     ),
 )
-opts_vec = vcat(
-    [merge(
+opts_vec = vcat([
+    merge(
         opts,
         Dict(
             "reserve_shedding_limit" => v,
             "save_path" => joinpath(opts["save_path"], "RSL=$(v)_L=$L"),
         ),
-    ) v in 1.0:-0.1:0]...,
-)
+    ) for v in 1.0:-0.1:0
+])
 gep_vec = run_GEPPR(opts_vec)
 GC.gc() # Who knows, maybe this will help
 map(

### 2022-06-14 papers/main.tex
Add more description of data

@@
 \title{Resource Adequacy and Operational Security Interaction in the EPOC 2030-50 Project}
+
+\author{Sebastian Gonzato}
 
@@
 
-\section{Description of a Deterministic Unit Commitment Model with Probabilistic Reserve (DUC-PR) constraints}
+The rest of this document is organised as follows.
+\secref{sec:model} describes the \ac{DUCPR} model used to simulate power system operations and obtain day ahead adequacy and very approximate real time operational security indicators, namely scheduled load shedding and reserve shedding.
+\secref{sec:data} describes the stylised Belgian system analysed.
+\secref{sec:initial_results} analyses a single day in order to identify sources of adequacy and security issues, while \secref{sec:adequacy_vs_security} analyses the trade-off between these two.
+
+\section{Description of a Deterministic Unit Commitment Model with Probabilistic Reserve (DUC-PR) constraints} \label{sec:model}
 
@@
     \item $GD$ - Dispatchable / thermal / conventional generators
-    \item $GR$ - Renewable / variable generators
+    \item $GR$ - Renewable / variable generators (\ac{VRES}).
     \item $N$ - Nodes or buses in the network
@@
     \centering
-    \includegraphics[width=0.9\textwidth]{../plots/nodal_imbalance_investigation/imbalance_range}
+    % \includegraphics[width=0.9\textwidth]{../plots/nodal_imbalance_investigation/imbalance_range}
     \caption{Range of possible imbalances for all hours of day 309.\label{fig:imbalance_range}}
@@
     \centering
-    \includegraphics[width=0.9\textwidth]{../plots/nodal_imbalance_investigation/convex_hull_TIHANGE 1_TIHANGE 2}
+    % \includegraphics[width=0.9\textwidth]{../plots/nodal_imbalance_investigation/convex_hull_TIHANGE 1_TIHANGE 2}
     \caption{Illustration of convex hull constraints on $d_{nlt}^{L^+}$ and $d_{nlt}^{L^-}$.
@@
 
-\section{Input data - stylised Belgian grid with a large amount of variable renewable energy sources}
+\section{Input data - stylised Belgian grid with a large amount of variable renewable energy sources} \label{sec:data}
 
-\todo[inline]{For now, see README.md}
+All data used for this exercise can be found at: \href{https://www.dropbox.com/sh/mdvmc082gwng0tr/AABRyc3fZpxAFycmUfZmh8Csa?dl=0}.
 
-\section{Analysis of DUC-PR model results}
+The grid and resource data used for this interaction is inspired by the case study in \cite{Belderbos2020}, omitting the gas network and power to gas technologies.
+This data gives a stylised Belgian system with a high penetration (~8s0\%) of \ac{VRES}, which includes 75.7 GW of solar PV, 7.0 GW of Onshore Wind and 7.3 GW of Offshore Wind.
+The grid consists of 46 nodes connected by 69 lines.
+The total amount of conventional thermal generation is 9.9 GW, with all generators based on new \ac{CCGT} units apart from one new \ac{OCGT} unit at Lixhe.
+The nuclear generators of Doel and Tihange are therefore replaced with new \ac{CCGT} units.
 
-This analysis relates only to the 309th day of the year (November 5th).
+The residual load time series, aggregated across all nodes, for this system is plotted in \todo[inline]{Plot this}.
+
+A complete overview of the data is given in \cite{Belderbos2020}, Appendix B.
+
+\section{Analysis of \acl{DUCPR} model results} \label{sec:initial_results}
+
+This analysis relates only to the 309th day of the year (November 5th). The aim is to investigate what are the causes of load shedding.
 
@@
         \centering
-        \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base/dispatch}
+        % \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base/dispatch}
         \caption{Economic dispatch model}
@@
         \centering
-        \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_UC=true/dispatch}
+        % \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_UC=true/dispatch}
         \caption{Unit commitment model}
@@
         \centering
-        \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_UC=true_DANet=true/dispatch}
+        % \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_UC=true_DANet=true/dispatch}
         \caption{Unit commitment model with day ahead network constraints}
@@
 
-\subsection{Probabilistic operating reserves with no reserve activation network constraints}
+\subsection{Probabilistic operating reserves without reserve activation network constraints}
 
@@
         \centering
-        \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_RSV=0.0/dispatch}
+        % \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_RSV=0.0/dispatch}
         \caption{No UC, no DANet, no PSCD, RSL is 0.0}
@@
         \centering
-        \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_UC=true_RSV=0.0/dispatch}
+        % \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_UC=true_RSV=0.0/dispatch}
         \caption{UC, no DANet, no PSCD, RSL is 0.0}
@@
         \centering
-        \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_UC=true_DANet=true_RSV=0.0/dispatch}
+        % \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_UC=true_DANet=true_RSV=0.0/dispatch}
         \caption{UC, DANet, no PSCD, RSL is 0.0}
@@
         \centering
-        \includegraphics[width=\textwidth]{../plots/main_model_runs/ls_timeseries_idx=56}
+        % \includegraphics[width=\textwidth]{../plots/main_model_runs/ls_timeseries_idx=56}
         \caption{}
@@
         \centering
-        \includegraphics[width=\textwidth]{../plots/main_model_runs/ls_per_node_abs_idx=56}
+        % \includegraphics[width=\textwidth]{../plots/main_model_runs/ls_per_node_abs_idx=56}
         \caption{}
@@
         \centering
-        \includegraphics[width=\textwidth]{../plots/main_model_runs/ls_per_node_idx=56}
+        % \includegraphics[width=\textwidth]{../plots/main_model_runs/ls_per_node_idx=56}
         \caption{}
@@
 
+\section{Day ahead adequacy and real time operational security trade offs} \label{sec:adequacy_vs_security}
+
 \end{document}

### 2022-06-14 papers/main.tex
Mention storage amount

@@
 
+Power to Gas is omitted but 14.1 GW of storage power capacity (1.1 GW provided by pumped hydro at Coo, the rest batteries) and 101 GWh of storage energy capacity (8.2 GWh provided by pumped hydro at Coo, the rest batteries) is included.
+The batteries therefore have a duration of 7.2 hours (which could be considered atypical, given current durations of 1 - 2 hours \addcite{MIT paper for this?}).
+
 The residual load time series, aggregated across all nodes, for this system is plotted in \todo[inline]{Plot this}.

### 2022-06-30 src/GEPPR.jl
Reduce penalty on slack variables

@@ -390,17 +390,19 @@ function absolute_limit_on_nodal_imbalance!(gep::GEPM, opts::Dict)
     )
 
     # Slack variables
-    gep[:M, :variables, :abs_slack_L⁺] =
-        abs_slack_L⁺ = @variable(
+    abs_slack_L⁺ =
+        gep[:M, :variables, :abs_slack_L⁺] = @variable(
             gep.model,
-            [n = N, l = L⁺, y = Y, p = P, i = 1:length(T)],
-            lower_bound = 0
+            [n = N, l = L⁺, y = Y, p = P, t = T],
+            lower_bound = 0,
+            base_name = "abs_slack_L⁺"
         )
-    gep[:M, :variables, :abs_slack_L⁻] =
-        abs_slack_L⁻ = @variable(
+    abs_slack_L⁻ =
+        gep[:M, :variables, :abs_slack_L⁻] = @variable(
             gep.model,
-            [n = N, l = L⁻, y = Y, p = P, i = 1:length(T)],
-            lower_bound = 0
+            [n = N, l = L⁻, y = Y, p = P, t = T],
+            lower_bound = 0,
+            base_name = "abs_slack_L⁻"
         )
 
     gep[:M, :constraints, :MaxAbsNodalImbalance] = @constraint(
@@ -416,7 +418,9 @@ function absolute_limit_on_nodal_imbalance!(gep::GEPM, opts::Dict)
 
     # Overload objective
     obj = gep[:M, :objective]
-    gep[:M, :objective] = @objective(Min, obj + 10^6 * (sum(abs_slack_L⁺) + sum(abs_slack_L⁻)))
+    gep[:M, :objective] = @objective(
+        gep.model, Min, obj + 10^3 * (sum(abs_slack_L⁺) + sum(abs_slack_L⁻))
+    )
 
     return gep
 end

### 2022-06-30 scripts/main_model_runs.jl
Fix what I think is a bug in the line constraints

@@ -18,6 +18,7 @@ end
 
 opts = options_diff_days(sn)[4]
 opts["initial_state_of_charge"] = 0.5
+opts["time_out"] = 100
 opts_vec = [
     merge(
         opts,

### 2022-06-30 src/GEPPR.jl

@@ -394,23 +394,32 @@ function absolute_limit_on_nodal_imbalance!(gep::GEPM, opts::Dict)
         gep[:M, :variables, :abs_slack_L⁺] = @variable(
             gep.model,
             [n = N, l = L⁺, y = Y, p = P, t = T],
-            lower_bound = 0,
             base_name = "abs_slack_L⁺"
         )
     abs_slack_L⁻ =
         gep[:M, :variables, :abs_slack_L⁻] = @variable(
             gep.model,
             [n = N, l = L⁻, y = Y, p = P, t = T],
-            lower_bound = 0,
             base_name = "abs_slack_L⁻"
         )
 
-    gep[:M, :constraints, :MaxAbsNodalImbalance] = @constraint(
+    # Constraints
+    gep[:M, :constraints, :MaxAbsNodalImbalanceUp] = @constraint(
         gep.model,
         [n = N, l = L⁺, y = Y, p = P, i = 1:length(T)],
-        dL⁺[n, l, y, p, T[i]] - abs_slack_L⁺[n, l, y, p, T[i]] <= d_max[n][i]
+        dL⁺[n, l, y, p, T[i]] + abs_slack_L⁺[n, l, y, p, T[i]] <= d_max[n][i]
     )
-    gep[:M, :constraints, :MinAbsNodalImbalance] = @constraint(
+    gep[:M, :constraints, :MinAbsNodalImbalanceUp] = @constraint(
+        gep.model,
+        [n = N, l = L⁺, y = Y, p = P, i = 1:length(T)],
+        dL⁺[n, l, y, p, T[i]] + abs_slack_L⁺[n, l, y, p, T[i]] >= d_min[n][i]
+    )
+    gep[:M, :constraints, :MaxAbsNodalImbalanceDown] = @constraint(
+        gep.model,
+        [n = N, l = L⁻, y = Y, p = P, i = 1:length(T)],
+        dL⁻[n, l, y, p, T[i]] + abs_slack_L⁻[n, l, y, p, T[i]] <= d_max[n][i]
+    )
+    gep[:M, :constraints, :MinAbsNodalImbalanceDown] = @constraint(
         gep.model,
         [n = N, l = L⁻, y = Y, p = P, i = 1:length(T)],
         dL⁻[n, l, y, p, T[i]] + abs_slack_L⁻[n, l, y, p, T[i]] >= d_min[n][i]
[…]

### 2022-06-30 scripts/main_model_runs.jl
Save just in case

@@ -18,7 +18,7 @@ end
 
 opts = options_diff_days(sn)[4]
 opts["initial_state_of_charge"] = 0.5
-opts["time_out"] = 100
+opts["time_out"] = 600
 opts_vec = [
     merge(
         opts,

### 2022-06-30 scripts/nodal_imbalance_investigation.jl

@@ -22,6 +22,7 @@ net_load_forecast_error_dict = scenarios_2_net_load_forecast_error(
     opts, scens, mult
 )
 
+gr()
 n1 = "TIHANGE 1"
 n2 = "TIHANGE 2"
 plt = scatter_nodal_imbalance(net_load_forecast_error_dict, "TIHANGE 1", "TIHANGE 2")
@@ -33,7 +34,7 @@ temp = mean_and_std.(eachcol(df))
 m = first.(temp)
 std = last.(temp)
 node_names = [n for n in names(df)]
-x = 0:2:length(node_names)*2
+x = 0:2:(length(node_names) - 1)*2
 Plots.bar(
     x,
     m;
@@ -49,7 +50,7 @@ Plots.savefig(plotsdir(sn, "mean_imbalance.png"))
 y_min = minimum.(eachcol(df))
 y_max = maximum.(eachcol(df))
 StatsPlots.groupedbar(
-    x[1:end-1],
+    x,
     hcat(y_min, y_max);
     lab="",
     bar_position=:stack,

### 2022-06-30 src/GEPPR.jl
Save just in case

@@ -428,7 +428,7 @@ function absolute_limit_on_nodal_imbalance!(gep::GEPM, opts::Dict)
     # Overload objective
     obj = gep[:M, :objective]
     gep[:M, :objective] = @objective(
-        gep.model, Min, obj + 10^3 * (sum(el^2 for el in abs_slack_L⁺) + sum(el^2 for el in abs_slack_L⁻))
+        gep.model, Min, obj + 1e-3 * (sum(el^2 for el in abs_slack_L⁺) + sum(el^2 for el in abs_slack_L⁻))
     )
 
     return gep

### 2022-06-30 scripts/interaction_selected_days.jl
Add interaction for all 4 days

@@ -1,0 +1,17 @@
+include(joinpath(@__DIR__, "..", "intro.jl"))
+sn = script_name(@__FILE__)
+mkrootdirs.(plotsdir(sn), simsdir(sn))
+
+opts_vec = options_diff_days(sn)
+opts_vec = [
+    merge(
+        opts,
+        Dict(
+            "initial_state_of_charge" => 0.5,
+            "time_out" => 600,
+            "absolute_limit_on_nodal_imbalance" => true,
+        ),
+    ),
+]
+gep = run_GEPPR(opts_vec)
+d_vec = [save_gep_for_security_analysis(gep, opts) for opts in opts_vec]

### 2022-07-01 scripts/process_input.jl
Changing day selection based on #12

@@ -10,6 +10,7 @@ powermodels_2_GEPPR(grid_path, grid_red_path)
 
 # Run a linear operational model for the entire year
 opts = options(
+    "load_multiplier" => 1.5,
     "include_storage" => true,
     "unit_commitment_type" => "none",
     "operating_reserves_sizing_type" => "given",
@@ -18,14 +19,16 @@ opts = options(
     "optimization_horizon" => [1, 8_760],
     "save_path" => datadir("sims", "linear_w_storage"),
 )
-# gep = run_GEPPR(opts)
+gep = run_GEPPR(opts)
 
 # Identify 4 days to investigate - no load shedding at all, almost load shedding and load shedding
-# Because I realised there were mistakes in the model later on, I save this to a "days_for_analysis_new.csv" instead of "days_for_analysis.csv", so that I don't have to recreate all the scenarios.
-# days, TVec = days_to_run_models_on(gep, "days_for_analysis_new.csv")
+# Because I realised there were mistakes in the model later on, I save this to a "days_for_analysis_$(now()).csv" instead of "days_for_analysis.csv", so that I don't have to recreate all the scenarios.
+days, TVec = days_to_run_models_on(gep, "days_for_analysis_$(now()).csv")
 
-# Choose days based on net load, so that I don't have to come back to this again
-days, TVec = days_to_run_models_on(opts, "days_for_analysis_$(now()).csv")
+# Choose days based on net load
+# Idea being if I do this, it's not dependent on any bugs I made in the above
+# however, load shedding depends on network as well / in particular, so the below is kind of useless
+# days, TVec = days_to_run_models_on(opts, "days_for_analysis_$(now()).csv")
 
 opts = options(
     "include_storage" => true,

### 2022-07-01 src/analysis.jl

@@ -92,7 +92,7 @@ function days_to_run_models_on(opts::Dict, filename::String)
     gep = gepm(opts)
     rl = demand_net_of_total_supply(gep)
     rl = sum(collect(values(rl)))
-    rl = sum(reshape(rl, 24, :), dims=1)[:]
+    rl = sum(reshape(rl, N_HR_PER_DAY, :), dims=1)[:]
     srt_idx = sortperm(rl)
     day_least_scarce = srt_idx[1]
     day_little_scarce = srt_idx[Int(round(quantile(1:length(srt_idx), 1/3)))]

### 2022-07-01 scripts/process_input.jl
Allow for diff day CSVs to be used

@@ -52,7 +52,7 @@ end
 ls = sum(
     reshape(
         dropdims(sum(gep[:loadShedding].data; dims=(1, 2, 3)); dims=(1, 2, 3)),
-        24,
+        N_HR_PER_DAY,
         :,
     );
     dims=1,
@@ -69,13 +69,13 @@ storage_dispatch_2_node_injection(
 
 # Check that dispatches make sense
 opts["save_path"] = ""
-opts["optimization_horizon"] = [1, 48]
+opts["optimization_horizon"] = [1, N_HR_PER_DAY*2]
 opts["include_storage"] = true
 gep = run_GEPPR(opts)
-plt_1 = plot_dispatch(gep, 1; T=1:48, N=["RODENHUIZE"])
+plt_1 = plot_dispatch(gep, 1; T=1:N_HR_PER_DAY*2, N=["RODENHUIZE"])
 Plots.savefig(plt_1, datadir("sims", sn, "dispatch_with_storage.pdf"))
 
 opts["include_storage"] = false
 gep = run_GEPPR(opts)
-plt_2 = plot_dispatch(gep, 1; T=1:48, N=["RODENHUIZE"])
+plt_2 = plot_dispatch(gep, 1; T=1:N_HR_PER_DAY*2, N=["RODENHUIZE"])
 Plots.savefig(plt_2, datadir("sims", sn, "dispatch_without_storage.pdf"))

### 2022-07-01 src/GEPPR.jl

@@ -427,8 +427,8 @@ function absolute_limit_on_nodal_imbalance!(gep::GEPM, opts::Dict)
 
     # Overload objective
     obj = gep[:M, :objective]
-    gep[:M, :objective] = @objective(
-        gep.model, Min, obj + 1e-3 * (sum(el^2 for el in abs_slack_L⁺) + sum(el^2 for el in abs_slack_L⁻))
+    gep[:M, :objective_w_slack] = @objective(
+        gep.model, Min, obj + 1e-2 * (sum(el^2 for el in abs_slack_L⁺) + sum(el^2 for el in abs_slack_L⁻))
     )
 
     return gep

### 2022-07-01 src/opts.jl

@@ -38,8 +38,8 @@ function options(pairs...)
     return opts
 end
 
-function options_diff_days(sn)
-    df = CSV.read(datadir("pro", "days_for_analysis.csv"), DataFrame)
+function options_diff_days(sn, file="days_for_analysis.csv")
+    df = CSV.read(datadir("pro", file), DataFrame)
     opts = options(
         "include_storage" => true,
         "operating_reserves_sizing_type" => "given",

### 2022-07-01 scripts/interaction_selected_days.jl
Allow for diff day CSVs to be used

@@ -9,9 +9,20 @@ opts_vec = [
         Dict(
             "initial_state_of_charge" => 0.5,
             "time_out" => 600,
-            "absolute_limit_on_nodal_imbalance" => true,
+            "absolute_limit_on_nodal_imbalance" => false,
+            "save_path" => opts["save_path"] * "_AbsIm=false"
         ),
     ),
+    merge(
+        opts,
+        Dict(
+            "initial_state_of_charge" => 0.5,
+            "time_out" => 600,
+            "absolute_limit_on_nodal_imbalance" => true,
+            "save_path" => opts["save_path"] * "_AbsIm=true"
+        ),
+    )
+    for opts in opts_vec
 ]
 gep = run_GEPPR(opts_vec)
 d_vec = [save_gep_for_security_analysis(gep, opts) for opts in opts_vec]

### 2022-07-01 scripts_debug/debug_UC_edge_effect_because_of_store.jl

@@ -1,0 +1,17 @@
+include(joinpath(@__DIR__, "..", "intro.jl"))
+sn = script_name(@__FILE__)
+mkrootdirs.(plotsdir(sn), simsdir(sn))
+
+opts = options_diff_days(sn, "days_for_analysis_2022_06_21.csv")[4]
+opts["initial_state_of_charge"] = 0.0
+opts["time_out"] = 600
+opts["save_path"] *= "SOC_init=0.0"
+gep = run_GEPPR(opts)
+d = save_gep_for_security_analysis(gep, opts)
+
+opts = options_diff_days(sn, "days_for_analysis_2022_06_21.csv")[4]
+opts["initial_state_of_charge"] = 0.5
+opts["time_out"] = 600
+opts["save_path"] *= "SOC_init=0.5"
+gep = run_GEPPR(opts)
+d = save_gep_for_security_analysis(gep, opts)

### 2022-07-01 src/GEPPR.jl
More options for absimb slacks

@@ -363,8 +363,9 @@ function apply_initial_state_of_charge!(gep::GEPM, opts::Dict)
 end
 
 function absolute_limit_on_nodal_imbalance!(gep::GEPM, opts::Dict)
-    @unpack absolute_limit_on_nodal_imbalance = opts
+    @unpack absolute_limit_on_nodal_imbalance, allow_absolute_imbalance_slacks, absolute_imbalance_slack_penalty = opts
     absolute_limit_on_nodal_imbalance == false && return gep
+    sp = absolute_imbalance_slack_penalty
 
     @info "Applying absolute limits on nodal imbalance..."
     scen_id = month_day(opts)
@@ -403,6 +404,13 @@ function absolute_limit_on_nodal_imbalance!(gep::GEPM, opts::Dict)
             base_name = "abs_slack_L⁻"
         )
 
+    # Fix slacks if necessary
+    if allow_absolute_imbalance_slacks == false
+        JuMP.fix.(abs_slack_L⁺, 0.0; force=true)
+        JuMP.fix.(abs_slack_L⁻, 0.0; force=true)
+    end
+
+
     # Constraints
     gep[:M, :constraints, :MaxAbsNodalImbalanceUp] = @constraint(
         gep.model,
@@ -428,7 +436,7 @@ function absolute_limit_on_nodal_imbalance!(gep::GEPM, opts::Dict)
     # Overload objective
     obj = gep[:M, :objective]
     gep[:M, :objective_w_slack] = @objective(
-        gep.model, Min, obj + 1e-2 * (sum(el^2 for el in abs_slack_L⁺) + sum(el^2 for el in abs_slack_L⁻))
+        gep.model, Min, obj + sp * (sum(el^2 for el in abs_slack_L⁺) + sum(el^2 for el in abs_slack_L⁻))
     )
 
     return gep

### 2022-07-01 src/opts.jl

@@ -31,6 +31,8 @@ function options(pairs...)
         "prevent_simultaneous_charge_and_discharge" => true,
         "rate_a_multiplier" => missing,
         "load_multiplier" => missing,
+        "allow_absolute_imbalance_slacks" => false,
+        "absolute_imbalance_slack_penalty" => 1e3
     )
     for (k, v) in pairs
         opts[k] = v

### 2022-07-01 scripts_debug/debug_abs_imb_infeasible.jl
More options for absimb slacks

@@ -1,0 +1,16 @@
+include(joinpath(@__DIR__, "..", "intro.jl"))
+sn = script_name(@__FILE__)
+mkrootdirs(plotsdir(sn))
+
+opts = options_diff_days(sn, "days_for_analysis_2022_07_01.csv")[4]
+opts["initial_state_of_charge"] = 0.5
+opts["time_out"] = 600
+opts["absolute_limit_on_nodal_imbalance"] = true
+opts["allow_absolute_imbalance_slacks"] = true
+opts["absolute_imbalance_slack_penalty"] = 1e2
+gep_UC = run_GEPPR(opts)
+
+opts["unit_commitment_type"] = "none"
+opts["save_path"] *= "_no_UC"
+gep_ED = run_GEPPR(opts)
+

### 2022-07-01 scripts_debug/debug_abs_imb_infeasible.jl
Include abs imb level redispatch

@@ -5,12 +5,16 @@ mkrootdirs(plotsdir(sn))
 opts = options_diff_days(sn, "days_for_analysis_2022_07_01.csv")[4]
 opts["initial_state_of_charge"] = 0.5
 opts["time_out"] = 600
+opts["upward_reserve_levels_included_in_redispatch"] = [10]
+opts["downward_reserve_levels_included_in_redispatch"] = [10]
 opts["absolute_limit_on_nodal_imbalance"] = true
 opts["allow_absolute_imbalance_slacks"] = true
 opts["absolute_imbalance_slack_penalty"] = 1e2
+rm(opts["save_path"]; force=true, recursive=true)
 gep_UC = run_GEPPR(opts)
 
 opts["unit_commitment_type"] = "none"
 opts["save_path"] *= "_no_UC"
+rm(opts["save_path"]; force=true, recursive=true)
 gep_ED = run_GEPPR(opts)
 

### 2022-07-01 scripts_debug/debug_abs_imb_infeasible.jl
Seems feasible again (???)

@@ -13,6 +13,12 @@ opts["absolute_imbalance_slack_penalty"] = 1e2
 rm(opts["save_path"]; force=true, recursive=true)
 gep_UC = run_GEPPR(opts)
 
+opts = options_diff_days(sn, "days_for_analysis_2022_07_01.csv")[4]
+opts["save_path"] *= "_no_slack"
+opts["allow_absolute_imbalance_slacks"] = false
+rm(opts["save_path"]; force=true, recursive=true)
+gep_UC_no_sl = run_GEPPR(opts)
+
 opts["unit_commitment_type"] = "none"
 opts["save_path"] *= "_no_UC"
 rm(opts["save_path"]; force=true, recursive=true)

### 2022-07-01 scripts/interaction_selected_days.jl
Clean up abs imb debug script

@@ -3,12 +3,14 @@ sn = script_name(@__FILE__)
 mkrootdirs.(plotsdir(sn), simsdir(sn))
 
 opts_vec = options_diff_days(sn)
+opts["time_out"] = 600
+opts["initial_state_of_charge"] = 0.5
+opts["upward_reserve_levels_included_in_redispatch"] = 1:10
+opts["downward_reserve_levels_included_in_redispatch"] = 1:10
 opts_vec = [
     merge(
         opts,
         Dict(
-            "initial_state_of_charge" => 0.5,
-            "time_out" => 600,
             "absolute_limit_on_nodal_imbalance" => false,
             "save_path" => opts["save_path"] * "_AbsIm=false"
         ),
@@ -16,8 +18,6 @@ opts_vec = [
     merge(
         opts,
         Dict(
-            "initial_state_of_charge" => 0.5,
-            "time_out" => 600,
             "absolute_limit_on_nodal_imbalance" => true,
             "save_path" => opts["save_path"] * "_AbsIm=true"
         ),

### 2022-07-01 scripts_debug/debug_abs_imb_infeasible.jl
Clean up abs imb debug script

@@ -2,25 +2,28 @@ include(joinpath(@__DIR__, "..", "intro.jl"))
 sn = script_name(@__FILE__)
 mkrootdirs(plotsdir(sn))
 
-opts = options_diff_days(sn, "days_for_analysis_2022_07_01.csv")[4]
-opts["initial_state_of_charge"] = 0.5
-opts["time_out"] = 600
-opts["upward_reserve_levels_included_in_redispatch"] = [10]
-opts["downward_reserve_levels_included_in_redispatch"] = [10]
-opts["absolute_limit_on_nodal_imbalance"] = true
-opts["allow_absolute_imbalance_slacks"] = true
-opts["absolute_imbalance_slack_penalty"] = 1e2
-rm(opts["save_path"]; force=true, recursive=true)
-gep_UC = run_GEPPR(opts)
+function run_debug_abs_imb_infeasible(L)
+    opts = options_diff_days(sn, "days_for_analysis_2022_07_01.csv")[4]
+    opts["save_path"] *= "_L=$L"
+    opts["initial_state_of_charge"] = 0.5
+    opts["time_out"] = 600
+    opts["upward_reserve_levels_included_in_redispatch"] = L
+    opts["downward_reserve_levels_included_in_redispatch"] = L
+    opts["absolute_limit_on_nodal_imbalance"] = true
+    opts["allow_absolute_imbalance_slacks"] = true
+    opts["absolute_imbalance_slack_penalty"] = 1e2
+    gep_UC = run_GEPPR(opts)
 
-opts = options_diff_days(sn, "days_for_analysis_2022_07_01.csv")[4]
-opts["save_path"] *= "_no_slack"
-opts["allow_absolute_imbalance_slacks"] = false
-rm(opts["save_path"]; force=true, recursive=true)
-gep_UC_no_sl = run_GEPPR(opts)
+    opts = options_diff_days(sn, "days_for_analysis_2022_07_01.csv")[4]
+    opts["save_path"] *= "_no_slack"
+    opts["allow_absolute_imbalance_slacks"] = false
+    gep_UC_no_sl = run_GEPPR(opts)
 
-opts["unit_commitment_type"] = "none"
-opts["save_path"] *= "_no_UC"
-rm(opts["save_path"]; force=true, recursive=true)
-gep_ED = run_GEPPR(opts)
+    opts["unit_commitment_type"] = "none"
[…]

### 2022-07-01 scripts/interaction_selected_days.jl
Increase timeout selected days

@@ -3,7 +3,7 @@ sn = script_name(@__FILE__)
 mkrootdirs.(plotsdir(sn), simsdir(sn))
 
 opts_vec = options_diff_days(sn)
-opts["time_out"] = 600
+opts["time_out"] = 1200
 opts["initial_state_of_charge"] = 0.5
 opts["upward_reserve_levels_included_in_redispatch"] = 1:10
 opts["downward_reserve_levels_included_in_redispatch"] = 1:10

### 2022-07-06 src/GEPPR.jl
Improved way of adapting GEPPR paths

@@ -688,3 +688,38 @@ function change_gep_root_path(
 
     return dictConfig
 end
+
+function change_gep_root_path_full(
+    topdir,
+    from_paths=["/vsc-hard-mounts/leuven-data/331/vsc33168/ASoSEPOC/", ],
+    to_path="/home/u0128861/Desktop/ASoSEPOC/"
+)
+    for (root, dirs, file) in walkdir(topdir)
+        for dir in dirs
+            dirfl = joinpath(root, dir)
+            if isfile(joinpath(dirfl, "config.jld2"))
+                for from_path in from_paths
+                    change_gep_root_path(dirfl, from_path, to_path)
+                end
+            end
+        end
+    end
+end
+
+function change_gep_config_files_full(topdir)
+    for (root, dirs, file) in walkdir(topdir)
+        for dir in dirs
+            dirfl = joinpath(root, dir)
+            opts_file = joinpath(dirfl, "opts.json")
+            config_file = joinpath(dirfl, "config.jld2")
+            if isfile(config_file) && isfile(opts_file)
+                opts = JSON.parsefile(opts_file)
+                config, ~ = param_and_config(opts)
+                @load eval(config_file) dictConfig
+                dc = dictConfig
+                dc["configFile"] = config
+                @save eval(config_file) dictConfig
+            end
+        end
+    end
+end

### 2022-07-06 papers/main.tex
Fix what I think is a bug in the line constraints
Save just in case
More info on the input data

@@
 
+Note that $rinj^{L^+}_{nlt}$, $rinj^{L^-}_{nlt}$, $d^{L^+}_{nlt}$ and $d^{L^-}_{nlt}$ are all free variables and there is a change in sign between upward and downward reserve levels to ensure that the resulting addition to the line flows, $rf^{L^+}_{blt}$ and $rf^{L^-}_{blt}$, are correct.
+
 Since imbalances are aggregated across the network, a particular reserve level activation is not associated with an imbalance at the nodal level.
@@
 The batteries therefore have a duration of 7.2 hours (which could be considered atypical, given current durations of 1 - 2 hours \addcite{MIT paper for this?}).
+
+The network topology was obtained from the Belgian TSO, Elia \cite{Elia2018grid}. It consists of 46 nodes connected by 69 lines.
+
+The load time series used was the load as seen by Elia on the high-voltage grid from the year 2015 \todo{add cite.}.
+For this year, the minimum and maximum load was 5,777 MW and 13,670 MW respectively, with an average of 9,934 MW.
+The electrical load is assigned to the different nodes in the system according to the rated capacity of the transformer feeding each node.
+
+The \ac{VRES} availability factor time series are taken from Elia data for the year 2015, same as the load.
+This was done by dividing generation from offshore wind, wind and solar by the total installed capacities available.
+This allows scaling up the generation of \ac{VRES} to the previously mentioned capacities.
+The distribution of onshore wind and solar photovoltaic (PV) capacity was done according to the distribution made in \cite{DelarueErik2016Dtio}.
 

### 2022-07-06 papers/main.tex
Realised what's happenning with my model

@@
 
+\subsubsection{Methodology used to select days for analysis}
+
+Four days were selected for analysis based on the results of a full year economic dispatch model which included network constraints.
+The load was multiplied uniformly across all nodes by a factor of 1.5 to ensure that some load shedding occurred and then four days were selected which represented the 0, 33, 66 and 100th percentile in terms of daily load shedding.
+The resulting days are presented in \tabref{tab:day_selection}.
+
+\begin{table}[H]
+    \centering
+    \begin{tabular}{cccc}
+        \toprule
+        Day ahead daily load shedding [MWh/day] & Day of year & Month & Day of month \\
+        \midrule
+        0                                       & 161         & 6     & 10           \\
+        449                                     & 319         & 11    & 15           \\
+        1,247                                   & 285         & 10    & 12           \\
+        10,062                                  & 12          & 1     & 12           \\
+        \bottomrule
+    \end{tabular}
+    \caption{Days selected for analysis.
+These were selected based on the results of a full year economic dispatch model with load multiplied uniformly across all nodes by a factor of 1.5.
+The 0, 33, 66 and 100th percentiles in terms of daily load shedding were then selected.}\label{tab:day_selection}
+\end{table}
+
+It should be noted that of the four days selected, only day 285 exhibited load shedding when the load was not scaled but operating reserves were present.
+This finding highlights how day ahead adequacy may occur for reasons other than insufficient capacity aggregated across the network, i.e.
+analysis of the aggregated residual load (the load net of renewable generation) may be uninformative.
+
 \section{Analysis of \acl{DUCPR} model results} \label{sec:initial_results}
 
-This analysis relates only to the 309th day of the year (November 5th). The aim is to investigate what are the causes of load shedding.
+This analysis relates only to the 285th day of the year (12th of October).
+This day was selected since of the 4 days selected  The aim is to investigate what are the causes of load shedding to better understand the system.
 
@@
 \tabref{tab:results_no_OR} lists results for an increasing level of technical constraints or increasingly inflexible system.
-While the objective more than doubles going from the simple linear economic dispatch model with no network constraints to the unit commitment model with network constraints, no load shedding occurs.
-Preventing simultaneous charging and discharging constraints does not affect results at all, implying that additional energy consumption (which is not physically possible) from storage does not aid congestion in this context.
+While the objective increases by \euro 1,267,000 from the simple linear economic dispatch model with no network constraints to the unit commitment model with network constraints, no load shedding occurs.
+Preventing simultaneous charging and discharging constraints actually decreases, though this difference is within the optimality gap (0.01\%).
+This suggests that additional energy consumption (which is not physically possible) from storage does not aid congestion in this context.
 
@@
         \toprule
-        UC  & DANet & PSCD & Load shedding [MWh] & Objective \\
+        UC  & DANet & PSCD & Load shedding [MWh] & Objective [\euro] \\
         \midrule
-            &       &      & 0.0                 & 101,915   \\
-        \xm &       &      & 0.0                 & 109,148   \\
-        \xm & \xm   &      & 0.0                 & 227,495   \\
-        \xm & \xm   & \xm  & 0.0                 & 227,492   \\
+            &       &      & 0.0                 & 3,127,814         \\
+        \xm &       &      & 0.0                 & 4,003,654         \\
+        \xm & \xm   &      & 0.0                 & 4,394,693         \\
+        \xm & \xm   & \xm  & 0.0                 & 4,394,513         \\
         \bottomrule
@@
 
-The dispatch for the model with and without unit commitment constraints and with network constraints is shown in \figref{fig:dispatch_base}.
-The addition of unit commitment constraints leads to storage dispatching and conventional units generating throughout the day.
-Interestingly, the amount of solar generation and storage dispatch is higher in the economic dispatch model and the model with unit commitment and network constraints.
-This may be because batteries and solar generators are located in buses where there is also load, favouring their usage.
+The dispatch aggregated over the entire network for the model with and without unit commitment constraints and with network constraints is shown in \figref{fig:dispatch_base}.
+The addition of unit commitment constraints leads to less storage discharging and more conventional (thermal) units generating at the start of the day.
+The aggregated storage state of charge at 10AM (hour 6826) is approximately the same however.
+The addition of network constraints does not appear to change the aggregated dispatch significantly.
 
@@
         \centering
-        % \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base/dispatch}
+        \includegraphics[width=\textwidth]{../data/sims/main_model_runs/285/base/dispatch}
         \caption{Economic dispatch model}
@@
         \centering
-        % \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_UC=true/dispatch}
+        \includegraphics[width=\textwidth]{../data/sims/main_model_runs/285/base_UC=true/dispatch}
         \caption{Unit commitment model}
@@
         \centering
-        % \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_UC=true_DANet=true/dispatch}
+        \includegraphics[width=\textwidth]{../data/sims/main_model_runs/285/base_UC=true_DANet=true/dispatch}
         \caption{Unit commitment model with day ahead network constraints}
     \end{subfigure}
-    \caption{Dispatch schedules with and without operating reserves.\label{fig:dispatch_base}}
+    \caption{Dispatch schedules for day 285 and increasing model complexity.\label{fig:dispatch_base}}
 \end{figure}
 
-\subsection{Probabilistic operating reserves without reserve activation network constraints}
+\subsection{Probabilistic operating reserves}
 
-\begin{itemize}
-    \item Unit commitment constraints lead to unavoidable load shedding.
-    \item This result should be interpreted with caution however, since reserve provision is quite strict (storage can't provide reserves, non spinning reserves not modelled).
-If more flexibility would be available in terms of reserve provision then this might not occur.
-    \item Network constraints are less problematic than unit commitment (cost increase of \~20,000 euros instead of \~500,000).
-    \item Preventing simultaneous charge and discharge has a small but noticeable effect.
-\end{itemize}
+\tabref{tab:results_no_RANet} summarises results for the model runs with probabilistic operating reserves but no reserve activation network constraints.
+Clearly unit commitment constraints lead to greatly increased load shedding, 3.8 GWh to be precise.
+They also lead to reserve shedding where there wasn't any before.
+Neither day ahead network constraints or preventing simultaneous charging or discharging have an effect on the objective function which is greater than the MIP gap.
 
@@
         \midrule
-            &       &      & 0.0 & 0.0                 & 0.0                    & 101,345   \\
-            &       &      & 0.5 & 0.0                 & 0.0                    & 101,014   \\
-            &       &      & 1.0 & 0.0                 & 0.0                    & 101,014   \\
+            &       &      & 0.0 & 1,575               & 0.0                    & 19,508,913 \\
+            &       &      & 0.5 & 1,575               & 0.0                    & 19,508,913 \\
+            &       &      & 1.0 & 1,575               & 0.0                    & 19,508,913 \\
         \midrule
-        \xm &       &      & 0.0 & 1.669               & 0.0                    & 617,536   \\
-        \xm &       &      & 0.5 & 1.669               & 0.0                    & 617,536   \\
-        \xm &       &      & 1.0 & 1.669               & 0.0                    & 617,536   \\
+        \xm &       &      & 0.0 & 5,410               & 0.0                    & 63,198,973 \\
+        \xm &       &      & 0.5 & 5,410               & 5,683                  & 63,136,103 \\
+        \xm &       &      & 1.0 & 5,410               & 5,616                  & 63,136,444 \\
         \midrule
-        \xm & \xm   &      & 0.0 & 1.669               & 0.0                    & 650,647   \\
-        \xm & \xm   &      & 0.5 & 1.669               & 136                    & 650,231   \\
-        \xm & \xm   &      & 1.0 & 1.669               & 136                    & 650,231   \\
+        \xm & \xm   &      & 0.0 & 5,410               & 0.0                    & 63,198,620 \\
+        \xm & \xm   &      & 0.5 & 5,410               & 5,440                  & 63,137,301 \\
+        \xm & \xm   &      & 1.0 & 5,410               & 5,700                  & 63,140,438 \\
         \midrule
-        \xm & \xm   & \xm  & 0.0 & 1.669               & 0.0                    & 650,647   \\
-        \xm & \xm   & \xm  & 0.5 & 1.669               & 136                    & 650,231   \\
-        \xm & \xm   & \xm  & 1.0 & 1.669               & 136                    & 650,263   \\
+        \xm & \xm   & \xm  & 0.0 & 5,410               & 0.0                    & 63,198,858 \\
+        \xm & \xm   & \xm  & 0.5 & 5,410               & 5,112                  & 63,139,965 \\
+        \xm & \xm   & \xm  & 1.0 & 5,410               & 5,114                  & 63,140,337 \\
         \bottomrule
@@
 
-\figref{fig:dispatch_simple_reserves} shows the dispatch for 3 models with operating reserves, the first a simple economic dispatch, then a unit commitment and then a unit commitment with network constraints.
-Similarly to before, network constraints lead to greater usage of solar and storage compared to the unit commitment model without.
+\figref{fig:dispatch_with_operating_reserves} shows the dispatch schedule for the case with unit commitment constraints, day ahead network constraints and operating reserve requirements but no limit on reserve shedding.
+Load shedding occurs at 8 - 10 AM, even though there is still energy in the storage units.
+This is not due to congestion issues, since some of these storage units are located at the same node as the load being shed.
+Rather, this is likely due to the constraint which specifies that the storage state of charge at the end of the day must be greater than or equal to that at the start of the day.
+This constraint was included so as not to overestimate the flexibility that storage could provide, though it is unlikely that such a situation would occur in reality.
 
@@
     \centering
-    \begin{subfigure}[t]{0.7\textwidth}
-        \centering
-        % \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_RSV=0.0/dispatch}
-        \caption{No UC, no DANet, no PSCD, RSL is 0.0}
-    \end{subfigure}
-    \begin{subfigure}[t]{0.7\textwidth}
-        \centering
-        % \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_UC=true_RSV=0.0/dispatch}
-        \caption{UC, no DANet, no PSCD, RSL is 0.0}
-    \end{subfigure}
-    \begin{subfigure}[t]{0.7\textwidth}
-        \centering
-        % \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_UC=true_DANet=true_RSV=0.0/dispatch}
-        \caption{UC, DANet, no PSCD, RSL is 0.0}
-    \end{subfigure}
-    \caption{Dispatch schedules with and without operating reserves.\label{fig:dispatch_simple_reserves}}
+    \includegraphics[width=0.7\textwidth]{../data/sims/main_model_runs/285/base_UC=true_DANet=true_RSV=1.0/dispatch}
+    \caption{Dispatch schedule for day 285, unit commitment constraints, day ahead network constraints and operating reserve requirements but no limit on reserve shedding.\label{fig:dispatch_with_operating_reserves}}
 \end{figure}
 
-\begin{itemize}
-    \item What's happening to reduce reserve shedding without reducing load shedding? Consider only the difference between $RSL$ 0.5 and 0.0 for the UC and DANet case.
-    \item For $RSL = 0.5 $, load shedding occurs across all nodes for the 10th hour of the day. Reserve shedding occurs on one node, LATOUR, for the 9th hour.
-    \item The commitment for the two values of $RSL$ is not different.
-    \item The load shed in both cases is the same i.e. it is not shed in different places. It is also equal everywhere, which seems strange...
-    \item To avoid reserve shedding, the cost increase appears to be simply from regulating the generation of 6 conventional generators - Eeklo Nord, Meerhout, Monce and the generators at TIHANGE 1 and TIHANGE 3.
-Various renewable generators also modify their output.
-    \item This leads me to believe that economic load shedding is occurring, and that more conventional generators are generating and providing reserves in order to satisfy the reserve shedding limit.
-\end{itemize}
+Considering the case with just unit commitment constraints, the reserve shedding limit can be decreased without increasing load shedding, though there is an increase (~0.09\%) in the objective function greater than the MIP gap.
+This is counterintuitive, as you would expect the DUC-PR model to shed all operating reserves before shedding load in the day ahead stage.
+The former is less costly than the latter after all, due to activation probabilities being < 1.
+However, what is occuring is most likely `economic reserve shedding' - shedding reserves because they are highly unlikely to be activated.
+This phenomenon was also observed in \cite{Hermans2018}.
+\figref{fig:reserves_with_operating_reserves} illustrates that almost all reserve shedding occurs between 14:00 and 16:00 (hour 6830 to 6832) and not 8:00 to 10:00 as with load shedding.
+These are hours with a lot of solar generation and very little conventional, so to avoid this reserve shedding only requires additional generator commitment and not increasing the amount of load shedding in those hours.
 
-The fact that upward operating reserves can be entirely satisfied while load shedding occurs signals to me that reserve provision is easier to provide than satisfying load.
-This is almost certainly due to the weak reserve activation network constraints.
-Further constraining these using the methods described in \secref{sec:further_constraining_operating_reserve_network_activation_constraints} attempts to address this, as will be detailed in the next section.
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.7\textwidth]{../data/sims/main_model_runs/285/base_UC=true_DANet=true_RSV=1.0/reserves}
+    \caption{Dispatch schedule for day 285, unit commitment constraints, day ahead network constraints and operating reserve requirements but no limit on reserve shedding.\label{fig:reserves_with_operating_reserves}}
+\end{figure}
 

### 2022-07-06 scripts/main_model_runs.jl

@@ -260,8 +260,8 @@ function how_is_reserve_shed_without_increasing_load_shedding(gep_vec, opts_vec)
     rs_0_5 = dropdims(
         sum(gep_vec[i + 4][:rsL⁺].data; dims=(2, 3, 4)); dims=(2, 3, 4)
     )
-    @show ls_0_5_idx = findall(ls_0_5 .> 1e-3)
-    @show rs_0_5_idx = findall(rs_0_5 .> 1e-6)
+    ls_0_5_idx = findall(ls_0_5 .> 1e-3)
+    rs_0_5_idx = findall(rs_0_5 .> 1e-6)
 
     # Load shedding occurs same time when RSL = 0 and RSL = 0.5
     ls_0 = dropdims(
@@ -270,8 +270,7 @@ function how_is_reserve_shed_without_increasing_load_shedding(gep_vec, opts_vec)
     ls_0_idx = findall(ls_0_5 .> 1e-3)
     @show ls_0_idx == ls_0_5_idx
 
-    # Plot load shedding
-    @show sum(ls_0[:, 10] .- ls_0_5[:, 10])
+    # Plot load shedding, which occurs at hour 8
     N = GEPPR.get_set_of_nodes(gep_vec[i])
     x = eachindex(N)
     display(
@@ -300,7 +299,7 @@ function how_is_reserve_shed_without_increasing_load_shedding(gep_vec, opts_vec)
     @show GN[q_idx]
 end
 
-# how_is_reserve_shed_without_increasing_load_shedding(gep_vec, opts_vec)
+how_is_reserve_shed_without_increasing_load_shedding(gep_vec, opts_vec)
 
 function load_shedding_distribution(gep_vec, opts_vec, idx)
     pyplot()

### 2022-07-07 src/opts.jl
Investigate effect of cyclic storage SOC

@@ -23,6 +23,7 @@ function options(pairs...)
         "vars_2_save" => Symbol[],
         "exprs_2_save" => Symbol[],
         "initial_state_of_charge" => missing,
+        "cyclic_state_of_charge_constraint" => true,
         "replace_storage_dispatch_with_node_injection" => false,
         "reserve_provision_cost" => 0.0,
         "absolute_limit_on_nodal_imbalance" => false,

### 2022-07-07 scripts/main_model_runs.jl
Investigate effect of cyclic storage SOC

@@ -355,3 +355,13 @@ function load_shedding_distribution(gep_vec, opts_vec, idx)
 end
 
 load_shedding_distribution(gep_vec, opts_vec, 56)
+
+function re_run_but_without_SOC(opts_vec, idx)
+    opts = copy(opts_vec[idx])
+    opts["cyclic_state_of_charge_constraint"] = false
+    opts["save_path"] *= "_wo_cyclic_SOC"
+
+    return main_model_run(opts; make_plots=true)
+end
+
+re_run_but_without_SOC(opts_vec, 15)

### 2022-07-07 src/GEPPR.jl

@@ -10,7 +10,7 @@ function param_and_config(opts::Dict)
     copperplate,
     initial_state_of_charge,
     reserve_provision_cost,
-    operating_reserves_type = opts
+    operating_reserves_type, cyclic_state_of_charge_constraint = opts
 
     is_linear = (opts["unit_commitment_type"] == "none")
     init_soc = opts["initial_state_of_charge"]
@@ -45,7 +45,7 @@ function param_and_config(opts::Dict)
             "end" => [1, 1, optimization_horizon[end]],
         ),
         "imposePeriodicityConstraintOnStorage" =>
-            rolling_horizon ? false : true,
+            rolling_horizon || (cyclic_state_of_charge_constraint == false) ? false : true,
         "dispatchableGeneration" => Dict(),
         "intermittentGeneration" => Dict(),
         "reserveType" => operating_reserves_type,

### 2022-07-07 papers/main.tex
Investigate effect of cyclic storage SOC
Saving

@@
 
-\begin{figure}[H]
+\todo[inline]{Confirm the above suspicion}
+
+\begin{figure}[H]I
     \centering
@@
 
-Including network reserve activation constraints, that is Constraints \ref{eq:reserve_network_activation_constraints_1} through \ref{eq:reserve_network_activation_constraints_1}, led to no discernible change in objective.
-This particular instance is not able to make a trade-off between day ahead adequacy and real time operational security.
-As a first attempt to overcome this issue, the load at all nodes was multiplied by a factor of 1.5.
-This gave the results shown in \tabref{tab:results_LM_1_5}.
+Including network reserve activation constraints, that is Constraints \ref{eq:reserve_network_activation_constraints_1} through \ref{eq:reserve_network_activation_constraints_2}, led to no discernible changes in objective value within the 0.01\% MIP gap, as can be seen from \tabref{tab:results_with_RANet} (see the difference between first three and second three rows).
+Similarly including absolute limits on the nodal imbalance realisations has very little effect on the results.
 
@@
     \footnotesize
-    \begin{tabular}{cccccccc}
+    \begin{tabular}{ccccccccc}
         \toprule
-        UC  & DANet & PSCD & AbsImb & RSL & Load shedding [MWh] & Reserve Shedding [MWh] & Objective \\
+        UC  & DANet & PSCD & RANet & AbsImb & RSL & Load shedding [MWh] & Reserve Shedding [MWh] & Objective [\euro] (MIP Gap [\%]) \\
         \midrule
-            &       &      &        & 0.0 & 32,366              & 0                      & 7,290,909 \\
-            &       &      &        & 0.5 & 0.0                 & 36,269                 & 447,102   \\
-            &       &      &        & 1.0 & 0.0                 & 36,082                 & 440,467   \\
+        \xm & \xm   & \xm  &       &        & 0.0 & 5,410               & 0.0                    & 63,198,858                       \\
+        \xm & \xm   & \xm  &       &        & 0.5 & 5,410               & 5,112                  & 63,139,965                       \\
+        \xm & \xm   & \xm  &       &        & 1.0 & 5,410               & 5,114                  & 63,140,337                       \\
         \midrule
-        \xm &       &      &        & 0.0 & 32,366              & 0                      & 7,433,785 \\
-        \xm &       &      &        & 0.5 & 1,669               & 34,379                 & 924,686   \\
-        \xm &       &      &        & 1.0 & 1,669               & 34,308                 & 919,363   \\
+        \xm & \xm   & \xm  & \xm   &        & 0.0 & 5,410               & 0                      & 63,199,116                       \\
+        \xm & \xm   & \xm  & \xm   &        & 0.5 & 5,410               & 4,898                  & 63,139,874                       \\
+        \xm & \xm   & \xm  & \xm   &        & 1.0 & 5,410               & 5,611                  & 63,137,767                       \\
         \midrule
-        \xm & \xm   &      &        & 0.0 & 33,929              & 0                      & 7,822,959 \\
-        \xm & \xm   &      &        & 0.5 & 1,967               & 39,234                 & 1,249,422 \\
-        \xm & \xm   &      &        & 1.0 & 1,941               & 39,160                 & 1,245,647 \\
-        \midrule
-        \xm & \xm   & \xm  &        & 0.0 & 33,929              & 0                      & 7,822,959 \\
-        \xm & \xm   & \xm  &        & 0.5 & 1,967               & 39,234                 & 1,249,422 \\
-        \xm & \xm   & \xm  &        & 1.0 & 1,941               & 39,160                 & 1,245,647 \\
-        \midrule
-        \xm & \xm   & \xm  & \xm    & 0.0 & 33,929              & 0                      & 7,822,959 \\
-        \xm & \xm   & \xm  & \xm    & 0.5 & 1,967               & 39,234                 & 1,249,422 \\
-        \xm & \xm   & \xm  & \xm    & 1.0 & 1,941               & 39,160                 & 1,245,647 \\
+        \xm & \xm   & \xm  & \xm   & \xm    & 0.0 & 5,410               & 0                      & 63,232,938 (0.04)
+        )                                                                                                                           \\
+        \xm & \xm   & \xm  & \xm   & \xm    & 0.5 & 5,410               & 5,687                  & 63,150,183                       \\
+        \xm & \xm   & \xm  & \xm   & \xm    & 1.0 & 5,410               & 6,452                  & 63,155,353 (0.02)                \\
         \bottomrule
     \end{tabular}
-    \caption{Analysis of model results with operating reserves and with reserve activation network constraints with load multiplied by a factor of 1.5.
-UC = Unit Commitment constraints, DANet = Day Ahead Network constraints, PSCD = Prevent Simultaneous Charging and Discharging constraints, RSL = Reserve Shedding Limit.}\label{tab:results_no_RANet}
+    \caption{Analysis of model results with operating reserves, with and without reserve activation network constraints and with and without absolute limits on nodal imbalances.
+MIP gap is shown in the case where this was greater than 0.01\%.
+UC = Unit Commitment constraints, DANet = Day Ahead Network constraints, PSCD = Prevent Simultaneous Charging and Discharging constraints, RANet = Reserve Activation Network constraints, AbsImb = Absolute limits on nodal Imbalance realisation, RSL = Reserve Shedding Limit.}\label{tab:results_with_RANet}
 \end{table}
 
-\todo[inline]{Objective is weird for some cases...}
+Multiplying the load by a factor of 1.5 gives the results shown in \tabref{tab:results_with_RANet_LM_15}.
+Here it is possible to discern a trade-off between load shedding and reserve shedding, albeit a small one - reducing reserve shedding to 0 from approximately 12,000 MWh increases load shedding only by 914 MWh.
+That this effect is slight is for similar reasons to that listed previously - reserve shedding occurs in different hours to load shedding and load shedding occurs in the same hours as before (8:00 to 11:00) and is likely due to the cyclic constraint on storage.
 
-With reference to \tabref{tab:results_LM_1_5}:
-
-\begin{itemize}
-    \item Even in this case, the reserve activation network constraints are not binding, since the AbsImb constraints do not alter the objective function.
-    \item On the other hand, the trade-off between reserve shedding and load shedding can be seen now in all cases. This trade-off is also obvious when observing the objective function.
-    \item Unit commitment constraints do relatively little to increase total costs, but this is unsurprising given that costs are dominated by load shedding.
-    \item Day ahead network constraints increase load shedding considerably i.e. congestion also plays a role in increasing load shedding.
-\end{itemize}
-
-Closer investigation at load shedding patterns for the model with UC and DANet (but no PSCD and AbsImb) and $RSL = 0.0$ reveals that load shedding is at it's highest in the mornings (8 - 10 AM) and in the evenings (17 - 20).
-No load shedding occurs in the middle of the day, which happens to be when load is lowest and solar generation is greatest.
-Load shedding is roughly equally distributed throughout the day in absolute though not relative terms.
-
-\begin{figure}[H]
+\begin{table}[ht]
     \centering
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        % \includegraphics[width=\textwidth]{../plots/main_model_runs/ls_timeseries_idx=56}
-        \caption{}
-    \end{subfigure}
-    \newline
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        % \includegraphics[width=\textwidth]{../plots/main_model_runs/ls_per_node_abs_idx=56}
-        \caption{}
-    \end{subfigure}
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        % \includegraphics[width=\textwidth]{../plots/main_model_runs/ls_per_node_idx=56}
-        \caption{}
-    \end{subfigure}
-    \caption{Load shedding distribution in time (top figure, aggregated over all nodes) and space (aggregated over the entire day) for UC, DANet, LM = 1.5 and $RSL = 0.5$.}
-\end{figure}
+    \footnotesize
+    \begin{tabular}{ccccccccc}
+        \toprule
+        UC  & DANet & PSCD & RANet & AbsImb & RSL & Load shedding [MWh] & Reserve Shedding [MWh] & Objective [\euro] (MIP Gap [\%]) \\
+        \midrule
+            &       &      & \xm   &        & 0.0 & 5,825               & 0.0                    & 69,902,500                       \\
+        \xm &       &      & \xm   &        & 0.0 & 5,825               & 0.0                    & 77,491,233                       \\
+        \xm & \xm   &      & \xm   &        & 0.0 & 6,314               & 0.0                    & 82,523,020                       \\
+        \xm & \xm   & \xm  & \xm   &        & 0.0 & 6,314               & 0.0                    & 82,523,020                       \\
+        \midrule
+        \xm & \xm   & \xm  & \xm   &        & 0.0 & 6,314               & 0.0                    & 82,523,020                       \\
+        \xm & \xm   & \xm  & \xm   &        & 0.5 & 5,410               & 12,187                 & 72,771,511                       \\
+        \xm & \xm   & \xm  & \xm   &        & 1.0 & 5,410               & 12,244                 & 72,771,897                       \\
+        \midrule
+        \xm & \xm   & \xm  & \xm   & \xm    & 0.0 & 6,314               & 0                      & 83,163,868                       \\
+        \xm & \xm   & \xm  & \xm   & \xm    & 0.5 & 5,410               & 13,115                 & 73,298,008                       \\
+        \xm & \xm   & \xm  & \xm   & \xm    & 1.0 & 5,410               & 13,430                 & 73,296,805                       \\
+        \bottomrule
+    \end{tabular}
+    \caption{Analysis of model results with operating reserves, with and without reserve activation network constraints and with and without absolute limits on nodal imbalances and for load multiplied by a factor of 1.5.
+MIP gap is shown in the case where this was greater than 0.01\%.
+UC = Unit Commitment constraints, DANet = Day Ahead Network constraints, PSCD = Prevent Simultaneous Charging and Discharging constraints, RANet = Reserve Activation Network constraints, AbsImb = Absolute limits on nodal Imbalance realisation, RSL = Reserve Shedding Limit.}\label{tab:results_with_RANet_LM_15}
+\end{table}
 

### 2022-07-07 scripts/main_model_runs.jl

@@ -115,8 +115,9 @@ function main_model_run(opts; make_plots=false)
 
     dis_file = joinpath(opts["save_path"], "dispatch.png")
     res_file = joinpath(opts["save_path"], "ls_and_rs.png")
+    comm_file = joinpath(opts["save_path"], "commitment.png")
 
-    if make_plots && (isfile(dis_file) == false && isfile(res_file) == false)
+    if make_plots && (any(isfile.([dis_file,res_file, comm_file]) .== false))
         # Plot dispatch
         plt = plot_dispatch(gep, 1)
         Plots.savefig(plt, dis_file)
@@ -125,13 +126,13 @@ function main_model_run(opts; make_plots=false)
         if opts["operating_reserves_type"] != "none"
             has_OR(gep) == false && apply_operating_reserves!(gep, opts)
             plt = plot_reserves_simple(gep, 1)
-            Plots.savefig(plt, joinpath(opts["save_path"], "reserves.png"))
+            Plots.savefig(plt, res_file)
         end
 
         # If has commitment variables, also plot the commitment heatmap
         if opts["unit_commitment_type"] != "none"
-            plt = plot_reserves_simple(gep, 1)
-            Plots.savefig(plt, joinpath(opts["save_path"], "commitment.png"))
+            plt = plot_commitment(gep, 1)
+            Plots.savefig(plt, comm_file)
         end
 
         # Plot the reserve shedding and load shedding time series
@@ -148,7 +149,7 @@ function main_model_run(opts; make_plots=false)
             [ls rs];
             lab=["Load shedding" "Reserve shedding"],
             ylabel="Power",
-            ylims=(0, max(100, maximum(ls))),
+            ylims=(0, max(100, maximum(ls), maximum(rs))),
         )
         Plots.savefig(plt, res_file)
     end

### 2022-07-08 src/GEPPR.jl
Fix bug for changing GEPPR paths

@@ -714,6 +714,7 @@ function change_gep_config_files_full(topdir)
             config_file = joinpath(dirfl, "config.jld2")
             if isfile(config_file) && isfile(opts_file)
                 opts = JSON.parsefile(opts_file)
+                opts = options(collect(opts)...)
                 config, ~ = param_and_config(opts)
                 @load eval(config_file) dictConfig
                 dc = dictConfig

### 2022-09-07 papers/main.tex
Update model description

@@
 In doing so, we deviate from typical practices in a number of ways.
-First, the adequacy assessment is done for a limited number of days instead of many Monte Carlo years for the sake of clarity and brevity.
+First, the adequacy assessment was done for a limited number of days instead of many Monte Carlo years for the sake of clarity and brevity.
 Indeed, doing this exercise for many Monte Carlo years would drastically increase the computational complexity while the aim of this exercise is to illustrate how such an interaction could be performed.
@@
     \item $GR$ - Renewable / variable generators (\ac{VRES}).
+    \item $S$ - Storage technologies
     \item $N$ - Nodes or buses in the network
@@
     \item $P^{max}_g$ - Maximum power output
+    \item $MUT_g$ - Minimum up time
+    \item $MDT_g$ - Minimum down time
+    \item $RAMP_g$ - Maximum ramping limit (up and down)
     \item $D^{L^+}_{lt}$ - Upward reserve requirement
     \item $D^{L^-}_{lt}$ - Downward reserve requirement
+    \item $P^{L^+}_{lt}$ - Upward reserve level activation probability
+    \item $P^{L^-}_{lt}$ - Downward reserve level activation probability
 \end{itemize}
@@
 \begin{itemize}
+    \item $z_{gt}$ - Commitment
+    \item $v_{gt}$ - start up
+    \item $w_{gt}$ - Shut down
     \item $q_{gt}$ - Generation
     \item $\hat{q}_{gt}$ - Generation above the minimum stable operating point (0 in the case of renewables).
+    \item $d_{st}$ - Storage discharge
+    \item $c_{st}$ - Storage charge
+    \item $y_{st}$ - Binary variable, 1 if storage discharges and 0 if it charges
+    \item $e_{st}$ - Storage state of charge variable
     \item $ls_{nt}$ - Load shedding
@@
 
+Objective is to minimise the sum of all day ahead and real time expected operational costs, the terms of which are defined in the following sections.
+
+\subsubsection{Commitment, start up and shut down costs}
+
+These costs are only defined for dispatchable generators $g \in GD$.
+
+\begin{equation}
+    \sum_{t \in T} \sum_{g \in GD} \left ( C^{comm} \cdot z_{gt} + C^{start} \cdot v_{gt} + C^{shut} \cdot w_{gt} \right ) \\
+\end{equation}
+
+\subsubsection{Day ahead generation costs}
+
+These costs are defined for all generators $g \in G$.
+
+\begin{equation}
+    \sum_{t \in T} \sum_{g \in G} C^{var}_{g} \cdot \hat{q}_{gt}
+\end{equation}
+
+\subsubsection{Probabilistic reserve activation costs}
+
+These are defined for all generators $g \in G$.
+
 \begin{align}
-    \min \quad & \sum_{t \in T} \sum_{g \in G} C^{var}_{g} \cdot \hat{q}_{gt} \nonumber                                                                  \\
-               & + \sum_{t \in T} \sum_{l \in L^+} \sum_{n \in N} P^{L^+} \cdot \sum_{g \in G} C^{var}_{g} \cdot r^{L^+}_{glt} + C^{shed} \cdot rs_{nlt} \\
-               & - \sum_{t \in T} \sum_{l \in L^-} \sum_{n \in N} P^{L^-} \cdot C^{var}_{g} \cdot r^{L^-}_{gnlt} + C^{shed} \cdot rc_{nlt} \nonumber
+     & + \sum_{g \in G, \; t \in T} \sum_{l \in L^+} \sum_{n \in N} \left ( P^{L^+}_{lt} \cdot \sum_{g \in G} C^{var}_{g} \cdot r^{L^+}_{glt} + C^{shed} \cdot rs_{nlt} \right ) \\
+     & - \sum_{g \in G, \; t \in T} \sum_{l \in L^-} \sum_{n \in N} \left ( P^{L^-}_{lt} \cdot C^{var}_{g} \cdot r^{L^-}_{gnlt} + C^{shed} \cdot rc_{nlt} \right ) \nonumber
 \end{align}
 
-From the top line to the bottom, the costs are those of dispatching generators and activating upwards or downwards reserves.
+These terms attempt to take into account the expected cost of activating operating reserves.
+The expectation is obtained by weighting the cost of activating a particular upward reserve level $l \in L^+$ (increase in costs) or downward reserve level $l \in L^-$ (decrease in costs) by the probability of that reserve level being activated, $P^{L^+}_{lt}$ and $P^{L^-}_{lt}$ respectively.
+
+Both upward and downward reserve levels can be `shed', i.e.
+not provided.
+In the case of upward reserves, load shedding in real time is expected so as to satisfy the power balance.
+In the case of downward reserves, load shedding in the day ahead stage provides the additional downward reserves.
+The assumption (perhaps implausible) is that this load could in fact be served i.e.
+not shed depending on the imbalance realisation.
+Since load shedding in day ahead is the most expensive decision that the model can take, downward reserve provided in this way is a last resort, even though this term of the objective on its own leads to a decrease in objective costs.
 
@@
 
-\subsubsection{Day ahead constraints}
+\subsubsection{Day ahead power balance}
 
@@
 \begin{equation}
-    \sum_{g \in G_n} q_{gt} + ls_{nt} = D_{nt} + inj_{nt} \quad n \in N, \; t \in T
+    \sum_{g \in G_n} q_{gt} + \sum_{s \in S_n} (d_{st} - c_{st}) + ls_{nt} = D_{nt} + inj_{nt} \quad n \in N, \; t \in T
 \end{equation}
 
-Note the use of the set $G_n$ to only allow generators at node $n$ to contribute to the power balance. Another way of describing this would have been through an incidence matrix.
+Note the use of the set $G_n$ to only allow generators at node $n$ to contribute to the power balance (similarly for storage technologies).
+Another way of describing this would have been through an incidence matrix.
 
-Network constraints:
+The equation below limits load shedding to the demand at that node.
+In copper plate systems this constraint is not needed, however in nodal systems this constraint is required as unphysical load shedding may be exploited to improve congestion (as was the case during this exercise).
+
+\begin{equation}
+    ls_{nt} \leq D_{nt} \quad n \in N, \; t \in T
+\end{equation}
+
+\subsubsection{Day ahead network constraints}
+
+The DC power flow approximation used in this exercise is described below:
 
@@
 
-Constraints on generator output:
+\subsubsection{Generators}
+
+Constraints on generation:
 
@@
 
-For brevity and clarity, constraints on ramping and minimum up and down times are omitted.
+Minimum up and down times:
+
+\begin{align}
+    1 - z_{gt} \geq \sum_{\tau \in 1:MDT_g - 1} w_{gt-\tau} & \quad g \in GD, \; t \in T_{MDT:end} \\
+    z_{gt} \geq \sum_{\tau \in 1:MUT_g - 1} v_{gt-\tau}     & \quad g \in GD, \; t \in T_{MUT:end} \\
+\end{align}
+
+Relationship between commitment, startup and shutdown:
+
+\begin{align}
+    z_{gt-1} = z_{gt} - v_{gt} + w_{gt} & \quad g \in GD, \; t \in T_{2:end} \\
+    w_{gt} + v_{gt} \leq 1              & \quad g \in GD, \; t \in T
+\end{align}
+
+Ramping constraints:
+
+\begin{align}
+    q_{gt} - q_{gt-1} + r^+_{gt} \leq RAMP_g & \quad g \in GD, \; t \in T_{2:end} \\
+    q_{gt-1} - q_{gt} + r^-_{gt} \leq RAMP_g & \quad g \in GD, \; t \in T_{2:end} \\
+\end{align}
+
+\subsubsection{Storage}
+
+\begin{align}
+    e_{t+1} = e_t + \sqrt{\eta} \cdot c_{st} - \sqrt{\eta} \cdot d_{st} & \quad s \in S, \; t \in T_{1:end} \\
+    e_{end} \geq e_0                                                    & \quad s \in S                     \\
+    y_{st} \cdot d_{st} \geq 0                                          & \quad s \in S, \; t \in T         \\
+    y_{st} \cdot d_{st} \leq K_s                                        & \quad s \in S, \; t \in T         \\
+    (1 - y_{st}) \cdot c_{st} \geq 0                                    & \quad s \in S, \; t \in T         \\
+    (1 - y_{st}) \cdot c_{st} \leq K_s                                  & \quad s \in S, \; t \in T
+\end{align}
+
+The binary variable $y_{st}$ prevents simultaneous charging and discharging.
+While such unphysical behaviour is not typical, in this case the constraint was deemed necessary since such behaviour could be exploited to reduce congestion.
+This was also investigated in the results.
 
@@
 
-At the time of writing, this convex hull restriction led to model infeasibilities and so it is not
+At the time of writing, this convex hull restriction led to model infeasibilities and so it was not implemented.
 
@@
 
+\clearpage
+
+\bibliography{bibliography}
+
 \end{document}

### 2022-09-07 papers/main.tex
Continued writing the model

@@
     -F_b \leq f_{bt} \leq F_b                        & \quad b \in B, \; t \in T \\
-    \sum_{n \in N} inj_{nt} = 0                      & \quad n \in N, \; t \in T \\
+    \sum_{n \in N} inj_{nt} = 0                      & \quad n \in N, \; t \in T
 \end{align}
@@
     1 - z_{gt} \geq \sum_{\tau \in 1:MDT_g - 1} w_{gt-\tau} & \quad g \in GD, \; t \in T_{MDT:end} \\
-    z_{gt} \geq \sum_{\tau \in 1:MUT_g - 1} v_{gt-\tau}     & \quad g \in GD, \; t \in T_{MUT:end} \\
+    z_{gt} \geq \sum_{\tau \in 1:MUT_g - 1} v_{gt-\tau}     & \quad g \in GD, \; t \in T_{MUT:end}
 \end{align}
@@
     q_{gt} - q_{gt-1} + r^+_{gt} \leq RAMP_g & \quad g \in GD, \; t \in T_{2:end} \\
-    q_{gt-1} - q_{gt} + r^-_{gt} \leq RAMP_g & \quad g \in GD, \; t \in T_{2:end} \\
+    q_{gt-1} - q_{gt} + r^-_{gt} \leq RAMP_g & \quad g \in GD, \; t \in T_{2:end}
 \end{align}
@@
     e_{t+1} = e_t + \sqrt{\eta} \cdot c_{st} - \sqrt{\eta} \cdot d_{st} & \quad s \in S, \; t \in T_{1:end}                   \\
-    e_{end} \geq e_0                                                    & \quad s \in S                     \\
+    e_{end} \geq e_0                                                    & \quad s \in S   \label{eq:storage_cyclic_constrant} \\
     y_{st} \cdot d_{st} \geq 0                                          & \quad s \in S, \; t \in T                           \\
@@
 This was also investigated in the results.
+
+When carrying out this exercise it was noticed that the system being studied was very sensitive to assumptions on storage due to the high penetration of batteries (~92 GWh).
+Assuming that all storage technologies were empty at the beginning of the day led to load shedding at this time because the system relies on charging storage during the day.
+Assuming they were full led to no scarcity issues at all.
+Assuming they were half full sometimes led to a single spike in load shedding \emph{despite there being sufficient energy in the storage units to avoid this}.
+This is because of the cyclic constraint on the state of charge given by \eqref{eq:storage_cyclic_constrant}, that is that the load shedding was occurring so as to satisfy this constraint and not because there was any real scarcity.
 
@@
 \begin{equation}
-    \sum_{n \in N, l \in L^+} rs_{nlt} \leq RSL \cdot \sum_{l \in L^+} D^{L^+}_{l't} \quad t \in T
+    \sum_{n \in N, l \in L^+} rs_{nlt} \leq RSL \cdot \sum_{l \in L^+} D^{L^+}_{lt} \quad t \in T
 \end{equation}
@@
 Since imbalances are aggregated across the network, a particular reserve level activation is not associated with an imbalance at the nodal level.
-The above constraints therefore enforce that for each reserve level $l$ and node $n$, there exists some combination of nodal imbalance, node injections, generator dispatches and line flows which would satisfy the network constraints AND the imbalance across the entire network.
+The above constraints therefore enforce that for each reserve level $l$ and node $n$, \emph{there exists some combination} of nodal imbalance, node injections, generator dispatches and line flows which would satisfy the network constraints AND the imbalance across the entire network.
+This weakly constrains the reserve activation network flows since the optimizer is free to choose this combination.
 

### 2022-11-22 papers/main.tex
Saving

@@
 \maketitle
+
+\begin{abstract}
+    To be filled in
+\end{abstract}
 
@@
 
-Secondly, the adequacy assessment is done using an unconventional \ac{UC} model, a modified version of the \ac{DUCPR} model described in \cite{Bruninx2017}.
+Secondly, the adequacy assessment is done using an unconventional \ac{UC} model, a modified version of the \ac{DUCPR} model described in~\cite{Bruninx2017}.
 This model can be seen as a compromise between a computationally expensive stochastic \ac{UC} model and a more tractable though less accurate deterministic \ac{UC} model with operating reserve requirements.
 The treatment of operating reserves in this model allows a trade-off between shedding load in the day ahead stage, which would reduce day ahead adequacy, and shedding operating reserves, which should reduce operational security.
-By making this trade-off explicit and verifying the results in an operational security analysis we aim to elaborate on the results of \cite{Hermans2018}, which did not consider the possibility that shedding operating reserves would lead to insecure operation of the electric power system.
+By making this trade-off explicit and verifying the results in an operational security analysis we aim to elaborate on the results of~\cite{Hermans2018}, which did not consider the possibility that shedding operating reserves would lead to insecure operation of the electric power system.
 
@@
     \centering
-    % \includegraphics[width=0.9\textwidth]{../plots/nodal_imbalance_investigation/imbalance_range}
+    \includegraphics[width=0.9\textwidth]{../plots/nodal_imbalance_investigation/imbalance_range}
     \caption{Range of possible imbalances for all hours of day 309.\label{fig:imbalance_range}}
@@
     \centering
-    % \includegraphics[width=0.9\textwidth]{../plots/nodal_imbalance_investigation/convex_hull_TIHANGE 1_TIHANGE 2}
+    \includegraphics[width=0.9\textwidth]{../plots/nodal_imbalance_investigation/convex_hull_TIHANGE 1_TIHANGE 2}
     \caption{Illustration of convex hull constraints on $d_{nlt}^{L^+}$ and $d_{nlt}^{L^-}$.
@@
 
-The grid and resource data used for this interaction is inspired by the case study in \cite{Belderbos2020}, omitting the gas network and power to gas technologies.
+The grid and resource data used for this interaction is inspired by the case study in~\cite{Belderbos2020}, omitting the gas network and power to gas technologies.
 This data gives a stylised Belgian system with a high penetration (~8s0\%) of \ac{VRES}, which includes 75.7 GW of solar PV, 7.0 GW of Onshore Wind and 7.3 GW of Offshore Wind.
@@
 Power to Gas is omitted but 14.1 GW of storage power capacity (1.1 GW provided by pumped hydro at Coo, the rest batteries) and 101 GWh of storage energy capacity (8.2 GWh provided by pumped hydro at Coo, the rest batteries) is included.
-The batteries therefore have a duration of 7.2 hours (which could be considered atypical, given current durations of 1 - 2 hours \addcite{MIT paper for this?}).
+The batteries therefore have a duration of 7.2 hours (which could be considered atypical, given current durations of 1 - 2 hours~\cite{MITEI2022}).
 
-The network topology was obtained from the Belgian TSO, Elia \cite{Elia2018grid}. It consists of 46 nodes connected by 69 lines.
+The network topology was obtained from the Belgian TSO, Elia~\cite{Elia2018grid}. It consists of 46 nodes connected by 69 lines.
 
-The load time series used was the load as seen by Elia on the high-voltage grid from the year 2015 \todo{add cite.}.
+The load time series used was the load as seen by Elia on the high-voltage grid from the year 2015~\cite{Belderbos2020}.
 For this year, the minimum and maximum load was 5,777 MW and 13,670 MW respectively, with an average of 9,934 MW.
@@
 This allows scaling up the generation of \ac{VRES} to the previously mentioned capacities.
-The distribution of onshore wind and solar photovoltaic (PV) capacity was done according to the distribution made in \cite{DelarueErik2016Dtio}.
+The distribution of onshore wind and solar photovoltaic (PV) capacity was done according to the distribution made in~\cite{DelarueErik2016Dtio}.
 
@@
 
-A complete overview of the data is given in \cite{Belderbos2020}, Appendix B.
-
-\subsubsection{Methodology used to select days for analysis}
+A complete overview of the data is given in~\cite{Belderbos2020}, Appendix B.
 
@@
 
-\begin{table}[H]
+\begin{table}[ht]
     \centering
@@
 
-\todo[inline]{Confirm the above suspicion}
-
-\begin{figure}[H]I
+\begin{figure}[H]
     \centering
@@
 However, what is occuring is most likely `economic reserve shedding' - shedding reserves because they are highly unlikely to be activated.
-This phenomenon was also observed in \cite{Hermans2018}.
+This phenomenon was also observed in~\cite{Hermans2018}.
 \figref{fig:reserves_with_operating_reserves} illustrates that almost all reserve shedding occurs between 14:00 and 16:00 (hour 6830 to 6832) and not 8:00 to 10:00 as with load shedding.
@@
     \centering
-    \footnotesize
+    \tiny
     \begin{tabular}{ccccccccc}
@@
 
+\subsection{Simple illustratve case}
+
+We begin by illustrating the tradeoff that exists between real time load shedding and day ahead reserve shedding with a simple model.
+No network or unit commitment constraints are included and no storage either.
+The latter choice is made since in this model formulation storage is not able to contribute to operating reserve requirements, complicating the interpretation of the results.
+Day 319 is chosen since resource use is not as `tight' as the other days and the absence of storage increases the probability of load or reserve shedding.
+
+A comparison of the real time operational security from the DUC-PR model compared to those obtained from a power flow performed by the team ULiege is given in the figure below.
+
 \clearpage

### 2022-11-22 scripts/copperplate_adequacy_security_tradeoff.jl

@@ -1,0 +1,111 @@
+include(joinpath(@__DIR__, "..", "intro.jl"))
+sn = script_name(@__FILE__)
+mkrootdirs(plotsdir(sn))
+
+opts = options_diff_days(sn, "days_for_analysis.csv")[2]
+opts = merge(
+    opts,
+    Dict(
+        "copperplate" => true,
+        "unit_commitment" => false,
+        # "downward_reserve_levels_included_in_redispatch" => 1:10,
+        # "upward_reserve_levels_included_in_redispatch" => 1:10,
+        "include_storage" => false,
+        "vars_2_save" => [:z, :q, :ls, :rL⁺, :rL⁻, :rsL⁺, :rsL⁻],
+    ),
+)
+opts_vec = [
+    merge(
+        opts,
+        Dict(
+            "reserve_shedding_limit" => rsl,
+            "save_path" => opts["save_path"] * "_RSL=$rsl",
+        ),
+    ) for rsl in 1.0:-0.5:0.0
+]
+gep_vec = run_GEPPR(opts_vec)
+rsl_vec = [opts_vec[i]["reserve_shedding_limit"] for i in 1:length(opts_vec)]
+
+function plot_DUCPR_reserve_shedding_sensitivity(
+    gep_vec, opts_vec, rsl_vec, days=["161"]
+)
+    ls = Dict(
+        rsl_vec[i] => try
+            sum(gep_vec[i][:loadShedding].data; dims=(1, 2, 3))[:]
+        catch
+            GEPPR.SVC(NaN)
+        end for i in 1:length(opts_vec)
+    )
+    rsLt = Dict(rsl_vec[i] => try
+        gep_vec[i][:rsL⁺]
+    catch
+        GEPPR.SVC(NaN)
+    end for i in 1:length(opts_vec))
+    rsL⁺ = Dict(
+        rsl_vec[i] => try
+            sum(
+                rsLt[(rsl_vec[i])][n, l, 1, 1, t]
+                # rsLt[(rsl_vec[i])][n, l, 1, 1, t] *
+                # P⁺[(rsl_vec[i])][l, 1, 1, t] 
+                for n in GEPPR.get_set_of_nodes(gep_vec[i]),
+                l in GEPPR.get_set_of_upward_reserve_levels(gep_vec[i]),
+                t in GEPPR.get_set_of_time_indices(gep_vec[i])[3]
+            )
+        catch
+            NaN
+        end for i in 1:length(opts_vec)
+    )
+    nd = length(days)
[…]

### 2022-11-22 scripts/interaction_selected_days.jl

@@ -36,7 +36,7 @@ opts_vec = vcat(
     ]...,
 )
 gep_vec = run_GEPPR(opts_vec)
-d_vec = [save_gep_for_security_analysis(gep, opts) for opts in opts_vec]
+d_vec = [save_gep_for_security_analysis(gep, opts) for (gep,opts) in zip(gep_vec,opts_vec)]
 
 # Some short analysis
 function analyse_interaction_selected_days()

### 2022-11-22 papers/main.tex
Update, copperplate sort of done

@@
 
-This analysis relates only to the 285th day of the year (12th of October).
-This day was selected since of the 4 days selected  The aim is to investigate what are the causes of load shedding to better understand the system.
+This analysis relates only to the 285th day of the year (12th of October). The aim is to investigate what are the causes of load shedding to better understand the system.
 
@@
 
-\subsection{Simple illustratve case}
+\subsection{Simple illustrative case}
 
@@
 The latter choice is made since in this model formulation storage is not able to contribute to operating reserve requirements, complicating the interpretation of the results.
-Day 319 is chosen since resource use is not as `tight' as the other days and the absence of storage increases the probability of load or reserve shedding.
 
-A comparison of the real time operational security from the DUC-PR model compared to those obtained from a power flow performed by the team ULiege is given in the figure below.
+\subsection{Comparison with ULiege operational security results}
+
+A comparison of the real time operational security as percieved by the DUC-PR model compared to those obtained from a power flow performed by the team ULiege is given in the figure below.
+
+Clearly the two models disagree. We suggest the following non-exhaustive list of reasons why this could be:
+
+\begin{itemize}
+    \item The approximate modeling of real time imbalances in the DUC-PR model (see \secref{sec:operating_reserve_network_activation_constraints} and \secref{sec:further_constraining_operating_reserve_network_activation_constraints}).
+As extensively discussed, this means the DUC-PR model overestimates the ease with which real time imbalances may be dealt with.
+    \item The omission of N-1 contingencies in the DUC-PR model. This omission also leads to an overestimation of the ease of dealing with real time imbalances.
+    \item The extra-liberty for storage to deal with real time imbalances in ULiege's power flow model. This has the opposite effect.
+\end{itemize}
 

### 2022-11-22 scripts/copperplate_adequacy_security_tradeoff.jl

@@ -2,15 +2,14 @@ include(joinpath(@__DIR__, "..", "intro.jl"))
 sn = script_name(@__FILE__)
 mkrootdirs(plotsdir(sn))
 
-opts = options_diff_days(sn, "days_for_analysis.csv")[2]
+opts = options_diff_days(sn, "days_for_analysis.csv")[3]
 opts = merge(
     opts,
     Dict(
         "copperplate" => true,
-        "unit_commitment" => false,
-        # "downward_reserve_levels_included_in_redispatch" => 1:10,
-        # "upward_reserve_levels_included_in_redispatch" => 1:10,
+        "unit_commitment_type" => "none",
         "include_storage" => false,
+        "load_multiplier" => 1.5,
         "vars_2_save" => [:z, :q, :ls, :rL⁺, :rL⁻, :rsL⁺, :rsL⁻],
     ),
 )
@@ -21,7 +20,7 @@ opts_vec = [
             "reserve_shedding_limit" => rsl,
             "save_path" => opts["save_path"] * "_RSL=$rsl",
         ),
-    ) for rsl in 1.0:-0.5:0.0
+    ) for rsl in 1.0:-0.1:0.0
 ]
 gep_vec = run_GEPPR(opts_vec)
 rsl_vec = [opts_vec[i]["reserve_shedding_limit"] for i in 1:length(opts_vec)]
@@ -109,3 +108,5 @@ function plot_DUCPR_reserve_shedding_sensitivity(
 
     return nothing
 end
+
+plot_DUCPR_reserve_shedding_sensitivity(gep_vec, opts_vec, rsl_vec, ["161"])

### 2022-11-23 scripts/copperplate_adequacy_security_tradeoff.jl
Finished copperplate analysis

@@ -28,6 +28,7 @@ rsl_vec = [opts_vec[i]["reserve_shedding_limit"] for i in 1:length(opts_vec)]
 function plot_DUCPR_reserve_shedding_sensitivity(
     gep_vec, opts_vec, rsl_vec, days=["161"]
 )
+    @assert length(days) == 1 "This is horrific coding, apologies, but I can't have more than one day here."
     ls = Dict(
         rsl_vec[i] => try
             sum(gep_vec[i][:loadShedding].data; dims=(1, 2, 3))[:]
@@ -35,15 +36,30 @@ function plot_DUCPR_reserve_shedding_sensitivity(
             GEPPR.SVC(NaN)
         end for i in 1:length(opts_vec)
     )
-    rsLt = Dict(rsl_vec[i] => try
-        gep_vec[i][:rsL⁺]
-    catch
-        GEPPR.SVC(NaN)
-    end for i in 1:length(opts_vec))
+    scens = load_scenarios(first(opts_vec))
+    D⁺, D⁻, P⁺, P⁻, Dmid⁺, Dmid⁻ = scenarios_2_GEPPR(first(opts_vec), scens)
+    rsLt = Dict(
+        rsl_vec[i] => gep_vec[i][:rsL⁺, GEPPR.SVC(NaN)] for
+        i in 1:length(opts_vec)
+    )
+    T = GEPPR.get_set_of_time_indices(first(gep_vec))[3]
     rsL⁺ = Dict(
         rsl_vec[i] => try
             sum(
-                rsLt[(rsl_vec[i])][n, l, 1, 1, t]
+                rsLt[(rsl_vec[i])][n, l, 1, 1, t] *
+                P⁺[l, t - T[1] + 1] for
+                n in GEPPR.get_set_of_nodes(gep_vec[i]),
+                l in GEPPR.get_set_of_upward_reserve_levels(gep_vec[i]),
+                t in GEPPR.get_set_of_time_indices(gep_vec[i])[3]
+            )
+        catch
+            NaN
+        end for i in 1:length(opts_vec)
+    )
+    rsLexp = Dict(
+        rsl_vec[i] => try
+            sum(
+                rsLt[(rsl_vec[i])][n, l, 1, 1, t] * P⁺[l, t]
                 # rsLt[(rsl_vec[i])][n, l, 1, 1, t] *
                 # P⁺[(rsl_vec[i])][l, 1, 1, t] 
[…]

### 2022-11-23 papers/main.tex
Include comparison of Uliege model results

@@
 The latter choice is made since in this model formulation storage is not able to contribute to operating reserve requirements, complicating the interpretation of the results.
+The results are shown in \figref{fig:simple_reserve_vs_load_shedding}.
 
-\subsection{Comparison with ULiege operational security results}
+\begin{figure}[h]
+    \centering
+    \begin{subfigure}[t]{0.9\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{../plots/copperplate_adequacy_security_tradeoff/load_shedding_vs_reserve_shedding}
+        \caption{Reserve shedding plotted as a function of load shedding}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.9\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{../plots/copperplate_adequacy_security_tradeoff/load_shedding_vs_expected_load_shedding}
+        \caption{Expected additional real time load shedding (reserve shedding weighted by activation probability) plotted as a function of load shedding.}
+    \end{subfigure}
+\caption{Simple illustration of the potential trade-off between planned load shedding (i.e.
+day ahead adequacy) and expected load shedding (i.e.
+real time operational security).
+Results are for day 161, no storage, unit commitment or network constraints are included.\label{fig:simple_reserve_vs_load_shedding}}
+\end{figure}
+
+\figref{fig:simple_reserve_vs_load_shedding} is included here to illustrate the trade-off we wanted to be able to investigate.
+However, for the reasons described in \secref{sec:initial_results} and \bref{sec:ULiege_comparison}, this was not possible.
+
+\subsection{Comparison with ULiege operational security results} \label{sec:ULiege_comparison}
 
 A comparison of the real time operational security as percieved by the DUC-PR model compared to those obtained from a power flow performed by the team ULiege is given in the figure below.
+
+\begin{figure}[h]
+    \centering
+    \begin{subfigure}[t]{0.9\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{../plots/compare_ULiege_results/forecast_only_no_contingency.pdf}
+        \caption{Day ahead load shedding for DUC-PR and operational security power flow model.}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.9\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{../plots/compare_ULiege_results/with_scenarios_no_contingency.pdf}
+        \caption{Expected load shedding for DUC-PR and operational security power flow model.}
+    \end{subfigure}
+\caption{Results are for day 309, with storage, unit commitment and network (reserve activation) constraints.\label{fig:Uliege_compare}}
+\end{figure}
 
@@
 
+While the remark regarding storage could be resolved by allowing for storage to contribute to operating reserves as in \cite{Bruninx2016}, the remaining two issues are non-trivial to address in the DUC-PR framework.
+A stochastic unit commitment model, in which real time imbalances are fully defined within a scenario, would lend itself more easily to addressing these issues while coming at considerable computational cost.
+
 \clearpage

### 2022-11-23 scripts/copperplate_adequacy_security_tradeoff.jl

@@ -46,8 +46,7 @@ function plot_DUCPR_reserve_shedding_sensitivity(
     rsL⁺ = Dict(
         rsl_vec[i] => try
             sum(
-                rsLt[(rsl_vec[i])][n, l, 1, 1, t] *
-                P⁺[l, t - T[1] + 1] for
+                rsLt[(rsl_vec[i])][n, l, 1, 1, t] for
                 n in GEPPR.get_set_of_nodes(gep_vec[i]),
                 l in GEPPR.get_set_of_upward_reserve_levels(gep_vec[i]),
                 t in GEPPR.get_set_of_time_indices(gep_vec[i])[3]
@@ -59,7 +58,7 @@ function plot_DUCPR_reserve_shedding_sensitivity(
     rsLexp = Dict(
         rsl_vec[i] => try
             sum(
-                rsLt[(rsl_vec[i])][n, l, 1, 1, t] * P⁺[l, t]
+                rsLt[(rsl_vec[i])][n, l, 1, 1, t] * P⁺[l, t - T[1] + 1]
                 # rsLt[(rsl_vec[i])][n, l, 1, 1, t] *
                 # P⁺[(rsl_vec[i])][l, 1, 1, t] 
                 for n in GEPPR.get_set_of_nodes(gep_vec[i]),
@@ -81,6 +80,11 @@ function plot_DUCPR_reserve_shedding_sensitivity(
             reshape([rsL⁺[rsl_vec[i]] for i in 1:length(opts_vec)], nd, :)
         ),
     )
+    rsexp_mat = Matrix(
+        transpose(
+            reshape([rsLexp[rsl_vec[i]] for i in 1:length(opts_vec)], nd, :)
+        ),
+    )
     Plots.plot(
         ls_mat,
         rs_mat;
@@ -95,15 +99,17 @@ function plot_DUCPR_reserve_shedding_sensitivity(
 
     Plots.plot(
         ls_mat,
-        rs_mat .+ ls_mat;
+        rsexp_mat;
         lw=2,
         markerzise=16,
         markershape=:star5,
         xlab="Day ahead load shedding [MWh]",
-        ylab="Expected real time load shedding [MWh]",
-        lab=hcat(days...),
+        ylab="Additional real time expected load shedding [MWh]",
+        lab="",
+        legend=:topleft
     )
-    Plots.savefig(plotsdir(sn, "load_shedding_vs_expected_load__shedding.png"))
[…]

### 2022-11-23 papers/main.tex
I think I might be done?

@@
 \begin{abstract}
-    To be filled in
+    This report outlines an attempt to integrate a resource adequacy and operational security exercise.
+The idea is to use a unit commitment schedule which is then used as the input to a more detailed power flow model.
+An additional goal was to make explicit the trade-off between day ahead adequacy and real time reliability.
+We used an adapted version of KU Leuven's \ac{DUCPR} model which accounts for network constraints in the reserve activation phase for this purpose.
+However the \ac{DUCPR} model is not able to capture enough detail in the reserve activation phase for this tradeoff to be made.
+This and other omissions of the \ac{DUCPR} model when compared to the power flow model meant the results of the two models differed significantly.
 \end{abstract}
@@
 
+\section{Thoughts}
+
 \clearpage

### 2022-11-28 papers/main.tex
Added my reflections at the end

@@
 However the \ac{DUCPR} model is not able to capture enough detail in the reserve activation phase for this tradeoff to be made.
-This and other omissions of the \ac{DUCPR} model when compared to the power flow model meant the results of the two models differed significantly.
+This and other omissions of the \ac{DUCPR} model when compared to the power flow model meant the results of the two models differed.
+All code for the adequacy side of this exercise can be found at \href{https://github.com/junglegobs/ASoSEPOC}{https://github.com/junglegobs/ASoSEPOC}.
 \end{abstract}
@@
 
-\section{Thoughts}
+\section{Reflections on future work}
+
+Were an attempt at integrating an adequacy assessment and operational security analysis to be done again, it is my humble opinion that this should be done on a very simple system to begin with.
+Such a simple system would consist of only a few nodes (e.g.
+3) and without complicating technologies such as storage.
+A preliminary analysis could validate the agreement between the unit commitment and power flow model in an initial step, followed by increasing the complexity of the model and input data, e.g.
+including more forecast error scenarios and storage.
+As noted in this report, the \ac{DUCPR} model appears to be unable to account for real time nodal imbalances due to the way it models operating reserves (i.e.
+aggregating all forecast error scenarios over all nodes into one net load forecast error distribution for the entire system).
+A stochastic unit commitment would be therefore better suited for such an exercise, though this comes at the cost of dealing with the computational complexity of such a model.
 

### 2022-11-30 papers/main.tex
Made residual load plot, fixing full year ED

@@
 However the \ac{DUCPR} model is not able to capture enough detail in the reserve activation phase for this tradeoff to be made.
-This and other omissions of the \ac{DUCPR} model when compared to the power flow model meant the results of the two models differed.
+This and other omissions of the \ac{DUCPR} model when compared to the power flow model meant that the load shedding observed in the two models differed by an order of magntiude.
 All code for the adequacy side of this exercise can be found at \href{https://github.com/junglegobs/ASoSEPOC}{https://github.com/junglegobs/ASoSEPOC}.
@@
 Adequacy assessments and operational security analyses are two fundamental exercises in determining and ensuring the reliable operation of the electric power system.
-The former assesses the ability of the power system to satisfy demand for electricity in the day ahead stage, while the latter assesses (or ensures, depending on the context) that an unexpected event, such as an outage of a transmission line or generator or lower or greater than expected realisation of demand or renewable generation, does not lead to cascading failures and a possible blackout.
+The former assesses the ability of the power system to satisfy demand for electricity in the day ahead stage, while the latter assesses (or ensures, depending on the context) that an unexpected event, such as an outage of a transmission line or generator or lower or greater than expected realisation of demand or renewable generation, will lead to cascading failures and a possible blackout.
 
-Typically, these exercises are performed independently. This interaction combines the two by feeding the results of an adequacy assessment to an operational security analysis.
+Typically, these exercises are performed independently.
+This interaction combines the two by feeding the unit commitment of an adequacy assessment to an operational security analysis which takes these commitments as given.
 
@@
 
-Secondly, the adequacy assessment is done using an unconventional \ac{UC} model, a modified version of the \ac{DUCPR} model described in~\cite{Bruninx2017}.
+Secondly, we perform the adequacy assessment using an unconventional \ac{UC} model, a modified version of the \ac{DUCPR} model described in~\cite{Bruninx2017}.
 This model can be seen as a compromise between a computationally expensive stochastic \ac{UC} model and a more tractable though less accurate deterministic \ac{UC} model with operating reserve requirements.
-The treatment of operating reserves in this model allows a trade-off between shedding load in the day ahead stage, which would reduce day ahead adequacy, and shedding operating reserves, which should reduce operational security.
-By making this trade-off explicit and verifying the results in an operational security analysis we aim to elaborate on the results of~\cite{Hermans2018}, which did not consider the possibility that shedding operating reserves would lead to insecure operation of the electric power system.
+The treatment of operating reserves in this model in principle allows a trade-off between shedding load in the day ahead stage, which would reduce day ahead adequacy, and shedding operating reserves, which should reduce operational security.
+By making this trade-off explicit and verifying the results in an operational security analysis we wished to elaborate on the results of~\cite{Hermans2018}, which did not consider the possibility that shedding operating reserves would lead to insecure operation of the electric power system.
+
+As noted in the abstract, we were unsuccesful in illustrating the value of this interaction for two reasons.
+Firstly, we were unable to illustrate the tradeoff between adequacy and security with only the \ac{DUCPR} model due to its approximate treatment of real time network imbalances due to reserve activation.
+Secondly and related, the expected load shedding resulting from the \ac{DUCPR} differed by an order of magnitude to that found by the power flow, likely due to due to the difference in the treatment of storage as well as the more detailed network representation in the power flow model.
 
@@
 \secref{sec:data} describes the stylised Belgian system analysed.
-\secref{sec:initial_results} analyses a single day in order to identify sources of adequacy and security issues, while \secref{sec:adequacy_vs_security} analyses the trade-off between these two.
+\secref{sec:initial_results} analyses a single day in order to identify sources of adequacy and security issues, while \secref{sec:adequacy_vs_security} compares the \ac{DUCPR} model results to that of the power flow analysis.
 
 \section{Description of a Deterministic Unit Commitment Model with Probabilistic Reserve (DUC-PR) constraints} \label{sec:model}
-
-Sets, variables and costs related to unit commitment and storage operation have been omitted for brevity though they are included in the model.
-The interested reader is referred to \href{https://github.com/junglegobs/ASoSEPOC}{https://github.com/junglegobs/ASoSEPOC} for the relevant code.
 
@@
     \item $S$ - Storage technologies
+    \item $S$ - Storage technologies located at node $n$
     \item $N$ - Nodes or buses in the network
@@
 \begin{itemize}
-    \item $D_{nt}$ - Demand
-    \item $PTDF_{nl}$ - Power Transfer Distribution Factor
-    \item $AF_{gt}$ - Availability Factor
-    \item $K_g$ - Capacity
-    \item $P^{min}_g$ - Minimum power output
-    \item $P^{max}_g$ - Maximum power output
-    \item $MUT_g$ - Minimum up time
-    \item $MDT_g$ - Minimum down time
-    \item $RAMP_g$ - Maximum ramping limit (up and down)
-    \item $D^{L^+}_{lt}$ - Upward reserve requirement
-    \item $D^{L^-}_{lt}$ - Downward reserve requirement
-    \item $P^{L^+}_{lt}$ - Upward reserve level activation probability
-    \item $P^{L^-}_{lt}$ - Downward reserve level activation probability
+    \item $D_{nt}$ - Demand [MW]
+    \item $PTDF_{nl}$ - Power Transfer Distribution Factor [-]
+    \item $AF_{gt}$ - Availability Factor [-]
+    \item $K_g$ - Capacity [MW]
+    \item $P^{min}_g$ - Minimum power output [MW]
+    \item $P^{max}_g$ - Maximum power output [MW]
+    \item $MUT_g$ - Minimum up time [timesteps]
+    \item $MDT_g$ - Minimum down time [timesteps]
+    \item $RAMP_g$ - Maximum ramping limit (up and down) [MW/timestep]
+    \item $D^{L^+}_{lt}$ - Upward reserve requirement [MW]
+    \item $D^{L^-}_{lt}$ - Downward reserve requirement [MW]
+    \item $P^{L^+}_{lt}$ - Upward reserve level activation probability [-]
+    \item $P^{L^-}_{lt}$ - Downward reserve level activation probability [-]
 \end{itemize}
@@
 \begin{itemize}
-    \item $z_{gt}$ - Commitment
-    \item $v_{gt}$ - start up
-    \item $w_{gt}$ - Shut down
-    \item $q_{gt}$ - Generation
-    \item $\hat{q}_{gt}$ - Generation above the minimum stable operating point (0 in the case of renewables).
-    \item $d_{st}$ - Storage discharge
-    \item $c_{st}$ - Storage charge
-    \item $y_{st}$ - Binary variable, 1 if storage discharges and 0 if it charges
-    \item $e_{st}$ - Storage state of charge variable
-    \item $ls_{nt}$ - Load shedding
-    \item $inj_{nt}$ - Node injection
-    \item $f_{lt}$ - Branch flow
-    \item $r^+_{gt}$ - Total upward reserve provision
-    \item $r^-_{gt}$ - Total downward reserve provision
-    \item $r^{L^+}_{glt}$ - Upward reserve provision for reserve level $l$
-    \item $r^{L^-}_{glt}$ - Downward reserve provision for reserve level $l$
-    \item $rs_{nlt}$ - Upward reserve shedding for reserve level $l$
-    \item $rc_{nlt}$ - Downward reserve provided by day ahead load shedding for reserve level $l$
-    \item $rinj_{nlt}^{L^+}$ - Possible node injection due to activation of upward reserve level $l$
-    \item $rinj_{nlt}^{L^-}$ - Possible node injection due to activation of downward reserve level $l$
-    \item $rf_{nbt}^{L^+}$ - Possible branch flow due to activation of upward reserve level $l$
-    \item $rf_{nbt}^{L^-}$ - Possible branch flow due to activation of downward reserve level $l$
-    \item $d_{nlt}^{L^+}$ - Possible imbalance on node $n$ for upward reserve level $l$
-    \item $d_{nlt}^{L^-}$ - Possible imbalance on node $n$ for downward reserve level $l$
+    \item $z_{gt}$ - Commitment [-]
+    \item $v_{gt}$ - Start up [-]
+    \item $w_{gt}$ - Shut down [-]
+    \item $q_{gt}$ - Generation [MW]
+    \item $\hat{q}_{gt}$ - Generation above the minimum stable operating point (0 in the case of renewables) [MW]
+    \item $d_{st}$ - Storage discharge [MW]
+    \item $c_{st}$ - Storage charge [MW]
+    \item $y_{st}$ - Binary variable, 1 if storage discharges and 0 if it charges [-]
+    \item $e_{st}$ - Storage state of charge variable [MWh]
+    \item $ls_{nt}$ - Load shedding [MW]
+    \item $inj_{nt}$ - Node injection [MW]
+    \item $f_{lt}$ - Branch flow [MW]
+    \item $r^+_{gt}$ - Total upward reserve provision [MW]
+    \item $r^-_{gt}$ - Total downward reserve provision [MW]
+    \item $r^{L^+}_{glt}$ - Upward reserve provision for reserve level $l$ [MW]
+    \item $r^{L^-}_{glt}$ - Downward reserve provision for reserve level $l$ [MW]
+    \item $rs_{nlt}$ - Upward reserve shedding for reserve level $l$ [MW]
+    \item $rc_{nlt}$ - Downward reserve provided by day ahead load shedding for reserve level $l$ [MW]
+    \item $rinj_{nlt}^{L^+}$ - Possible node injection due to activation of upward reserve level $l$ [MW]
+    \item $rinj_{nlt}^{L^-}$ - Possible node injection due to activation of downward reserve level $l$ [MW]
+    \item $rf_{nbt}^{L^+}$ - Possible branch flow due to activation of upward reserve level $l$ [MW]
+    \item $rf_{nbt}^{L^-}$ - Possible branch flow due to activation of downward reserve level $l$ [MW]
+    \item $d_{nlt}^{L^+}$ - Possible imbalance on node $n$ for upward reserve level $l$ [MW]
+    \item $d_{nlt}^{L^-}$ - Possible imbalance on node $n$ for downward reserve level $l$ [MW]
 \end{itemize}
@@
 Note the use of the set $G_n$ to only allow generators at node $n$ to contribute to the power balance (similarly for storage technologies).
-Another way of describing this would have been through an incidence matrix.
+Another way of describing this would have been through an incidence matrix~\cite{Belderbos2020}.
+
+\subsubsection{Limiting day ahead load shedding}
 
@@
 While such unphysical behaviour is not typical, in this case the constraint was deemed necessary since such behaviour could be exploited to reduce congestion.
-This was also investigated in the results.
+This was also investigated in the results, where including this binary variable is referred to as PSCD (`Prevent Simultaneous Charge / Discharge').
 
@@
 
-At the time of writing, this convex hull restriction led to model infeasibilities and so it was not implemented.
+This convex hull restriction led to model infeasibilities and so it was not implemented.
 
@@
 
-All data used for this exercise can be found at: \href{https://www.dropbox.com/sh/mdvmc082gwng0tr/AABRyc3fZpxAFycmUfZmh8Csa?dl=0}.
+All data used for this exercise can be found \href{https://www.dropbox.com/sh/mdvmc082gwng0tr/AABRyc3fZpxAFycmUfZmh8Csa?dl=0}{here}.
 
@@
 
-The residual load time series, aggregated across all nodes, for this system is plotted in \todo[inline]{Plot this}.
+The residual load time series, aggregated across all nodes, for this system is plotted in \figref{fig:residual_load_ts}.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.9/textwidth]{../plots/process_input/residual_load_timeseries}
+    \caption{Residual load time series of test system.\label{fig:residual_load_ts}}
+\end{figure}
 
@@
 
-It should be noted that of the four days selected, only day 285 exhibited load shedding when the load was not scaled but operating reserves were present.
+Of the four days selected, only day 285 exhibited load shedding when the load was not scaled but operating reserves were present.
 This finding highlights how day ahead adequacy may occur for reasons other than insufficient capacity aggregated across the network, i.e.
 analysis of the aggregated residual load (the load net of renewable generation) may be uninformative.
+
+We changed the days under investigation several times during this interaction.
+For this reason (and also for brevity), we only analysed day 285 in detail (see \secref{sec:initial_results}) and only have a comparison of the \ac{DUCPR} and power flow results for day 309.
 

### 2022-11-30 scripts/process_input.jl

@@ -1,6 +1,7 @@
 include(joinpath(@__DIR__, "..", "intro.jl"))
 sn = script_name(@__FILE__)
 mkrootdirs(datadir("sims", sn))
+mkrootdirs(plotsdir(sn))
 
 # Process Excel sheets into a PowerModels.json format
 process_belderbos_data(grid_path)
@@ -17,9 +18,21 @@ opts = options(
     "operating_reserves_type" => "none",
     "prevent_simultaneous_charge_and_discharge" => false,
     "optimization_horizon" => [1, 8_760],
-    "save_path" => datadir("sims", "linear_w_storage"),
+    "rolling_horizon" => true,
+    "save_path" => datadir("sims", "linear_w_storage_v2"),
 )
 gep = run_GEPPR(opts)
+
+# Plot the residual load duration curve
+function plot_residual_load_curve(gep)
+    rl = demand_net_of_total_supply(gep)
+    rl = demand_net_of_total_supply(gep)
+    rl = sum(collect(values(rl)))
+    rl = sum(reshape(rl, N_HR_PER_DAY, :), dims=1)[:]
+    plt = Plots.plot(rl, xlab="Day of year", ylab="Residual load [MWh]", lab="")
+    Plots.savefig(plt, plotsdir(sn, "residual_load_timeseries.png"))
+end
+plot_residual_load_curve(gep)
 
 # Identify 4 days to investigate - no load shedding at all, almost load shedding and load shedding
 # Because I realised there were mistakes in the model later on, I save this to a "days_for_analysis_$(now()).csv" instead of "days_for_analysis.csv", so that I don't have to recreate all the scenarios.

### 2022-11-30 src/GEPPR.jl

@@ -165,7 +165,9 @@ function run_GEPPR(opts::Dict; load_only=false)
                 optimize_GEP_model!(gep)
                 save_optimisation_values!(gep)
             else
-                run_rolling_horizon(gep)
+                modify_network!(gep, opts)
+                modify_timeseries!(gep, opts)
+                run_rolling_horizon(gep; scheduleHorizon=168, slackVariables=(:sc, :sd, :z, :d⁺, :Δq, :v))
             end
             nothing
         end
@@ -562,7 +564,7 @@ function save_gep_for_security_analysis(gep::GEPM, path::String)
     e = gep[:e, SVC(missing)]
     sc = gep[:sc, SVC(missing)]
     sd = gep[:sd, SVC(missing)]
-    ls = gep[:loadShedding]
+    ls = gep[:loadShedding, gep[:ls, SVC(missing)] .+ gep[:lsel, SVC(missing)]]
     UC_results = Dict{Integer,Dict}()
     N, Y, P, T = GEPPR.get_set_of_nodes_and_time_indices(gep)
     GDN = GEPPR.get_set_of_nodal_dispatchable_generators(gep)

### 2022-11-30 scripts/process_input.jl
Fix RH full year ED

@@ -19,6 +19,7 @@ opts = options(
     "prevent_simultaneous_charge_and_discharge" => false,
     "optimization_horizon" => [1, 8_760],
     "rolling_horizon" => true,
+    "vars_2_save" => [:q, :sc, :sd, :e, :ls, :lsel],
     "save_path" => datadir("sims", "linear_w_storage_v2"),
 )
 gep = run_GEPPR(opts)
@@ -75,7 +76,7 @@ Plots.savefig(simsdir(sn, "load_shedding_daily_timeseries.png"))
 Plots.plot(sort(ls; rev=true); xlab="Day", ylab="Load shedding [MW]", lab="")
 Plots.savefig(simsdir(sn, "load_shedding_daily_timeseries_sorted.png"))
 
-# Save storage dispatch as a node injection
+# Save storage dispatch as a node injection, not used in the end
 storage_dispatch_2_node_injection(
     gep, grid_path, grid_w_store_ts_path, grid_wo_store_path
 )

### 2022-11-30 src/GEPPR.jl

@@ -167,7 +167,7 @@ function run_GEPPR(opts::Dict; load_only=false)
             else
                 modify_network!(gep, opts)
                 modify_timeseries!(gep, opts)
-                run_rolling_horizon(gep; scheduleHorizon=168, slackVariables=(:sc, :sd, :z, :d⁺, :Δq, :v))
+                run_rolling_horizon(gep; scheduleLength=168, slackVariables=(:sc, :sd, :z, :d⁺, :Δq, :v))
             end
             nothing
         end

### 2022-11-30 scripts/process_input.jl
Finish draft

@@ -25,7 +25,10 @@ opts = options(
 gep = run_GEPPR(opts)
 
 # Plot the residual load duration curve
-function plot_residual_load_curve(gep)
+function plot_residual_load_curve(opts)
+    opts_cp = deepcopy(opts)
+    opts_cp["load_multiplier"] = 1.0
+    gep = gepm(opts)
     rl = demand_net_of_total_supply(gep)
     rl = demand_net_of_total_supply(gep)
     rl = sum(collect(values(rl)))
@@ -33,7 +36,7 @@ function plot_residual_load_curve(gep)
     plt = Plots.plot(rl, xlab="Day of year", ylab="Residual load [MWh]", lab="")
     Plots.savefig(plt, plotsdir(sn, "residual_load_timeseries.png"))
 end
-plot_residual_load_curve(gep)
+plot_residual_load_curve(opts)
 
 # Identify 4 days to investigate - no load shedding at all, almost load shedding and load shedding
 # Because I realised there were mistakes in the model later on, I save this to a "days_for_analysis_$(now()).csv" instead of "days_for_analysis.csv", so that I don't have to recreate all the scenarios.

### 2022-11-30 papers/main.tex

@@
 We used an adapted version of KU Leuven's \ac{DUCPR} model which accounts for network constraints in the reserve activation phase for this purpose.
-However the \ac{DUCPR} model is not able to capture enough detail in the reserve activation phase for this tradeoff to be made.
+However, the presence of storage and the lack of detail captured by the the \ac{DUCPR} model in the reserve activation phase made it difficult to illustrate this tradeoff.
 This and other omissions of the \ac{DUCPR} model when compared to the power flow model meant that the load shedding observed in the two models differed by an order of magntiude.
-All code for the adequacy side of this exercise can be found at \href{https://github.com/junglegobs/ASoSEPOC}{https://github.com/junglegobs/ASoSEPOC}.
+All code for the adequacy aspect of this exercise can be found at \href{https://github.com/junglegobs/ASoSEPOC}{https://github.com/junglegobs/ASoSEPOC}.
 \end{abstract}
@@
 As noted in the abstract, we were unsuccesful in illustrating the value of this interaction for two reasons.
-Firstly, we were unable to illustrate the tradeoff between adequacy and security with only the \ac{DUCPR} model due to its approximate treatment of real time network imbalances due to reserve activation.
+Firstly, we were unable to illustrate the tradeoff between adequacy and security with only the \ac{DUCPR} model.
+This was partly because load shedding occurred at different moments to reserve shedding and possibly also due to the approximate treatment of real time network imbalances due to reserve activation in the \ac{DUCPR} model.
 Secondly and related, the expected load shedding resulting from the \ac{DUCPR} differed by an order of magnitude to that found by the power flow, likely due to due to the difference in the treatment of storage as well as the more detailed network representation in the power flow model.
@@
 \secref{sec:data} describes the stylised Belgian system analysed.
-\secref{sec:initial_results} analyses a single day in order to identify sources of adequacy and security issues, while \secref{sec:adequacy_vs_security} compares the \ac{DUCPR} model results to that of the power flow analysis.
+\secref{sec:initial_results} analyses a single day in order to identify the sources of adequacy and security issues, while \secref{sec:adequacy_vs_security} compares the \ac{DUCPR} model results to that of the power flow analysis.
 
@@
 Clearly it is simply not possible to have an imbalance on some nodes, such as Coo (the location of Belgium's pumped hydro unit).
-Constraining $d_{nlt}^{L^+}$ and $d_{nlt}^{L^-}$ in this way is referred to as AbsImb later on.
+Constraining $d_{nlt}^{L^+}$ and $d_{nlt}^{L^-}$ in this way is referred to as `AbsImb' later on.
 
@@
     \centering
-    \includegraphics[width=0.9/textwidth]{../plots/process_input/residual_load_timeseries}
+    \includegraphics[width=0.9\textwidth]{../plots/process_input/residual_load_timeseries}
     \caption{Residual load time series of test system.\label{fig:residual_load_ts}}
@@
 
+\section{Day ahead adequacy and real time operational security trade offs - a simple illustrative case} \label{sec:adequacy_vs_security}
+
+We begin by illustrating the tradeoff that exists between real time load shedding and day ahead reserve shedding with a simple model.
+No network or unit commitment constraints are included and no storage either.
+The latter choice is made since in this model formulation storage is not able to contribute to operating reserve requirements, complicating the interpretation of the results.
+The results are shown in \figref{fig:simple_reserve_vs_load_shedding}.
+
+\begin{figure}[ht]
+    \centering
+    \begin{subfigure}[t]{0.7\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{../plots/copperplate_adequacy_security_tradeoff/load_shedding_vs_reserve_shedding}
+        \caption{Reserve shedding plotted as a function of load shedding}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.7\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{../plots/copperplate_adequacy_security_tradeoff/load_shedding_vs_expected_load_shedding}
+        \caption{Expected additional real time load shedding (reserve shedding weighted by activation probability) plotted as a function of load shedding.}
+    \end{subfigure}
+\caption{Simple illustration of the potential trade-off between planned load shedding (i.e.
+day ahead adequacy) and expected load shedding (i.e.
+real time operational security).
+Results are for day 161, no storage, unit commitment or network constraints are included.\label{fig:simple_reserve_vs_load_shedding}}
+\end{figure}
+
+\figref{fig:simple_reserve_vs_load_shedding} is included here to illustrate the trade-off we wanted to be able to investigate.
+However, for the reasons we will describe in \secref{sec:initial_results} and \ref{sec:ULiege_comparison}, this was not possible.
+
 \section{Analysis of \acl{DUCPR} model results} \label{sec:initial_results}
 
-This analysis relates only to the 285th day of the year (12th of October). The aim is to investigate what are the causes of load shedding to better understand the system.
+This analysis relates only to the 285th day of the year (12th of October).
+The aim is to investigate what are the causes of load shedding to better understand the system.
+We will see that for this particular day, load shedding occurred due to constraint \ref{eq:storage_cyclic_constrant}.
+This meant that there was no trade-off to be made between load and reserve shedding, unless the load was multiplied by a factor of 1.5.
+The reserve network activation constraints also had a limited impact on the results.
 
@@
 While the objective increases by \euro 1,267,000 from the simple linear economic dispatch model with no network constraints to the unit commitment model with network constraints, no load shedding occurs.
-Preventing simultaneous charging and discharging constraints actually decreases, though this difference is within the optimality gap (0.01\%).
+Preventing simultaneous charging and discharging constraints actually decreases the objective function, though this difference is within the optimality gap (0.01\%).
 This suggests that additional energy consumption (which is not physically possible) from storage does not aid congestion in this context.
@@
 They also lead to reserve shedding where there wasn't any before.
-Neither day ahead network constraints or preventing simultaneous charging or discharging have an effect on the objective function which is greater than the MIP gap.
+Neither day ahead network constraints nor preventing simultaneous charging or discharging have an effect on the objective function greater than the MIP gap.
 
@@
 This is not due to congestion issues, since some of these storage units are located at the same node as the load being shed.
-Rather, this is likely due to the constraint which specifies that the storage state of charge at the end of the day must be greater than or equal to that at the start of the day.
+Rather, this is likely due to constraint \ref{eq:storage_cyclic_constrant} which specifies that the storage state of charge at the end of the day must be greater than or equal to that at the start of the day.
 This constraint was included so as not to overestimate the flexibility that storage could provide, though it is unlikely that such a situation would occur in reality.
@@
 This is counterintuitive, as you would expect the DUC-PR model to shed all operating reserves before shedding load in the day ahead stage.
-The former is less costly than the latter after all, due to activation probabilities being < 1.
+The former is less costly than the latter after all, due to activation probabilities being $< 1$.
 However, what is occuring is most likely `economic reserve shedding' - shedding reserves because they are highly unlikely to be activated.
@@
 That this effect is slight is for similar reasons to that listed previously - reserve shedding occurs in different hours to load shedding and load shedding occurs in the same hours as before (8:00 to 11:00) and is likely due to the cyclic constraint on storage.
+The absolute imbalance limits also appear to have a limited impact (less than a million euro increase in objective function) on the results.
 
@@
 
-\section{Day ahead adequacy and real time operational security trade offs} \label{sec:adequacy_vs_security}
+\section{Comparison of \acs{DUCPR} and power flow model results} \label{sec:ULiege_comparison}
 
-\subsection{Simple illustrative case}
-
-We begin by illustrating the tradeoff that exists between real time load shedding and day ahead reserve shedding with a simple model.
-No network or unit commitment constraints are included and no storage either.
-The latter choice is made since in this model formulation storage is not able to contribute to operating reserve requirements, complicating the interpretation of the results.
-The results are shown in \figref{fig:simple_reserve_vs_load_shedding}.
-
-\begin{figure}[h]
-    \centering
-    \begin{subfigure}[t]{0.9\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../plots/copperplate_adequacy_security_tradeoff/load_shedding_vs_reserve_shedding}
-        \caption{Reserve shedding plotted as a function of load shedding}
-    \end{subfigure}
-    \begin{subfigure}[t]{0.9\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../plots/copperplate_adequacy_security_tradeoff/load_shedding_vs_expected_load_shedding}
-        \caption{Expected additional real time load shedding (reserve shedding weighted by activation probability) plotted as a function of load shedding.}
-    \end{subfigure}
-\caption{Simple illustration of the potential trade-off between planned load shedding (i.e.
-day ahead adequacy) and expected load shedding (i.e.
-real time operational security).
-Results are for day 161, no storage, unit commitment or network constraints are included.\label{fig:simple_reserve_vs_load_shedding}}
-\end{figure}
-
-\figref{fig:simple_reserve_vs_load_shedding} is included here to illustrate the trade-off we wanted to be able to investigate.
-However, for the reasons described in \secref{sec:initial_results} and \bref{sec:ULiege_comparison}, this was not possible.
-
-\subsection{Comparison with ULiege operational security results} \label{sec:ULiege_comparison}
-
-A comparison of the real time operational security as percieved by the DUC-PR model compared to those obtained from a power flow performed by the team ULiege is given in the figure below.
+A comparison of the real time operational security as percieved by the \ac{DUCPR} model compared to those obtained from a power flow performed by the team ULiege is given in the figure below.
 
@@
         \includegraphics[width=\textwidth]{../plots/compare_ULiege_results/forecast_only_no_contingency.pdf}
-        \caption{Day ahead load shedding for DUC-PR and operational security power flow model.}
+        \caption{Day ahead load shedding for \ac{DUCPR} and operational security power flow model.}
     \end{subfigure}
@@
 
-Clearly the two models disagree. We suggest the following non-exhaustive list of reasons why this could be:
+Clearly the two models disagree and by some margin. We suggest the following non-exhaustive list of reasons for why this could be:
 
@@
     \item The omission of N-1 contingencies in the DUC-PR model. This omission also leads to an overestimation of the ease of dealing with real time imbalances.
-    \item The extra-liberty for storage to deal with real time imbalances in ULiege's power flow model. This has the opposite effect.
+    \item The extra-liberty for storage to deal with real time imbalances in ULiege's power flow, since the storage state of charge is not tracked and there are no cyclical constraints.
 \end{itemize}
